### 2020-01-17 fitbit_cleaning.Rmd
fitbit cleaning initial

---
title: "Fitbit cleaning"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

# Load files
```{r}
list_excel_files <- list.files("Input/fitbit_navajo_excel")
# Set working directory and load files in list; extract file name and add as column
## run next 4 lines together
setwd("Input/fitbit_navajo_excel")
initial_data <- tibble(file_name = list_excel_files) %>%
  mutate(save_data = lapply(file_name, read_excel, skip = 0)) %>%
  unnest(save_data)

list_csv_files <- list.files("Input/fitbit_navajo_csv")
# Set working directory and load files in list; extract file name and add as column
## run next 4 lines together
setwd("Input/fitbit_navajo_csv")
initial_data <- tibble(file_name = list_csv_files) %>%
  mutate(save_data = lapply(file_name, read_csv, skip = 1)) %>%
  unnest(save_data) 
```