### 2020-02-14 indoor_temp_rh_cleaning.Rmd
major data cleaning updates

---
title: "EldersAIR initial home/adult data work"
author: "Ethan Walker"
date: "Started 14 Feb 2020, Updated 14 Feb 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

# Load, format, and save indoor temp/rh data for NP
```{r}
## Box location: Update Box location once final data loaded
np_temp_rh_initial <- read_csv("Input/NP/np_qtrack_20200214.csv") %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         rh_indoor_current = as.numeric(RelHumCur),
         rh_indoor_max = as.numeric(RelHumMax),
         rh_indoor_min = as.numeric(RelHumMin),
         temp_indoor_current = as.numeric(TempCCur),
         temp_indoor_max = as.numeric(TempCMax),
         temp_indoor_min = as.numeric(TempCMin),
         sampling_visit = as.factor(SamplingVisit),
         area = "NP") %>% 
  # replace -999 and 0 values with NA
  replace_with_na(replace = list(rh_indoor_current = -999, rh_indoor_max = -999,
                                 rh_indoor_min = -999, temp_indoor_current = -999,
                                 temp_indoor_max = -999, temp_indoor_min = -999)) %>% 
  replace_with_na(replace = list(rh_indoor_current = 0, rh_indoor_max = 0,
                                 rh_indoor_min = 0, temp_indoor_current = 0,
                                 temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
  select(home_winter_id, area, rh_indoor_current, 
         rh_indoor_max, rh_indoor_min, temp_indoor_current, 
         temp_indoor_max, temp_indoor_min, sampling_visit) %>% 
  arrange(home_winter_id)
#write_rds(np_temp_rh_initial, "Output/np_temp_rh.rds")
```

# Load, format, and save indoor temp/rh data for NN
```{r}

```

# Join temp/rh data from different areas
```{r}
np_temp_rh <- read_rds("Output/np_temp_rh.rds")
nn_temp_rh <- read_rds("Output/nn_temp_rh.rds")
adult_indoor_temp_rh <- rbind(np_temp_rh, nn_temp_rh)
# write_rds(adult_indoor_temp_rh, "Output/adult_indoor_temp_rh.rds")
```

# Clean up all temp/rh data
```{r}
adult_indoor_temp_rh <- read_rds("Output/adult_indoor_temp_rh.rds")

kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
indoor_temp_rh_clean <- kids_indoor_temp_rh %>% 
  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
# write_rds(indoor_temp_rh_clean, "Output/indoor_temp_rh_clean.rds")
```