### 2020-02-25 stove_usage_cleaning.Rmd

---
title: "EldersAIR initial stove grades/usage data work"
author: "Ethan Walker"
date: "Started 25 Feb 2020, Updated 25 Feb 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

# Load, format, and save stove grades data
```{r}
## Box location: Update Box location once final data loaded
np_stove_initial <- read_xlsx("Input/NP/np_stove_20200225.xlsx") %>% 
  # rename variables
  mutate(home_id = as.factor(HomeID),
         home_winter_id = as.factor(HomeWinterID),
         stove_grade = as.factor(Score1),
         area = "NP") %>% 
  select(home_id, home_winter_id, area, stove_grade) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

## Box location: Update Box location once final data loaded
nn_stove_initial <- read_xlsx("Input/NN/nn_stove_20200225.xlsx") %>% 
  # rename variables
  mutate(home_id = as.factor(HomeID),
         home_winter_id = as.factor(HomeWinterID),
         stove_grade = as.factor(Score1),
         area = "NN") %>% 
  select(home_id, home_winter_id, area, stove_grade) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

# Combine locations and save
elders_stove_grades <- rbind(np_stove_initial, nn_stove_initial)

#write_rds(elders_stove_grades, "Output/elders_stove_grades.rds")
```

# Clean up all stove grades data
```{r}
elders_stove_grades <- read_rds("Output/elders_stove_grades.rds")

elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")

elders_stove_grades_clean <- elders_stove_grades %>% 
  left_join(elders_ids_linked, by = c("area", "home_id", "home_winter_id")) %>% 
  filter(!is.na(home_winter_id))

# write_rds(elders_stove_grades_clean, "Output/elders_stove_grades_clean.rds")

elders_stove_grades_clean <- read_rds("Output/elders_stove_grades_clean.rds")
```

##########################################

# Load, format, and save stove usage data
```{r}
## Box location: Update Box location once final data loaded
np_stove_initial <- read_xlsx("Input/NP/np_woodstoveusage_20200225.xlsx") %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         burn_level = as.factor(Burning),
         wood_type = as.factor(Wood),
         wood_type_other = as.character(WhOtherWood),
         sampling_visit = as.factor(SamplingVisit),
         area = "NP") %>% 
  select(home_winter_id, area, sampling_visit, burn_level, wood_type, wood_type_other) %>% 
  arrange(home_winter_id, sampling_visit)

## Box location: Update Box location once final data loaded
nn_stove_initial <- read_xlsx("Input/NN/nn_woodstoveusage_20200225.xlsx") %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         burn_level = as.factor(Burning),
         wood_type = as.factor(Wood),
         wood_type_other = as.character(WhOtherWood),
         sampling_visit = as.factor(SamplingVisit),
         area = "NN") %>% 
  select(home_winter_id, area, sampling_visit, burn_level, wood_type, wood_type_other) %>% 
  arrange(home_winter_id, sampling_visit)

# Combine locations and save
elders_stove_usage <- rbind(np_stove_initial, nn_stove_initial)

#write_rds(elders_stove_usage, "Output/elders_stove_usage.rds")
```

# Clean up all stove usage data
```{r}
elders_stove_usage <- read_rds("Output/elders_stove_usage.rds")

elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")

elders_stove_usage_clean <- elders_stove_usage %>% 
  left_join(elders_ids_linked, by = c("area", "home_winter_id")) %>% 
  filter(!is.na(burn_level))

# write_rds(elders_stove_usage_clean, "Output/elders_stove_usage_clean.rds")

elders_stove_usage_clean <- read_rds("Output/elders_stove_usage_clean.rds")
```