### 2020-03-06 indoor_temp_rh_summary.Rmd

---
title: "Elders indoor temp/RH summary"
author: "Ethan Walker"
date: "Started 5 March 2020, Updated 5 March 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(knitr)
```

```{r}
# Load data
indoor_temp_rh_clean <- read_rds("Output/elders_temp_rh_clean.rds")
elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
elders_pm <- read_rds("Output/elders_pm_clean.rds")
```

# Summary stats
```{r}
temp_rh_summary <- indoor_temp_rh_clean %>%
  filter(winter_id == 1) %>% 
  group_by(area) %>% 
  summarize_if(is.numeric, list(mean = mean, sd = sd, min = min, 
                                med = median, max = max), na.rm = TRUE) %>% 
  select(-area) %>% 
  t() 
kable(temp_rh_summary, col.names = c("Navajo", "Nez Perce"), 
      digits = 1, align = "c")
```

```{r, fig.height=6, fig.width=5}
# Boxplot function
boxplots_temp_rh_function <- function(data, variable){

boxplots_temp_rh <- data %>% 
  filter(winter_id == 1) %>% 
  select(area, variable) %>% 
  rename(plot_var = variable) %>% 
  na.exclude() %>% 
  ggplot(aes(area, plot_var)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
    theme_minimal() +
    theme(axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank())
boxplots_temp_rh

}
```

## Current indoor humidity (%)
```{r}
boxplots_temp_rh_function(indoor_temp_rh_clean, "rh_indoor_current")
```

## Max indoor humidity (%)
```{r}
boxplots_temp_rh_function(indoor_temp_rh_clean, "rh_indoor_max")
```

## Min indoor humidity (%)
```{r}
boxplots_temp_rh_function(indoor_temp_rh_clean, "rh_indoor_min")
```

## Current indoor temperature (C)
```{r}
boxplots_temp_rh_function(indoor_temp_rh_clean, "temp_indoor_current")
```

## Max indoor temperature (C)
```{r}
boxplots_temp_rh_function(indoor_temp_rh_clean, "temp_indoor_max")
```

## Min indoor temperature (C)
```{r}
boxplots_temp_rh_function(indoor_temp_rh_clean, "temp_indoor_min")
```

#### Abstract/manuscript tables ####

# Filter for homes that have PM data
```{r}
# 93 Winter 1 placebo homes have PM data
pm_homes_placebo <- pm_clean %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
  filter(initial_obs == 1) %>% 
  mutate(pm_home = "true") %>% 
  select(area, home, home_winter_id, pm_home) %>% 
  ungroup() %>% 
  arrange(area, home)

temp_rh_filtered <- indoor_temp_rh_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  ungroup() %>% 
  arrange(area, home)
```

# Summarize home data
```{r}
# Summarize numeric vars
kids_temp_rh <- temp_rh_filtered %>%  
  distinct(home, .keep_all = TRUE) %>%  
  group_by(area) %>% 
  summarize("Mean" = mean(rh_indoor_min, na.rm = TRUE), 
            "SD" = sd(rh_indoor_min, na.rm = TRUE),
            "min" = min(rh_indoor_min, na.rm = TRUE), 
            "med" = median(rh_indoor_min, na.rm = TRUE), 
            "max" = max(rh_indoor_min, na.rm = TRUE))
kable(kids_temp_rh, digits = 1, align = "c")

```