### 2020-03-06 stove_grade_usage_summary.Rmd

---
title: "Elders stove grades and usage summary"
author: "Ethan Walker"
date: "Started 5 March 2020, Updated 5 March 2020"
output: pdf_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

```{r}
# Load data
stove_grades_clean <- read_rds("Output/elders_stove_grades_clean.rds")
stove_use_clean <- read_rds("Output/elders_stove_usage_clean.rds")
elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
elders_pm <- read_rds("Output/elders_pm_clean.rds")
```

############ Use only most recent stove grades for analysis ###############

```{r, eval=FALSE}
# Save this for potential future use

# Summarize factor vars
## Use this as a quick look at all vars
table_list <- stove_use_clean %>% 
  #filter(area == "WMT") %>% 
  filter(winter_id == 1) %>%  
  select(burn_level, wood_type, wood_type_other) 
lapply(table_list, table)
```

# Summary of stove grades
```{r}
## Look further into NA values? ##
stove_grade_summary <- stove_grades_clean %>% 
  #filter(winter_id == 1 & treatment == "Placebo") %>% 
  select(area, home, stove_grade, winter_id) %>% 
  #filter(!is.na(stove_grade)) %>% 
  group_by(area, winter_id, stove_grade) %>% 
  distinct(home) %>% 
  summarize(n()) %>% 
  filter(!is.na(stove_grade))
kable(stove_grade_summary, align = "c")

stove_grades_total <- stove_grades_clean %>% 
  filter(!is.na(stove_grade)) %>% 
  filter(!is.na(home)) %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  select(area, home, stove_grade) %>% 
  arrange(area, home) 
```

```{r}
multiple_grades <- stove_grades_clean %>% 
  select(area, home, stove_grade, winter_id) %>% 
  filter(winter_id == 1 | winter_id == 2) %>% 
  filter(!is.na(stove_grade)) %>%
  group_by(home) %>% 
  filter(n() > 1) %>% 
  arrange(home, winter_id)
kable(multiple_grades)
```

# Summary of stove use
```{r}
# Summarize numeric vars
stove_use_summary <- stove_use_clean %>% 
  filter(winter_id == 1) %>% 
  group_by(area) %>% 
  summarize("N" = n(),
            "Mean Cords" = mean(wood_cords, na.rm = TRUE),
            "SD Cords" = sd(wood_cords, na.rm = TRUE),
            "Min Cords" = min(wood_cords, na.rm = TRUE), 
            "Median Cords" = median(wood_cords, na.rm = TRUE), 
            "Max Cords" = max(wood_cords, na.rm = TRUE))
kable(stove_use_summary, digits = 1, align = "c")

## Select specific var to assess individually/group by area
table(stove_use_clean$burn_level, stove_use_clean$area, useNA = "ifany")
table(stove_use_clean$wood_type, stove_use_clean$area, useNA = "ifany")
table(stove_use_clean$wood_type_other, stove_use_clean$area, useNA = "ifany")
```

# Stove use summary plots
```{r}
stove_use_plot_function <- function(data, variable) {

stove_use_plot <- data %>% 
  filter(winter_id == 1) %>% 
  select(area, variable) %>% 
  rename(plot_var = variable) %>% 
  #na.exclude() %>% 
  ggplot() +
    geom_bar(aes(plot_var, fill = area)) +
    theme_minimal() +
    labs(title = variable) +
    theme(axis.text.x = element_text(size = 16, colour = "black", 
                                     angle = 40, vjust = 0.7, hjust = 0.7),
          axis.text.y = element_text(size = 16, colour = "black"),
          legend.text = element_text(size = 14, colour = "black"),
          legend.title = element_text(size = 16, colour = "black"),
          title = element_text(size = 18, colour = "black"),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(),
          panel.grid = element_blank()) +
    scale_fill_manual(values = jv_palette)
stove_use_plot

}

stove_use_plot_function(stove_use_clean, "burn_level")
stove_use_plot_function(stove_use_clean, "wood_type")
```

#### Abstract/manuscript tables ####

# Filter for homes that have PM data
```{r}
# 93 Winter 1 placebo homes have PM data
pm_homes_placebo <- pm_clean %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
  filter(initial_obs == 1) %>% 
  mutate(pm_home = "true") %>% 
  select(area, home, home_winter_id, pm_home) %>% 
  ungroup() %>% 
  arrange(area, home)

stove_use_filtered <- stove_use_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  ungroup() %>% 
  arrange(area, home)

stove_grades_filtered <- stove_grades_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  ungroup() %>% 
  arrange(area, home)
```

# Summarize home data
```{r}
# Summarize numeric vars
kids_use <- stove_use_filtered %>%  
  #distinct(home, .keep_all = TRUE) %>%  
  group_by(area) %>% 
  summarize("Mean" = mean(wood_cords, na.rm = TRUE), 
            "SD" = sd(wood_cords, na.rm = TRUE),
            "min" = min(wood_cords, na.rm = TRUE), 
            "med" = median(wood_cords, na.rm = TRUE), 
            "max" = max(wood_cords, na.rm = TRUE))
kable(kids_use, digits = 1, align = "c")

# Summarize factor vars
alaska_use <- stove_use_filtered %>%  
  #distinct(home, .keep_all = TRUE) %>% 
  filter(area == "AK") %>% 
  group_by(wood_collect_method) %>% 
  summarize(n = n(), percent = (n()/10)*100)
  #summarize(n = n(), percent = (n()/60)*100)
alaska_use

navajo_use <- stove_use_filtered %>%  
  #distinct(home, .keep_all = TRUE) %>% 
  filter(area == "NN") %>% 
  group_by(wood_collect_method) %>% 
  summarize(n = n(), percent = (n()/23)*100)
  #summarize(n = n(), percent = (n()/138)*100)
navajo_use

montana_use <- stove_use_filtered %>%  
  #distinct(home, .keep_all = TRUE) %>% 
  filter(area == "WMT") %>% 
  group_by(wood_collect_method) %>% 
  summarize(n = n(), percent = (n()/60)*100)
  #summarize(n = n(), percent = (n()/360)*100)
montana_use

kids_use <- rbind(alaska_use, navajo_use, montana_use)
kable(kids_use)

# stove grade table
stove_grades_filtered <- stove_grades_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  arrange(area, home, winter_id) %>% 
  group_by(home) %>% 
  mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade)) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>%
  filter(!is.na(stove_grade)) %>% 
  ungroup() %>% 
  arrange(area, home)

table(stove_grades_filtered$stove_grade, stove_grades_filtered$area)
```