### 2020-03-11 demographics_summary.Rmd
summary updates

---
title: "EldersAIR: demographics summary"
author: "Ethan Walker"
date: "Started 6 March 2020, Updated 10 March 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

# Load individual datasets
```{r}
elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
elders_pm <- read_rds("Output/elders_pm_clean.rds")
```

##### For Winter 1, there were 145 homes with demographic data (76 NN and 69 NP)
##### and 150 participants with demographic data (79 NN and 71 NP)
##### 5 homes had 2 participants (3 NN and 2 NP); all other homes had 1 participant

# Filter for homes that have PM data
```{r}
demographics_filtered <- elders_ids_linked %>% 
  filter(winter_id == 1) %>% 
  filter(!is.na(gender)) 
```

# Summarize demographics data
```{r}
# Summarize numeric vars: total_residents, residents_under18, age_nov1
elders_dems <- demographics_filtered %>%  
  #group_by(area) %>% 
  summarize("Mean" = mean(age_nov1, na.rm = TRUE), 
            "SD" = sd(age_nov1, na.rm = TRUE),
            "min" = min(age_nov1, na.rm = TRUE), 
            "max" = max(age_nov1, na.rm = TRUE))
kable(elders_dems, digits = 1, align = "c")

# summarize factor vars: gender, hispanic, race, education, income
total_dems <- demographics_filtered %>%  
  group_by(education) %>% 
  summarize(n = n(), percent = (n()/150)*100)
total_dems

nn_dems <- demographics_filtered %>%  
  filter(area == "NN") %>% 
  group_by(education) %>% 
  summarize(n = n(), percent = (n()/79)*100)
nn_dems

np_dems <- demographics_filtered %>%  
  filter(area == "NP") %>% 
  group_by(education) %>% 
  summarize(n = n(), percent = (n()/71)*100)
np_dems

## Select specific var to assess individually/grouped by area
table(demographics_filtered$education, demographics_filtered$area, useNA = "ifany")
```