### 2020-03-12 demographics_summary.Rmd
plot and summary updates

---
title: "EldersAIR: demographics summary"
author: "Ethan Walker"
date: "Started 6 March 2020, Updated 12 March 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

```{r}
# Load individual datasets

elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
elders_pm <- read_rds("Output/elders_pm_clean.rds")
```

## For Winter 1, there were 145 homes with demographic data (76 NN and 69 NP)
## and 150 participants with demographic data (79 NN and 71 NP)
## 5 homes had 2 participants (3 NN and 2 NP); all other homes had 1 participant

```{r}
# Filter for Winter 1; remove NAs

demographics_filtered <- elders_ids_linked %>% 
  filter(winter_id == 1) %>% 
  filter(!is.na(gender)) %>% 
  mutate(income = factor(income,
                         levels = c("Less than $20,000", "$20,000 to $29,999",
                                    "$30,000 to $39,999", "$40,000 to $49,999",
                                    "$50,000 to $74,999", "$75,000 to $99,999",
                                    "$100,000 or more")),
         education = factor(education,
                            levels = c("Less than high school",
                                       "High school diploma or GED",
                                       "Some college",
                                       "College degree")))
```

```{r}
# Demographics summary stats function

demographics_function <- function(data, var, title_label, y_label) {

  demographics_plot <- data %>% 
  filter(winter_id == 1) %>% 
  select(area, var) %>% 
  rename(plot_var = var) %>% 
  filter(!is.na(plot_var)) %>% 
  ggplot() +
    geom_bar(aes(plot_var, fill = area)) +
    theme_minimal() +
    theme(axis.text.x = element_text(size = 16, colour = "black", 
                                     angle = 40, vjust = 0.7, hjust = 0.7),
          axis.text.y = element_text(size = 16, colour = "black"),
          title = element_text(size = 18, colour = "black"),
          legend.text = element_text(size = 14, colour = "black"),
          legend.title = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16, colour = "black"),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(),
          panel.grid = element_blank()) +
    labs(title = title_label) +
    ylab(y_label) +
    scale_fill_manual(values = jv_palette)
  demographics_plot

}

demographics_function(demographics_filtered, "gender", "Participant sex", "count")
demographics_function(demographics_filtered, "hispanic", "Hispanic ethnicity", "count")
demographics_function(demographics_filtered, "race", "Participant race", "count")
demographics_function(demographics_filtered, "income", "Household income", "count")
demographics_function(demographics_filtered, "education", "Participant education", "count")
demographics_function(demographics_filtered, "total_residents", "Total residents", "count")
demographics_function(demographics_filtered, "residents_under18", "Residents under 18", "count")
```

```{r}
demo_summary_function <- function(data, var, title_label, y_label) {

  demo_summary <- data %>% 
    filter(winter_id == 1) %>% 
    #select(var) %>% 
    select(area, var) %>% 
    group_by(area) %>% 
    summarize_all(list(mean=mean, sd=sd, min=min, med=median, max=max), na.rm = TRUE)
  demo_summary <<- demo_summary

  demo_plot <- data %>% 
    filter(winter_id == 1) %>% 
    select(area, var) %>% 
    rename(plot_var = var) %>% 
    filter(!is.na(plot_var)) %>% 
    ggplot(aes(area, plot_var)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red",
                 position=position_nudge(x = -0.1, y = 0)) +
    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
                 size=7, position=position_nudge(x = -0.35, y = 0)) +
    stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
                 position=position_nudge(x = 0.1, y = 0)) +
    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
                 size=7, position=position_nudge(x = 0.35, y = 0)) +
    theme_minimal() +
    labs(title = title_label, subtitle = "Mean = Red, Median = Blue") +
    ylab(y_label) +
    theme(title = element_text(size = 20, colour = "black"),
          axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16, colour = "black"),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank()) 
  demo_plot

}

demo_summary_function(demographics_filtered, "age_nov1", "Age, Nov 1", "years")
demo_summary_function(demographics_filtered, "total_residents", "Total residents", "count")
demo_summary_function(demographics_filtered, "residents_under18", "Residents under 18", "count")
```

```{r, include=FALSE}
# Summarize demographics data

# Summarize numeric vars: total_residents, residents_under18, age_nov1
elders_dems <- demographics_filtered %>%  
  #group_by(area) %>% 
  summarize("Mean" = mean(age_nov1, na.rm = TRUE), 
            "SD" = sd(age_nov1, na.rm = TRUE),
            "min" = min(age_nov1, na.rm = TRUE), 
            "max" = max(age_nov1, na.rm = TRUE))
kable(elders_dems, digits = 1, align = "c")

# summarize factor vars: gender, hispanic, race, education, income
total_dems <- demographics_filtered %>%  
  group_by(education) %>% 
  summarize(n = n(), percent = (n()/150)*100)
total_dems

nn_dems <- demographics_filtered %>%  
  filter(area == "NN") %>% 
  group_by(education) %>% 
  summarize(n = n(), percent = (n()/79)*100)
nn_dems

np_dems <- demographics_filtered %>%  
  filter(area == "NP") %>% 
  group_by(education) %>% 
  summarize(n = n(), percent = (n()/71)*100)
np_dems

## Select specific var to assess individually/grouped by area
table(demographics_filtered$education, demographics_filtered$area, useNA = "ifany")
```