### 2020-03-12 stove_grade_usage_summary.Rmd

---
title: "Elders stove grades and usage summary"
author: "Ethan Walker"
date: "Started 5 March 2020, Updated 12 March 2020"
output: pdf_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

```{r}
# Load data
stove_grades_clean <- read_rds("Output/elders_stove_grades_clean.rds")
stove_use_clean <- read_rds("Output/elders_stove_usage_clean.rds")
elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
elders_pm <- read_rds("Output/elders_pm_clean.rds")
```

```{r, eval=FALSE}
# Save this for potential future use

# Summarize factor vars
## Use this as a quick look at all vars
table_list <- stove_use_clean %>% 
  filter(winter_id == 1) %>%  
  filter(!is.na(burn_level)) %>% 
  select(area, sampling_visit, burn_level, wood_type, wood_type_other) 
#lapply(table_list, table)

table(table_list$wood_type)
table(table_list$area, table_list$wood_type)
table(table_list$area, table_list$burn_level, table_list$sampling_visit)
```

```{r}
############ Use only most recent stove grades for analysis ###############

## Look further into NA values? ##
stove_grade_summary <- stove_grades_clean %>% 
  #filter(winter_id == 1 & treatment == "Placebo") %>% 
  select(area, home, stove_grade, winter_id) %>% 
  #filter(!is.na(stove_grade)) %>% 
  group_by(area, winter_id, stove_grade) %>% 
  distinct(home) %>% 
  summarize(n()) %>% 
  filter(!is.na(stove_grade))
kable(stove_grade_summary, align = "c")

stove_grades_total <- stove_grades_clean %>% 
  filter(!is.na(stove_grade)) %>% 
  filter(!is.na(home)) %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  select(area, home, stove_grade) %>% 
  arrange(area, home) 
```

```{r}
multiple_grades <- stove_grades_clean %>% 
  select(area, home, stove_grade, winter_id) %>% 
  filter(winter_id == 1 | winter_id == 2) %>% 
  filter(!is.na(stove_grade)) %>%
  group_by(home) %>% 
  filter(n() > 1) %>% 
  arrange(home, winter_id)
#kable(multiple_grades)
```

# Summary of stove use
```{r, include=FALSE}
table(stove_use_clean$wood_type, stove_use_clean$area, useNA = "ifany")
table(stove_use_clean$wood_type_other, stove_use_clean$area, useNA = "ifany")
```

# Stove use summary plots
```{r}
stove_use_plot_function <- function(data, variable, title_label, y_label) {

stove_use_plot <- data %>% 
  filter(winter_id == 1) %>% 
  select(area, variable) %>% 
  rename(plot_var = variable) %>% 
  filter(!is.na(plot_var)) %>% 
  ggplot() +
    geom_bar(aes(plot_var, fill = area)) +
    theme_minimal() +
    labs(title = variable) +
    theme(axis.text.x = element_text(size = 16, colour = "black", 
                                     angle = 40, vjust = 0.7, hjust = 0.7),
          axis.text.y = element_text(size = 16, colour = "black"),
          legend.text = element_text(size = 14, colour = "black"),
          legend.title = element_text(size = 16, colour = "black"),
          title = element_text(size = 18, colour = "black"),
          axis.title.y = element_text(size = 16, colour = "black"),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(),
          panel.grid = element_blank()) +
    labs(title = title_label) +
    ylab(y_label) +
    scale_fill_manual(values = jv_palette)
stove_use_plot

}

stove_use_plot_function(stove_use_clean, "burn_level", "Burn level", "count")
stove_use_plot_function(stove_use_clean, "wood_type", "Wood type", "count")
stove_use_plot_function(stove_grades_clean, "stove_grade", "Stove grade", "count")
```

```{r, include=FALSE}
# stove grade tables
stove_grades_filtered <- stove_grades_clean %>% 
  arrange(area, home, winter_id) %>% 
  group_by(home) %>% 
  #mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade)) %>% 
  #filter(winter_id == 1) %>%
  group_by(home, winter_id) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  #filter(!is.na(stove_grade)) %>% 
  ungroup() %>% 
  filter(!is.na(stove_grade)) %>% 
  arrange(area, home)

table(stove_grades_filtered$stove_grade)
table(stove_grades_filtered$stove_grade, stove_grades_filtered$area)
table(stove_grades_filtered$stove_grade, stove_grades_filtered$area, stove_grades_filtered$winter_id)
```