### 2020-06-09 elders_health_data_cleaning.Rmd

---
title: "Elders health and elder files data cleaning"
author: "Ethan Walker"
date: "Started 28 Feb 2020, Updated 28 Feb 2020"
output: pdf_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

# Load data
elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
```

################# At Home Log ###############

# Load and merge files from different locations
```{r}
nn_athomelog_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_athome_20200225.xlsx")) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         adult_id_num = as.character(AdultID),
         area = "NN",
         sampling_visit = as.numeric(SamplingVisit),
         at_home_day = as.factor(Day),
         at_home_7 = as.factor(AtHome1),
         at_home_8 = as.factor(AtHome2),
         at_home_9 = as.factor(AtHome3),
         at_home_10 = as.factor(AtHome4),
         at_home_11 = as.factor(AtHome5),
         at_home_12 = as.factor(AtHome6),
         at_home_13 = as.factor(AtHome7),
         at_home_14 = as.factor(AtHome8),
         at_home_15 = as.factor(AtHome9),
         at_home_16 = as.factor(AtHome10),
         at_home_17 = as.factor(AtHome11),
         at_home_18 = as.factor(AtHome12),
         at_home_19 = as.factor(AtHome13),
         at_home_20 = as.factor(AtHome14),
         at_home_21 = as.factor(AtHome15),
         at_home_22 = as.factor(AtHome16)) %>% 
  select(home_winter_id:at_home_22)

np_athomelog_initial <- read_xlsx(paste0(file_path, "Input/NP/np_athomelog_20200225.xlsx")) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         adult_id_num = as.character(AdultID),
         area = "NP",
         sampling_visit = as.numeric(SamplingVisit),
         at_home_day = as.factor(Day),
         at_home_7 = as.factor(AtHome1),
         at_home_8 = as.factor(AtHome2),
         at_home_9 = as.factor(AtHome3),
         at_home_10 = as.factor(AtHome4),
         at_home_11 = as.factor(AtHome5),
         at_home_12 = as.factor(AtHome6),
         at_home_13 = as.factor(AtHome7),
         at_home_14 = as.factor(AtHome8),
         at_home_15 = as.factor(AtHome9),
         at_home_16 = as.factor(AtHome10),
         at_home_17 = as.factor(AtHome11),
         at_home_18 = as.factor(AtHome12),
         at_home_19 = as.factor(AtHome13),
         at_home_20 = as.factor(AtHome14),
         at_home_21 = as.factor(AtHome15),
         at_home_22 = as.factor(AtHome16)) %>% 
  select(home_winter_id:at_home_22)

# Combine logs from all areas
elders_at_home_log <- rbind(nn_athomelog_initial, np_athomelog_initial) %>% 
  left_join(elders_ids_linked, c("area", "home_winter_id", "adult_id_num"))
# write_rds(elders_at_home_log, "Output/elders_at_home_log.rds")

save_var_names <- data.frame(t(elders_at_home_log)) %>% 
  rownames_to_column() %>% 
  select(rowname)

write_csv(save_var_names, paste0(file_path, "Output/at_home_var_names.csv"))
```

################# Health History ###############

# Load and merge files from different locations
```{r}
nn_healthhistory_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_healthhistory_20200225.xlsx")) %>% 
  rename_all(~str_replace_all(., "Date", "_date")) %>% 
  rename_all(tolower) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         adult_id_num = as.character(adultid),
         area = "NN",
         rheum_arth = ra,
         rheum_arth_date = ra_date) %>% 
  select(-id:-homewinterid, -ra, -ra_date, -whatmissing:-ignorereason) %>% 
  select(home_winter_id, adult_id_num, area, rheum_arth, rheum_arth_date,
         lupus:cigarspipes)

np_healthhistory_initial <- read_xlsx(paste0(file_path, "Input/NP/np_healthhistory_20200225.xlsx")) %>% 
  rename_all(~str_replace_all(., "Date", "_date")) %>% 
  rename_all(tolower) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         adult_id_num = as.character(adultid),
         area = "NP",
         rheum_arth = ra,
         rheum_arth_date = ra_date) %>% 
  select(-id:-homewinterid, -ra, -ra_date, -whatmissing:-ignoremissingdata) %>% 
  select(home_winter_id, adult_id_num, area, rheum_arth, rheum_arth_date,
         lupus:cigarspipes)

# Combine logs from all areas
elders_health_hx <- rbind(nn_healthhistory_initial, np_healthhistory_initial) %>% 
  left_join(elders_ids_linked, c("area", "home_winter_id", "adult_id_num"))
# write_rds(elders_health_hx, "Output/elders_health_hx.rds")

save_var_names <- data.frame(t(elders_health_hx)) %>% 
  rownames_to_column() %>% 
  select(rowname)

write_csv(save_var_names, paste0(file_path, "Output/health_hx_var_names.csv"))
```

################# Health Extra ###############

# Load and merge files from different locations
```{r}
nn_healthextra_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_healthextra_20200225.xlsx")) %>% 
  rename_all(tolower) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         adult_id_num = as.character(adultid),
         area = "NN",
         sampling_visit = samplingvisit) %>% 
  select( home_winter_id, adult_id_num, sampling_visit, area, cold2m:bronch2mdr)

np_healthextra_initial <- read_xlsx(paste0(file_path, "Input/NP/np_healthextra_20200225.xlsx")) %>% 
  rename_all(tolower) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         adult_id_num = as.character(adultid),
         area = "NP",
         sampling_visit = samplingvisit) %>% 
  select( home_winter_id, adult_id_num, sampling_visit, area, cold2m:bronch2mdr)

# Combine logs from all areas
elders_health_extra <- rbind(nn_healthextra_initial, np_healthextra_initial) %>% 
  left_join(elders_ids_linked, c("area", "home_winter_id", "adult_id_num"))
# write_rds(elders_health_extra, "Output/elders_health_extra.rds")

save_var_names <- data.frame(t(elders_health_extra)) %>% 
  rownames_to_column() %>% 
  select(rowname)

write_csv(save_var_names, paste0(file_path, "Output/health_extra_var_names.csv"))
```