### 2020-06-09 fitbit_cleaning.Rmd

---
title: "Fitbit cleaning"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
```

# Load files
```{r}
list_excel_files <- list.files(paste0(file_path, "Input/fitbit_navajo_excel"))
# Set working directory and load files in list; extract file name and add as column
## run next 4 lines together
setwd(paste0(file_path, "Input/fitbit_navajo_excel"))
initial_data <- tibble(file_name = list_excel_files) %>%
  mutate(save_data = lapply(file_name, read_excel, skip = 0)) %>%
  unnest(save_data)

list_csv_files <- list.files(paste0(file_path, "Input/fitbit_navajo_csv"))
# Set working directory and load files in list; extract file name and add as column
## run next 4 lines together
setwd(paste0(file_path, "Input/fitbit_navajo_csv"))
initial_data <- tibble(file_name = list_csv_files) %>%
  mutate(save_data = lapply(file_name, read_csv, skip = 1)) %>%
  unnest(save_data) 
```