### 2020-06-09 wood_moisture_cleaning.Rmd

---
title: "EldersAIR initial wood moisture content data work"
author: "Ethan Walker"
date: "Started 25 Feb 2020, Updated 25 Feb 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
```

# Load, format, and save wood moisture data
```{r}
## Box location: Update Box location once final data loaded
np_moisture_initial <- read_xlsx(paste0(file_path, "Input/NP/np_moisture_20200225.xlsx")) %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         moisture_date = ymd(CompletedDate),
         moisture_1 = as.numeric(Moisture1),
         moisture_2 = as.numeric(Moisture2),
         moisture_3 = as.numeric(Moisture3),
         moisture_ave = as.numeric(MoistureAve),
         moisture_split = as.factor(Split),
         area = "NP") %>% 
  # replace 0 and -999 values with NA
  replace_with_na(replace = list(moisture_1 = 0, moisture_2 = 0,
                                 moisture_3 = 0, moisture_ave = 0)) %>% 
  replace_with_na(replace = list(moisture_1 = -999, moisture_2 = -999,
                                 moisture_3 = -999)) %>% 
  # gather 3 moisture samples to calculate average
  gather("sample", "value", c("moisture_1":"moisture_3")) %>% 
  arrange(home_winter_id, moisture_date) %>% 
  group_by(home_winter_id, moisture_date) %>% 
  mutate(moisture_ave = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>% 
  select(home_winter_id, area, moisture_date, 
         moisture_1, moisture_2, moisture_3, 
         moisture_ave, moisture_split) %>% 
  replace_with_na(replace = list(moisture_ave = "NaN")) %>%
  filter(!is.na(moisture_ave)) %>% 
  arrange(home_winter_id, moisture_date)

## Box location: Update Box location once final data loaded
nn_moisture_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_moisture_20200225.xlsx")) %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         moisture_date = ymd(CompletedDate),
         moisture_1 = as.numeric(Moisture1),
         moisture_2 = as.numeric(Moisture2),
         moisture_3 = as.numeric(Moisture3),
         moisture_ave = as.numeric(MoistureAve),
         moisture_split = as.factor(Split),
         area = "NN") %>% 
  # replace 0 and -999 values with NA
  replace_with_na(replace = list(moisture_1 = 0, moisture_2 = 0,
                                 moisture_3 = 0, moisture_ave = 0)) %>% 
  replace_with_na(replace = list(moisture_1 = -999, moisture_2 = -999,
                                 moisture_3 = -999)) %>% 
  # gather 3 moisture samples to calculate average
  gather("sample", "value", c("moisture_1":"moisture_3")) %>% 
  arrange(home_winter_id, moisture_date) %>% 
  group_by(home_winter_id, moisture_date) %>% 
  mutate(moisture_ave = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>% 
  select(home_winter_id, area, moisture_date, 
         moisture_1, moisture_2, moisture_3, 
         moisture_ave, moisture_split) %>% 
  replace_with_na(replace = list(moisture_ave = "NaN")) %>%
  filter(!is.na(moisture_ave)) %>% 
  arrange(home_winter_id, moisture_date)

# Combine locations and save
elders_moisture <- rbind(np_moisture_initial, nn_moisture_initial)

#write_rds(elders_moisture, paste0(file_path, "Output/elders_moisture.rds"))
```

# Clean up all moisture data
```{r}
elders_moisture <- read_rds(paste0(file_path, "Output/elders_moisture.rds"))

elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))

elders_moisture_clean <- elders_moisture %>% 
  left_join(elders_ids_linked, by = c("area", "home_winter_id")) %>% 
  filter(!is.na(home_winter_id)) %>% 
  # lower range on instrument = 5%
  # therefore, changing values < 5 to 5
  mutate(moisture_1 = if_else(moisture_1 < 5, 5, moisture_1),
         moisture_2 = if_else(moisture_2 < 5, 5, moisture_2),
         moisture_3 = if_else(moisture_3 < 5, 5, moisture_3),
         moisture_ave = if_else(moisture_ave < 5, 5, moisture_ave))

# write_rds(elders_moisture_clean, paste0(file_path, "Output/elders_moisture_clean.rds"))

elders_moisture_clean <- read_rds(paste0(file_path, "Output/elders_moisture_clean.rds"))
```