### 2020-06-15 fitbit_cleaning.Rmd
update fitbit data cleaning

---
title: "Fitbit cleaning"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
               "#0072B2", "#D55E00", "#CC79A7")
```

```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

elders_micropem_log <- read_rds(paste0(file_path, "Output/elders_micropem_log.rds"))
```

# Load NN files
```{r}
list_excel_files <- list.files(paste0(file_path, "Input/NN/fitbit_nn_excel"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path, "Input/NN/fitbit_nn_excel"))
initial_nn_data1 <- tibble(file_name = list_excel_files) %>%
  mutate(save_data = lapply(file_name, read_excel, skip = 1,
                            col_names = c("date", "cals_burned", "steps",
                                          "distance", "floors", "min_sedentary",
                                          "min_light_act", "min_fairly_act",
                                          "min_very_act", "act_cals"),
                            col_types = c("text", "text", "text",
                                          "text", "text", "text",
                                          "text", "text", 
                                          "text", "text"))) %>%
  unnest(save_data) %>% 
  mutate(date = ymd(date))

list_csv_files <- list.files(paste0(file_path, "Input/NN/fitbit_nn_csv"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path, "Input/NN/fitbit_nn_csv"))
initial_nn_data2 <- tibble(file_name = list_csv_files) %>%
  mutate(save_data = lapply(file_name, read_csv, skip = 2,
                            col_names = c("date", "cals_burned", "steps",
                                          "distance", "floors", "min_sedentary",
                                          "min_light_act", "min_fairly_act",
                                          "min_very_act", "act_cals"),
                            cols(date = col_character(), 
                                 cals_burned = col_character(), 
                                 steps = col_character(),
                                 distance = col_character(), 
                                 floors = col_character(), 
                                 min_sedentary = col_character(),
                                 min_light_act = col_character(), 
                                 min_fairly_act = col_character(), 
                                 min_very_act = col_character(), 
                                 act_cals = col_character()))) %>%
  unnest(save_data) %>% 
  mutate(date = mdy(date))

nn_joined <- rbind(initial_nn_data1, initial_nn_data2)

nn_clean <- nn_joined %>% 
  mutate(date = ymd(date),
         cals_burned = gsub(",", "", cals_burned),
         cals_burned = as.numeric(cals_burned),
         steps = gsub(",", "", steps),
         steps = as.numeric(steps),
         distance = gsub(",", "", distance),
         distance = as.numeric(distance),
         floors = gsub(",", "", floors),
         floors = as.numeric(floors),
         min_sedentary = gsub(",", "", min_sedentary),
         min_sedentary = as.numeric(min_sedentary),
         min_light_act = gsub(",", "", min_light_act),
         min_light_act = as.numeric(min_light_act),
         min_fairly_act = gsub(",", "", min_fairly_act),
         min_fairly_act = as.numeric(min_fairly_act),
         min_very_act = gsub(",", "", min_very_act),
         min_very_act = as.numeric(min_very_act),
         act_cals = gsub(",", "", act_cals),
         act_cals = as.numeric(act_cals),
         area = "NN") %>% 
  rename(fb_file_name = file_name) 
```

# Load NP files
```{r}
list_excel_files <- list.files(paste0(file_path, "Input/NP/np_fitbit"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path, "Input/NP/np_fitbit"))
initial_np_data1 <- tibble(file_name = list_excel_files) %>%
  mutate(save_data = lapply(file_name, read_excel, skip = 1,
                            col_names = c("date", "cals_burned", "steps",
                                          "distance", "floors", "min_sedentary",
                                          "min_light_act", "min_fairly_act",
                                          "min_very_act", "act_cals"),
                            col_types = c("text", "text", "text",
                                          "text", "text", "text",
                                          "text", "text", 
                                          "text", "text"))) %>%
  unnest(save_data) %>% 
  mutate(date = ymd(date))

list_excel_files <- list.files(paste0(file_path, "Input/NP/np_fitbit_new"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path, "Input/NP/np_fitbit_new"))
initial_np_data2 <- tibble(file_name = list_excel_files) %>%
  mutate(save_data = lapply(file_name, read_excel, skip = 1,
                            col_names = c("date", "cals_burned", "steps",
                                          "distance", "floors", "min_sedentary",
                                          "min_light_act", "min_fairly_act",
                                          "min_very_act", "act_cals"),
                            col_types = c("date", "text", "text",
                                          "text", "text", "text",
                                          "text", "text", 
                                          "text", "text"))) %>%
  unnest(save_data)

np_joined <- rbind(initial_np_data1, initial_np_data2)

np_clean <- np_joined %>% 
  mutate(date = ymd(date),
         cals_burned = gsub(",", "", cals_burned),
         cals_burned = as.numeric(cals_burned),
         steps = gsub(",", "", steps),
         steps = as.numeric(steps),
         distance = gsub(",", "", distance),
         distance = as.numeric(distance),
         floors = gsub(",", "", floors),
         floors = as.numeric(floors),
         min_sedentary = gsub(",", "", min_sedentary),
         min_sedentary = as.numeric(min_sedentary),
         min_light_act = gsub(",", "", min_light_act),
         min_light_act = as.numeric(min_light_act),
         min_fairly_act = gsub(",", "", min_fairly_act),
         min_fairly_act = as.numeric(min_fairly_act),
         min_very_act = gsub(",", "", min_very_act),
         min_very_act = as.numeric(min_very_act),
         act_cals = gsub(",", "", act_cals),
         act_cals = as.numeric(act_cals),
         area = "NP") %>% 
  rename(fb_file_name = file_name)
```

```{r}
elders_fitbit_joined <- rbind(nn_clean, np_clean)

elders_fitbit_clean <- elders_fitbit_joined %>% 
  separate(fb_file_name, c("fb_file_name", "trash1"), sep = ".csv", remove = FALSE) %>%
  separate(fb_file_name, c("fb_file_name", "trash1"), sep = ".xlsx", remove = FALSE) %>%
  separate(fb_file_name, c("fb_file_name", "trash1"), sep = ".xls", remove = FALSE) %>%
  left_join(elders_micropem_log, by = c("area", "fb_file_name")) %>% 
  select(-trash1, -floors)

write_rds(elders_fitbit_clean, paste0(file_path, "Output/elders_fitbit_clean.rds"))
```