### 2020-08-10 home_adult_data_cleaning.Rmd

---
title: "EldersAIR initial home/adult data work"
author: "Ethan Walker"
date: "Started 10 Feb 2020, Updated 10 Aug 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
```

# Load, format, and save home and adult overview/demographic data 
```{r}
# Load files 
## Box location: Update Box location once final data loaded
np_adult_initial <- read_xlsx(paste0(file_path, "Input/NP/adult.xlsx"), skip = 1,
                               na = c("", "NULL"),
                              col_names = c("adult_id_num", "home_id_num", "adult_id_char",
                                            "hispanic", "age_nov", "race", "gender",
                                            "child_assent", "consent_date", "phi_date",
                                            "phi_dr", "comment", "missing", "incomplete",
                                            "enrolled", "ignore_missing", "ignore_reason"),
                              col_types = c("numeric", "numeric", "text",
                                            "text", "text", "text", "text", 
                                            "text", "date", "date",
                                            "text", "text", "text", "text", 
                                            "text", "text", "text")) %>% 
  mutate(adult_id_num = as.factor(adult_id_num),
         home_id_num = as.factor(home_id_num),
         adult_id_char = as.factor(adult_id_char),
         adult_id_char = gsub("_", "", adult_id_char),
         adult_id_char = gsub(" ", "", adult_id_char),
         consent_date = ymd(consent_date),
         area = "NPT") %>% 
  select(adult_id_num, home_id_num, area, adult_id_char, consent_date, phi_date) %>% 
  arrange(adult_id_num)

## Box location: Update Box location once final data loaded
nn_adult_initial <- read_xlsx(paste0(file_path, "Input/NN/adult.xlsx"), skip = 1,
                               na = c("", "NULL"),
                              col_names = c("adult_id_num", "home_id_num", "adult_id_char",
                                            "hispanic", "age_nov", "race", "gender",
                                            "child_assent", "consent_date", "phi_date",
                                            "phi_dr", "comment", "missing", "incomplete",
                                            "enrolled", "ignore_missing", "ignore_reason"),
                              col_types = c("numeric", "numeric", "text",
                                            "text", "text", "text", "text", 
                                            "text", "date", "date",
                                            "text", "text", "text", "text", 
                                            "text", "text", "text")) %>% 
  mutate(adult_id_num = as.factor(adult_id_num),
         home_id_num = as.factor(home_id_num),
         adult_id_char = as.factor(adult_id_char),
         adult_id_char = gsub("_", "", adult_id_char),
         adult_id_char = gsub(" ", "", adult_id_char),
         consent_date = ymd(consent_date),
         area = "NN") %>% 
  select(adult_id_num, home_id_num, area, adult_id_char, consent_date, phi_date) %>% 
  arrange(adult_id_num)

elders_adult <- rbind(np_adult_initial, nn_adult_initial)
summary(elders_adult)

# Load files 
## Box location: Update Box location once final data loaded
np_demographics_initial <- read_xlsx(paste0(file_path, "Input/NP/demographics.xlsx"),
                                      na = c("", "NULL")) %>% 
  rename_all(tolower) %>% 
  mutate(adult_id_num = as.factor(adultid),
         home_winter_id = as.factor(homewinterid),
         gender = factor(gender),
         hispanic = factor(hispanic),
         race = factor(race),
         income = factor(income),
         education = factor(education),
         total_residents = as.numeric(totalres),
         residents_under18 = as.numeric(undereighteen),
         age = as.numeric(agenov1),
         area = "NPT") %>% 
  select(adult_id_num, home_winter_id, area, gender, hispanic, race, income, 
         education, total_residents, residents_under18, age) %>% 
  arrange(adult_id_num)

## Box location: Update Box location once final data loaded
nn_demographics_initial <- read_xlsx(paste0(file_path, "Input/NN/demographics.xlsx"),
                                      na = c("", "NULL")) %>% 
  rename_all(tolower) %>% 
  mutate(adult_id_num = as.factor(adultid),
         home_winter_id = as.factor(homewinterid),
         gender = factor(gender),
         hispanic = factor(hispanic),
         race = factor(race),
         income = factor(income),
         education = factor(education),
         total_residents = as.numeric(totalres),
         residents_under18 = as.numeric(undereighteen),
         age = as.numeric(agenov1),
         area = "NN") %>% 
  select(adult_id_num, home_winter_id, area, gender, hispanic, race, income, 
         education, total_residents, residents_under18, age) %>% 
  arrange(adult_id_num)

elders_demographics <- rbind(np_demographics_initial, nn_demographics_initial)
summary(elders_demographics)

## Box location: Update Box location once final data loaded
np_homewinter_initial <- read_xlsx(paste0(file_path, "Input/NP/homewinter.xlsx"), 
                                   skip = 1, na = c("", "NULL"),
                              col_names = c("home_winter_id", "home_id_num", "winter_id",
                                            "intervention_date", "intervention_time", 
                                            "sampling_date", "sampling_time", 
                                            "pickup_date", "pickup_time", 
                                            "sampling_date2", "sampling_time2", 
                                            "pickup_date2", "pickup_time2"),
                              col_types = c("numeric", "numeric", "numeric",
                                            "date", "date", 
                                            "date", "date", 
                                            "date", "date", 
                                            "date", "date", 
                                            "date", "date")) %>% 
  separate(intervention_time, c("trash1", "intervention_time"), sep = " ", remove = FALSE) %>%
  separate(sampling_time, c("trash1", "sampling_time"), sep = " ", remove = FALSE) %>%
  separate(pickup_time, c("trash1", "pickup_time"), sep = " ", remove = FALSE) %>%
  separate(sampling_time2, c("trash1", "sampling_time2"), sep = " ", remove = FALSE) %>%
  separate(pickup_time2, c("trash1", "pickup_time2"), sep = " ", remove = FALSE) %>%
  mutate(home_id_num = as.factor(home_id_num),
         winter_id = as.factor(winter_id),
         home_winter_id = as.factor(home_winter_id),
         intervention_date = ymd(intervention_date),
         intervention_time = as.character(intervention_time),
         sampling_date = ymd(sampling_date),
         sampling_time = as.character(sampling_time),
         pickup_date = ymd(pickup_date),
         pickup_time = as.character(pickup_time),
         sampling_date2 = ymd(sampling_date2),
         sampling_time2 = as.character(sampling_time2),
         pickup_date2 = ymd(pickup_date2),
         pickup_time2 = as.character(pickup_time2),
         area = "NPT") %>% 
  select(home_id_num, winter_id, home_winter_id, area, intervention_date, intervention_time,
         sampling_date, sampling_time, pickup_date, pickup_time,
         sampling_date2, sampling_time2, pickup_date2, pickup_time2) %>% 
  filter(winter_id != 3) %>% 
  arrange(home_id_num, winter_id)

## Box location: Update Box location once final data loaded
nn_homewinter_initial <- read_xlsx(paste0(file_path, "Input/NN/homewinter.xlsx"), 
                                   skip = 1, na = c("", "NULL"),
                              col_names = c("home_winter_id", "home_id_num", "winter_id",
                                            "intervention_date", "intervention_time", 
                                            "sampling_date", "sampling_time", 
                                            "pickup_date", "pickup_time", 
                                            "sampling_date2", "sampling_time2", 
                                            "pickup_date2", "pickup_time2"),
                              col_types = c("numeric", "numeric", "numeric",
                                            "date", "date", 
                                            "date", "date", 
                                            "date", "date", 
                                            "date", "date", 
                                            "date", "date")) %>% 
  separate(intervention_time, c("trash1", "intervention_time"), sep = " ", remove = FALSE) %>%
  separate(sampling_time, c("trash1", "sampling_time"), sep = " ", remove = FALSE) %>%
  separate(pickup_time, c("trash1", "pickup_time"), sep = " ", remove = FALSE) %>%
  separate(sampling_time2, c("trash1", "sampling_time2"), sep = " ", remove = FALSE) %>%
  separate(pickup_time2, c("trash1", "pickup_time2"), sep = " ", remove = FALSE) %>%
  mutate(home_id_num = as.factor(home_id_num),
         winter_id = as.factor(winter_id),
         home_winter_id = as.factor(home_winter_id),
         intervention_date = ymd(intervention_date),
         intervention_time = as.character(intervention_time),
         sampling_date = ymd(sampling_date),
         sampling_time = as.character(sampling_time),
         pickup_date = ymd(pickup_date),
         pickup_time = as.character(pickup_time),
         sampling_date2 = ymd(sampling_date2),
         sampling_time2 = as.character(sampling_time2),
         pickup_date2 = ymd(pickup_date2),
         pickup_time2 = as.character(pickup_time2),
         area = "NN") %>% 
  select(home_id_num, winter_id, home_winter_id, area, intervention_date, intervention_time,
         sampling_date, sampling_time, pickup_date, pickup_time,
         sampling_date2, sampling_time2, pickup_date2, pickup_time2) %>% 
  filter(winter_id != 3) %>% 
  arrange(home_id_num, winter_id)

elders_homewinter <- rbind(np_homewinter_initial, nn_homewinter_initial)
summary(elders_homewinter)

## Box location: Update Box location once final data loaded
np_home_initial <- read_xlsx(paste0(file_path, "Input/NP/home.xlsx"), na = c("", "NULL")) %>% 
  rename_all(tolower) %>% 
  mutate(home_id_num = as.factor(homeid),
         home = as.factor(home),
         treatment = as.factor(txcondition),
         home = gsub("_", "", home),
         home = gsub(" ", "", home),
         area = "NPT") %>% 
  select(home_id_num, home, treatment, area)

## Box location: Update Box location once final data loaded
nn_home_initial <- read_xlsx(paste0(file_path, "Input/NN/home.xlsx"), na = c("", "NULL")) %>% 
  rename_all(tolower) %>% 
  mutate(home_id_num = as.factor(homeid),
         home = as.factor(home),
         treatment = as.factor(txcondition),
         home = gsub("_", "", home),
         home = gsub(" ", "", home),
         area = "NN") %>% 
  select(home_id_num, home, treatment, area)

elders_home <- rbind(np_home_initial, nn_home_initial)
```

# Combine adult, home, homewinter files
```{r}
elders_demographics_final <- elders_homewinter %>%
  left_join(elders_demographics, by = c("area", "home_winter_id")) %>% 
  left_join(elders_home, by = c("area", "home_id_num")) %>% 
  left_join(elders_adult, by = c("area", "home_id_num", "adult_id_num")) %>% 
  mutate(intervention_date = gsub("2014-03-14", "2017-03-14", intervention_date)) %>% 
  select(adult_id_num, adult_id_char, area, home, home_id_num, home_winter_id, 
         winter_id, treatment, intervention_date:phi_date) %>% 
  filter(!is.na(home)) %>% 
  mutate(income = factor(income,
                         levels = c("Less than $20,000", "$20,000 to $29,999",
                                    "$30,000 to $39,999", "$40,000 to $49,999",
                                    "$50,000 to $74,999", "$75,000 to $99,999",
                                    "$100,000 or more"),
                         labels = c("< $20,000", "$20,000 to $29,999",
                                    "$30,000 to $39,999", "$40,000 to $49,999",
                                    "$50,000 to $74,999", "$75,000 to $99,999",
                                    ">= $100,000")),
         education = factor(education,
                            levels = c("Less than high school", "High school diploma or GED",
                                       "Some college", "College degree"),
                            labels = c("< high school", "High school grad or GED",
                                       "Some college", "College degree"))) %>% 
  arrange(area, home, winter_id)

#summary(elders_demographics_final)

elders_ids_linked <- elders_demographics_final %>% 
  select(adult_id_num, adult_id_char, area, home, home_id_num, home_winter_id, 
         winter_id, treatment)

write_rds(elders_demographics_final, paste0(file_path, "Output/elders_demographics.rds"))
write_rds(elders_ids_linked, paste0(file_path, "Output/elders_ids_linked.rds"))
```