### 2020-08-24 pm_summary_dustrak.Rmd

---
title: "Elders PM Summary"
author: "Ethan Walker"
date: "Started 24 Feb 2020, Updated 24 Aug 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(readxl)
library(knitr)
library(lubridate)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
               "#0072B2", "#D55E00", "#CC79A7")
```

```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

# Load data
elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds")) 
```

# Summary stats
```{r}
summary_stats <- elders_pm %>% 
  group_by(area, winter_id) %>% 
  distinct(area, home, winter_id, sampling_visit, .keep_all = TRUE) %>% 
  summarize("n" = n(),
            "Mean PM" = mean(pm_mean_winter, na.rm = TRUE), 
            "SD PM" = sd(pm_mean_winter, na.rm = TRUE),
            "Min PM" = min(pm_mean_winter, na.rm = TRUE), 
            "25%" = quantile(pm_mean_winter, 0.25, na.rm = TRUE),
            "Median PM" = median(pm_mean_winter, na.rm = TRUE),
            "75%" = quantile(pm_mean_winter, 0.75, na.rm = TRUE),
            "Max PM" = max(pm_mean_winter, na.rm = TRUE),
            "Mean Sample Days" = mean(pm_sample_interval, na.rm = TRUE),
            "SD Sample Days" = sd(pm_sample_interval, na.rm = TRUE)) %>% 
   t() 
kable(summary_stats, digits = 1, align = "c")

pm_by_day <- elders_pm %>% 
  filter(winter_id == 1) %>% 
  mutate(day_of_week = factor(day_of_week,
                                 levels = c("Monday", "Tuesday", "Wednesday", "Thursday",
                                            "Friday", "Saturday", "Sunday"),
                                 labels = c("Mon", "Tues", "Wed", "Thur", "Fri", "Sat", "Sun"))) %>% 
  #filter(pm < 500) %>% 
  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
  group_by(day_of_week) %>% 
  summarize("Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "25%" = quantile(pm, 0.25, na.rm = TRUE),
            "Median PM" = median(pm, na.rm = TRUE),
            "75%" = quantile(pm, 0.75, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE))
pm_by_day

pm_by_month <- elders_pm %>% 
  filter(winter_id == 1) %>% 
  mutate(sampling_month = month(pm_datetime_new, abbr = TRUE)) %>%
  #filter(pm < 500) %>% 
  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
  group_by(sampling_month) %>% 
  summarize("Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "25%" = quantile(pm, 0.25, na.rm = TRUE),
            "Median PM" = median(pm, na.rm = TRUE),
            "75%" = quantile(pm, 0.75, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE))
pm_by_month
```

\pagebreak

```{r, fig.height=6, fig.width=5}
boxplots_elders_pm <- elders_pm %>% 
  #filter(winter_id == 1) %>% 
  distinct(area, home, winter_id, .keep_all = TRUE) %>% 
  ggplot(aes(area, pm_mean_winter)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red") +
    #scale_y_log10() +
    #ylim(0, 200) +
    theme_minimal() +
    labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
    theme(axis.text.x = element_text(size = 16, colour = "black",
                                      margin = margin(t = 0, r = 0, b = 30, l = 0)),
          axis.text.y = element_text(size = 16, colour = "black"),
          strip.text = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank()) +
    facet_wrap(.~winter_id)
boxplots_elders_pm

# ggsave("boxplots_elders_pm.jpg", height = 7, width = 4)
```

```{r}
# homes that exceed PM guidelines
pm_guidelines <- elders_pm %>% 
  filter(winter_id == 1) %>% 
  filter(pm_mean_winter > 12) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  group_by(area) %>% 
  count()
pm_guidelines
```

```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds")) 
elders_temp_rh_clean <- read_rds(paste0(file_path, "Output/elders_temp_rh_clean.rds"))

# Summary data for results letters to send to participants
home_summary <- elders_pm %>% 
  left_join(elders_temp_rh_clean, by = c("area", "home", "winter_id", "sampling_visit")) %>% 
  filter(winter_id == 1) %>% 
  group_by(area, home) %>% 
  mutate("Min RH (%)" = round(mean(rh_indoor_min, na.rm = TRUE), digits = 0),
         "Max RH (%)" = round(mean(rh_indoor_max, na.rm = TRUE), digits = 0),
         "Mean PM2.5 (ug/m3)" = round(pm_mean_winter), digits = 0) %>% 
  distinct(area, home, .keep_all = TRUE) %>% 
  select(area, home, winter_id, "Mean PM2.5 (ug/m3)", "Min RH (%)", "Max RH (%)") %>%
  rename(Area = area,
         Home = home,
         Winter = winter_id) %>% 
  replace_with_na(replace = list("Min RH (%)" = "NaN")) %>% 
  replace_with_na(replace = list("Max RH (%)" = "NaN"))

write_csv(home_summary, paste0(file_path, "Output/elders_letter_data.csv"), na = "")
```