### 2021-03-11 stove_usage_cleaning.Rmd

---
title: "EldersAIR initial stove grades/usage data work"
author: "Ethan Walker"
date: "Started 25 Feb 2020, Updated 25 Feb 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
```

# Load, format, and save stove grades data
```{r}
## Box location: Update Box location once final data loaded
npt_stove_initial <- read_xlsx(paste0(file_path, "Input/NP/stove.xlsx"),
                               na = "NULL") %>% 
  # rename variables
  mutate(home_id_num = as.factor(HomeID),
         home_winter_id = as.factor(HomeWinterID),
         stove_grade = as.factor(Score1),
         area = "NPT") %>% 
  select(home_id_num, home_winter_id, area, stove_grade) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

## Box location: Update Box location once final data loaded
nn_stove_initial <- read_xlsx(paste0(file_path, "Input/NN/stove.xlsx"),
                               na = "NULL") %>% 
  # rename variables
  mutate(home_id_num = as.factor(HomeID),
         home_winter_id = as.factor(HomeWinterID),
         stove_grade = as.factor(Score1),
         area = "NN") %>% 
  select(home_id_num, home_winter_id, area, stove_grade) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

# Combine locations and save
elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))

elders_stove_grades <- rbind(npt_stove_initial, nn_stove_initial) %>% 
  right_join(elders_ids_linked, by = c("area", "home_id_num", "home_winter_id")) %>% 
  arrange(area, home_id_num, winter_id) %>% 
  select(area, home_id_num, home, home_winter_id, adult_id_num:treatment, stove_grade) %>% 
  distinct(area, home, winter_id, .keep_all = T) %>% 
  arrange(area, home_id_num, stove_grade) %>% 
  distinct(area, home_id_num, .keep_all = T) %>% 
  select(area, home_id_num, home, treatment, stove_grade) %>% 
  mutate(area = as.factor(area),
         home = as.factor(home))

summary(elders_stove_grades)

# write_rds(elders_stove_grades, paste0(file_path, "Output/elders_stove_grades_clean.rds"))
```

##########################################

# Load, format, and save stove usage data
```{r}
## Box location: Update Box location once final data loaded
npt_stove_initial <- read_xlsx(paste0(file_path, "Input/NP/woodstoveusage.xlsx"),
                               na = "NULL") %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         burn_level = as.factor(Burning),
         wood_type = as.factor(Wood),
         wood_type_other = as.character(WhOtherWood),
         sampling_visit = as.factor(SamplingVisit),
         area = "NPT") %>% 
  select(home_winter_id, area, sampling_visit, burn_level, wood_type, wood_type_other) %>% 
  arrange(home_winter_id, sampling_visit)

## Box location: Update Box location once final data loaded
nn_stove_initial <- read_xlsx(paste0(file_path, "Input/NN/woodstoveusage.xlsx"),
                               na = "NULL") %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         burn_level = as.factor(Burning),
         wood_type = as.factor(Wood),
         wood_type_other = as.character(WhOtherWood),
         sampling_visit = as.factor(SamplingVisit),
         area = "NN") %>% 
  select(home_winter_id, area, sampling_visit, burn_level, wood_type, wood_type_other) %>% 
  arrange(home_winter_id, sampling_visit)

# Combine locations and save
elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
  distinct(area, home, winter_id, .keep_all = T) %>% 
  select(-adult_id_num, -adult_id_char)

elders_stove_usage <- rbind(npt_stove_initial, nn_stove_initial) %>% 
  right_join(elders_ids_linked, by = c("area", "home_winter_id")) %>% 
  arrange(area, home, winter_id, sampling_visit) %>% 
  select(area, home, home_id_num, home_winter_id, winter_id, treatment,
         sampling_visit:wood_type_other) %>% 
  mutate(area = as.factor(area),
         home = as.factor(home))

summary(elders_stove_usage)

# write_rds(elders_stove_usage, paste0(file_path, "Output/elders_stove_usage_clean.rds"))
```

#####################################

# Load, format, and KW monitor data
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

##### NP #####
# Load data
kw_initial_np <- read_xlsx(paste0(file_path, "Input/NP/filtretechange.xlsx"), 
                                  na = c("NULL", "", "...", -99, -999, -9999))

kw_np <- kw_initial_np %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         kw_date = ymd(ChangeDate),
         kw_hours = as.numeric(KWHours),
         area = "NPT") %>% 
  select(home_winter_id, area, kw_date, kw_hours) %>% 
  filter(!is.na(kw_hours)) %>% 
  ungroup() %>% 
  arrange(home_winter_id, kw_date)

# Save as RDS
write_rds(kw_np, paste0(file_path, "Output/kw_np.rds"))

##### NN #####
# Load data
kw_initial_nn <- read_xlsx(paste0(file_path, "Input/NN/filtretechange.xlsx"), 
                                  na = c("NULL", "", "...", -99, -999, -9999))

kw_nn <- kw_initial_nn %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         kw_date = ymd(ChangeDate),
         kw_hours = as.numeric(KWHours),
         area = "NN") %>% 
  select(home_winter_id, area, kw_date, kw_hours) %>% 
  filter(!is.na(kw_hours)) %>%  
  ungroup() %>% 
  arrange(home_winter_id, kw_date)

# Save as RDS
write_rds(kw_nn, paste0(file_path, "Output/kw_nn.rds"))
```

# Join KW hour data from different areas
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

kw_np <- read_rds(paste0(file_path, "Output/kw_np.rds"))
kw_nn <- read_rds(paste0(file_path, "Output/kw_nn.rds"))

elders_kw <- rbind(kw_np, kw_nn) %>% 
  ungroup() %>% 
  arrange(area, home_winter_id, kw_date)

write_rds(elders_kw, paste0(file_path, "Output/elders_kw.rds"))
```

# Clean up all KW hour data
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
elders_kw <- read_rds(paste0(file_path, "Output/elders_kw.rds"))
elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))

elders_kw_clean <- elders_kw %>% 
  left_join(elders_ids_linked, by = c("area", "home_winter_id")) %>%  
  filter(!is.na(home)) %>% 
  arrange(area, home_winter_id, kw_date) %>% 
  group_by(area, home_winter_id) %>% 
  mutate(kw_hours_last = last(kw_hours),
         diff_check = if_else(kw_hours_last < kw_hours, 1, 0),
         diff_check2 = kw_hours_last - kw_hours)  
  #distinct(area, home_winter_id, diff_check, .keep_all = TRUE)

summary(elders_kw_clean)

kw_summary <- elders_kw_clean %>% 
  select(-treatment_assigned, -treatment_actual) %>% 
  left_join(treatments_masked, by = "home") %>% 
  distinct(area, home, kw_hours_total) %>% 
  ungroup() %>% 
  filter(area == "NPT") %>% 
  #group_by(newtx) %>% 
  summarize(mean(kw_hours_total),
            sd(kw_hours_total),
            n(),
            min(kw_hours_total),
            median(kw_hours_total),
            max(kw_hours_total))
kw_summary

write_rds(elders_kw_clean, paste0(file_path, "Output/elders_kw_clean.rds"))
```

#####################################

# Load, format, and KW compliance report
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

##### NP #####
# Load data
kw_report_np <- read_csv(paste0(file_path, "Input/NP/filtrete_report.csv"), 
                                  na = c("NULL", "")) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         home = as.factor(Home),
         kw_date = ymd(ChangeDate),
         kw_hours = as.numeric(KWHours),
         kw_notes = as.character(Notes),
         hours_possible = as.numeric(Hours),
         kw_expected = as.numeric(Expected),
         kw_perc_expected = as.numeric(PercExp),
         filter_type = as.character(FType),
         area = "NPT") %>% 
  dplyr::select(area, home, home_winter_id, kw_date, kw_hours, kw_notes, filter_type, 
         hours_possible, kw_expected, kw_perc_expected) %>% 
  #filter(!is.na(kw_hours)) %>% 
  ungroup() %>% 
  arrange(home_winter_id, kw_date)

# Save as RDS
write_rds(kw_report_np, paste0(file_path, "Output/kw_report_np.rds"))

##### NN #####
# Load data
kw_report_nn <- read_csv(paste0(file_path, "Input/NN/filtrete_report.csv"), 
                                  na = c("NULL", "")) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         home = as.factor(Home),
         kw_date = ymd(ChangeDate),
         kw_hours = as.numeric(KWHours),
         kw_notes = as.character(Notes),
         hours_possible = as.numeric(Hours),
         kw_expected = as.numeric(Expected),
         kw_perc_expected = as.numeric(PercExp),
         filter_type = as.character(FType),
         area = "NN") %>% 
  dplyr::select(area, home, home_winter_id, kw_date, kw_hours, kw_notes, filter_type, 
         hours_possible, kw_expected, kw_perc_expected) %>% 
  #filter(!is.na(kw_hours)) %>% 
  ungroup() %>% 
  arrange(home_winter_id, kw_date)

# Save as RDS
write_rds(kw_report_nn, paste0(file_path, "Output/kw_report_nn.rds"))
```

# Join KW report data from different areas
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

kw_report_np <- read_rds(paste0(file_path, "Output/kw_report_np.rds"))
kw_report_nn <- read_rds(paste0(file_path, "Output/kw_report_nn.rds"))

elders_kw_report <- rbind(kw_report_np, kw_report_nn) %>% 
  dplyr::select(-home) %>% 
  ungroup() %>% 
  arrange(area, home_winter_id, kw_date)

write_rds(elders_kw_report, paste0(file_path, "Output/elders_kw_report.rds"))
```

# Clean up all KW report data
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
elders_kw_report <- read_rds(paste0(file_path, "Output/elders_kw_report.rds"))
elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))

elders_kw_report_clean <- elders_kw_report %>% 
  left_join(elders_ids_linked, by = c("area", "home_winter_id")) %>%  
  arrange(area, home_winter_id) %>% 
  mutate(filter_type = fct_collapse(filter_type,
                                    "Honeywell" = "Honeywell",
                                    "Large Filtrete" = "Large Filtrete",
                                    "Small Filtrete" = "Small Filtrete",
                                    "Winix" = c("Winix", "winix"))) %>% 
  mutate(kw_perc_expected = if_else(kw_hours == -999, 999999, kw_perc_expected),
         kw_perc_expected = if_else(kw_hours == 1000, 999999, kw_perc_expected),
         kw_perc_expected = abs(kw_perc_expected)) %>% 
  replace_with_na(replace = list(kw_perc_expected = 999999)) %>% 
  mutate(kw_perc_expected = as.numeric(kw_perc_expected)) 
  #filter(treatment_assigned == "Filter")

#summary(kids_kw_report_clean)

kw_perc_exp_summary <- elders_kw_report_clean %>% 
  filter(!is.na(kw_perc_expected)) %>% 
  #filter(area == "WMT") %>% 
  group_by(treatment) %>% 
  summarize("Mean KW % Exp" = mean(kw_perc_expected),
            "SD KW % Exp" = sd(kw_perc_expected),
            "N KW % Exp" = n(),
            "Min KW % Exp" = min(kw_perc_expected),
            "25Q KW % Exp" = quantile(kw_perc_expected, 0.25),
            "Med KW % Exp" = median(kw_perc_expected),
            "75Q KW % Exp" = quantile(kw_perc_expected, 0.75),
            "Max KW % Exp" = max(kw_perc_expected))
kw_perc_exp_summary

png("Kids_KW_perc_exp.png", width=960, height=480, bg = "white")
grid.table(kw_perc_exp_summary)
dev.off()

boxplot(log(kw_perc_expected) ~ filter_type, data = kids_kw_report_clean)

boxplot(kw_perc_expected ~ filter_type, data = kids_kw_report_clean)

kw_exp_summary <- kids_kw_report_clean %>% 
  filter(!is.na(kw_expected)) %>% 
  #filter(area == "WMT") %>% 
  group_by(filter_type) %>% 
  summarize("Mean KW Exp" = mean(kw_expected),
            "SD KW Exp" = sd(kw_expected),
            "N KW Exp" = n(),
            "Min KW Exp" = min(kw_expected),
            "25Q KW Exp" = quantile(kw_expected, 0.25),
            "Med KW Exp" = median(kw_expected),
            "75Q KW Exp" = quantile(kw_expected, 0.75),
            "Max KW Exp" = max(kw_expected))
kw_exp_summary

fivenum(kids_kw_report_clean$kw_perc_expected)

write_rds(elders_kw_report_clean, paste0(file_path, "Output/elders_kw_report_clean.rds"))
```

# Join KW data with PM data and summarize
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
pm_data <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds"))  %>% 
  arrange(area, home, pm_mean_winter) %>% 
  distinct(area, home_winter_id, .keep_all = TRUE) %>% 
  select(area, home, pm_mean_sampling_period)
elders_kw_report_clean <- 
  read_rds(paste0(file_path, "Output/elders_kw_report_clean.rds"))

kw_perc_exp_summary <- kids_kw_report_clean %>% 
  left_join(pm_data, by = c("area", "home")) %>% 
  filter(!is.na(kw_perc_expected)) %>% 
  #filter(area == "WMT") %>% 
  group_by(filter_type, treatment_assigned) %>% 
  summarize("Mean KW % Exp" = mean(kw_perc_expected),
            "SD KW % Exp" = sd(kw_perc_expected),
            "N KW % Exp" = n(),
            "Min KW % Exp" = min(kw_perc_expected),
            "Med KW % Exp" = median(kw_perc_expected),
            "Max KW % Exp" = max(kw_perc_expected),
            "Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE),
            "Min PM" = min(pm_mean_sampling_period, na.rm = TRUE),
            "Med PM" = median(pm_mean_sampling_period, na.rm = TRUE),
            "Max PM" = max(pm_mean_sampling_period, na.rm = TRUE))
kw_perc_exp_summary

png("Kids_KW_perc_exp.png", width=1200, height=480, bg = "white")
grid.table(kw_perc_exp_summary)
dev.off()
```