### 2021-03-26 micropem_data_cleaning.Rmd
Micropem data cleaning updates

---
title: "EldersAIR initial Micropem data work"
author: "Ethan Walker"
date: "Started 2 March 2020, Updated 26 March 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
library(readr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
```

# Load, format, and save micropem log files
```{r}
## Box location: Update Box location once final data loaded
np_micropem_initial <- read_xlsx(paste0(file_path, "Input/NP/micropem.xlsx"),
                                skip = 1, na = c("", "NULL"),
                                col_names = c("id", "home_winter_id", "home_id_num",
                                              "incomplete", "ignoreblank", "whatmissing",
                                              "ignoremissing", "ignorereason", "mpid",
                                              "last_cal", "cleaned", "zeroed", 
                                              "filter_installed", "initial_flow",
                                              "adjust", "final_flow", "location",
                                              "start_date", "start_time",
                                              "stop_date", "stop_time",
                                              "nephelometer", "laser_cycle_on",
                                              "laser_cycle_off", "sample_obs", "comments",
                                              "sampling_visit", "adult_id_char", 
                                              "battery", "attached", "fb_file_name",
                                              "mp_file_name", "qaqc", "qaqc_comments",
                                              "fitbit_qaqc", "fitbit_qaqc_comments", "fbid"),
                                col_types = c("text", "text", "text",
                                              "text", "text", "text",
                                              "text", "text", "text",
                                              "text", "text", "text", 
                                              "text", "numeric",
                                              "text", "numeric", "text",
                                              "date", "date",
                                              "date", "date",
                                              "text", "text",
                                              "text", "numeric", "text",
                                              "text", "text", 
                                              "text", "text", "text",
                                              "text", "text", "text",
                                              "text", "text", "text")) %>%
  # rename variables
  separate(start_time, c("trash1", "start_time"), sep = " ", remove = FALSE) %>% 
  separate(stop_time, c("trash1", "stop_time"), sep = " ", remove = FALSE)  %>% 
  mutate(start_date = ymd(start_date),
         stop_date = ymd(stop_date),
         area = "NPT") %>% 
  separate(mp_file_name, c("trash1", "mp_file_name"), sep = "uploads/", remove = FALSE) %>% 
  separate(mp_file_name, c("mp_file_name", "trash1"), sep = ".csv", remove = FALSE) %>%
  separate(mp_file_name, c("mp_file_name", "trash1"), sep = ".xlsx", remove = FALSE) %>%
  separate(mp_file_name, c("mp_file_name", "trash1"), sep = ".xls", remove = FALSE) %>%
  separate(fb_file_name, c("trash1", "fb_file_name"), sep = "uploads/", remove = FALSE) %>% 
  separate(fb_file_name, c("fb_file_name", "trash1"), sep = ".csv", remove = FALSE) %>%
  separate(fb_file_name, c("fb_file_name", "trash1"), sep = ".xlsx", remove = FALSE) %>%
  separate(fb_file_name, c("fb_file_name", "trash1"), sep = ".xls", remove = FALSE) %>%
  select(home_winter_id, home_id_num, area, mpid, sampling_visit, 
         start_date, start_time, stop_date, stop_time, sample_obs, 
         mp_file_name, fb_file_name) %>% 
  arrange(home_winter_id)
write_rds(np_micropem_initial, paste0(file_path, "Output/np_micropem_log.rds"))

## Box location: Update Box location once final data loaded
nn_micropem_initial <- read_xlsx(paste0(file_path, "Input/NN/micropem.xlsx"),
                                skip = 1, na = c("", "NULL"),
                                col_names = c("id", "home_winter_id", "home_id_num",
                                              "incomplete", "ignoreblank", "whatmissing",
                                              "ignoremissing", "ignorereason", "mpid",
                                              "last_cal", "cleaned", "zeroed", 
                                              "filter_installed", "initial_flow",
                                              "adjust", "final_flow", "location",
                                              "start_date", "start_time",
                                              "stop_date", "stop_time",
                                              "nephelometer", "laser_cycle_on",
                                              "laser_cycle_off", "sample_obs", "comments",
                                              "sampling_visit", "adult_id_char", 
                                              "battery", "attached", "fb_file_name",
                                              "mp_file_name", "qaqc", "qaqc_comments",
                                              "fitbit_qaqc", "fitbit_qaqc_comments", "fbid"),
                                col_types = c("text", "text", "text",
                                              "text", "text", "text",
                                              "text", "text", "text",
                                              "text", "text", "text", 
                                              "text", "numeric",
                                              "text", "numeric", "text",
                                              "date", "date",
                                              "date", "date",
                                              "text", "text",
                                              "text", "numeric", "text",
                                              "text", "text", 
                                              "text", "text", "text",
                                              "text", "text", "text",
                                              "text", "text", "text")) %>%
  # rename variables
  separate(start_time, c("trash1", "start_time"), sep = " ", remove = FALSE) %>% 
  separate(stop_time, c("trash1", "stop_time"), sep = " ", remove = FALSE)  %>% 
  mutate(start_date = ymd(start_date),
         stop_date = ymd(stop_date),
         area = "NN") %>% 
  separate(mp_file_name, c("trash1", "mp_file_name"), sep = "uploads/", remove = FALSE) %>% 
  separate(mp_file_name, c("mp_file_name", "trash1"), sep = ".csv", remove = FALSE) %>%
  separate(mp_file_name, c("mp_file_name", "trash1"), sep = ".xlsx", remove = FALSE) %>%
  separate(mp_file_name, c("mp_file_name", "trash1"), sep = ".xls", remove = FALSE) %>%
  separate(fb_file_name, c("trash1", "fb_file_name"), sep = "uploads/", remove = FALSE) %>% 
  separate(fb_file_name, c("fb_file_name", "trash1"), sep = ".csv", remove = FALSE) %>%
  separate(fb_file_name, c("fb_file_name", "trash1"), sep = ".xlsx", remove = FALSE) %>%
  separate(fb_file_name, c("fb_file_name", "trash1"), sep = ".xls", remove = FALSE) %>%
  select(home_winter_id, home_id_num, area, mpid, sampling_visit, 
         start_date, start_time, stop_date, stop_time, sample_obs, 
         mp_file_name, fb_file_name) %>% 
  arrange(home_winter_id)
write_rds(nn_micropem_initial, paste0(file_path, "Output/nn_micropem_log.rds"))

elders_micropem_log <- rbind(np_micropem_initial, nn_micropem_initial)
write_rds(elders_micropem_log, paste0(file_path, "Output/elders_micropem_log.rds"))

elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
  filter(winter_id == 1 | winter_id == 2) %>% 
  select(-adult_id_num, -adult_id_char) %>% 
  distinct(area, home_winter_id, .keep_all = TRUE)

elders_micropem_log_new <- elders_micropem_log %>% 
  left_join(elders_ids_linked, by = c("area", "home_id_num", "home_winter_id")) %>% 
  filter(!is.na(mp_file_name)) %>% 
  unite(start_datetime_pm, c("start_date", "start_time"), sep = " ", remove = FALSE) %>% 
  unite(stop_datetime_pm, c("stop_date", "stop_time"), sep = " ", remove = FALSE) %>% 
  arrange(home, winter_id)
write_rds(elders_micropem_log_new, paste0(file_path, "Output/elders_micropem_log.rds"))
```

# Initial load, format, and save EldersAIR DustTrak (PM) data
## Files were downloaded from Box: NoonanGroupData > EldersAIR
```{r}
##### Read in Micropem files #####

### NP ###
# List files
list_files <- list.files(paste0(file_path, "Input/NP/np_micropem"))

# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path, "Input/NP/np_micropem")) # set new working directory
initial_data <- tibble(list_files) %>% # use list of files from above
  mutate(save_data = lapply(list_files, read_xlsx, range = "A29:N20000",
                            col_names = c("date", "time", "pm",
                                          "temp", "rh", "battery",
                                          "in_pressure", "or_pressure",
                                          "flow", "x_axis", "y_axis",
                                          "z_axis", "vector_sum", "shutdown_reason"),
                            col_types = c("date", "date", "numeric",
                                          "numeric", "numeric", "numeric",
                                          "numeric", "numeric",
                                          "numeric", "numeric", "numeric",
                                          "numeric", "numeric", "text"))) %>%
  unnest(save_data)

# Format combined file from above; combine with other PM data below
npt_micropem <- initial_data %>% 
  separate(time, c("trash1", "time_pm"), sep = " ", remove = FALSE) %>% 
  separate(list_files, c("mp_file_name", "trash2"), sep = ".xlsx", remove = FALSE) %>%
  mutate(date_pm = ymd(date),
         pm_personal = as.numeric(pm),
         area = "NPT",
         mp_file_name = as.character(mp_file_name),
         file_type = "xlsx") %>% 
  unite("datetime", c("date_pm", "time_pm"), sep = " ", remove = FALSE) %>% 
  mutate(datetime_pm = ymd_hms(datetime)) %>% 
  select(mp_file_name, area, datetime_pm, date_pm, time_pm, pm_personal, file_type,
         temp:shutdown_reason) %>% 
  filter(!is.na(pm_personal)) %>% 
  filter(!is.na(rh)) %>% 
  ungroup() %>% 
  arrange(area, mp_file_name, datetime_pm)

# save files
write_rds(npt_micropem, paste0(file_path, "Output/npt_micropem.rds"))

### NN ###
# Only doing NPT micropem data for now, as NN data is a mess and unreliable
# List files
list_files <- list.files(paste0(file_path, "Input/NN/nn_micropem_small_20200302"))

# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path, "Input/NN/nn_micropem_small_20200302")) # set new working directory
initial_data <- tibble(list_files) %>% # use list of files from above
  mutate(save_data = lapply(list_files, read_csv, skip = 28,
                            col_names = c("date", "time", "pm", "pm2",
                                          "temp", "rh", "battery",
                                          "in_pressure", "or_pressure",
                                          "flow", "x_axis", "y_axis",
                                          "z_axis", "vector_sum", "shutdown_reason"),
                            col_types = list(col_character(), col_character(), col_double(), col_double(),
                                          col_double(), col_double(), col_double(),
                                          col_double(), col_double(),
                                          col_double(), col_double(), col_double(),
                                          col_double(), col_double(), col_character()))) %>%
  unnest(save_data)

# Format combined file from above; combine with other PM data below
nn_micropem_small <- initial_data %>% 
  separate(list_files, c("mp_file_name", "trash2"), sep = ".csv", remove = FALSE) %>%
  mutate(date_pm = mdy(date),
         pm_personal = as.numeric(pm),
         area = "NN",
         mp_file_name = as.character(mp_file_name),
         file_type = "csv") %>% 
  unite("datetime", c("date_pm", "time"), sep = " ", remove = FALSE) %>% 
  mutate(datetime_pm = ymd_hms(datetime)) %>% 
  select(mp_file_name, area, datetime_pm, date_pm, time, pm_personal, file_type,
         temp:shutdown_reason) %>% 
  filter(!is.na(pm_personal)) %>% 
  filter(!is.na(rh)) %>% 
  ungroup() %>% 
  arrange(area, mp_file_name, datetime_pm)

# save files
write_rds(nn_micropem_small, paste0(file_path, "Output/nn_micropem_small.rds"))

# Try running program on large files
list_files <- list.files(paste0(file_path, "Input/NN/nn_micropem_large_20200302"))

# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path, "Input/NN/nn_micropem_large_20200302")) # set new working directory
initial_data <- tibble(list_files) %>% # use list of files from above
  mutate(save_data = lapply(list_files, read_csv, skip = 28,
                            col_names = c("date", "time", "pm", "pm2",
                                          "temp", "rh", "battery",
                                          "in_pressure", "or_pressure",
                                          "flow", "x_axis", "y_axis",
                                          "z_axis", "vector_sum", "shutdown_reason"),
                            col_types = list(col_character(), col_character(), col_double(), col_double(),
                                          col_double(), col_double(), col_double(),
                                          col_double(), col_double(),
                                          col_double(), col_double(), col_double(),
                                          col_double(), col_double(), col_character()))) %>%
  unnest(save_data)

# Format combined file from above; combine with other PM data below
c <- initial_data %>% 
  separate(list_files, c("mp_file_name", "trash2"), sep = ".csv", remove = FALSE) %>%
  mutate(date_pm = mdy(date),
         pm_personal = as.numeric(pm),
         area = "NN",
         mp_file_name = as.character(mp_file_name),
         file_type = "csv") %>% 
  unite("datetime", c("date_pm", "time"), sep = " ", remove = FALSE) %>% 
  mutate(datetime_pm = ymd_hms(datetime)) %>% 
  select(mp_file_name, area, datetime_pm, date_pm, time, pm_personal, file_type,
         temp:shutdown_reason) %>% 
  filter(!is.na(pm_personal)) %>% 
  filter(!is.na(rh)) %>% 
  ungroup() %>% 
  arrange(area, mp_file_name, datetime_pm)

# save files
write_rds(nn_micropem_large, paste0(file_path, "Output/nn_micropem_large.rds"))

nn_micropem_small <- read_rds(paste0(file_path, "Output/nn_micropem_small.rds"))
nn_micropem_large <- read_rds(paste0(file_path, "Output/nn_micropem_large.rds"))

nn_micropem <- rbind(nn_micropem_large, nn_micropem_small)

write_rds(nn_micropem, paste0(file_path, "Output/nn_micropem_initial.rds"))
```

# Clean up and filter all micropem data
```{r}
# load NPT data
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

npt_micropem_log <- read_rds(paste0(file_path, "Output/elders_micropem_log.rds")) %>% 
  filter(area == "NPT") %>% 
  filter(!is.na(stop_date)) %>% 
  select(area, home, home_winter_id:winter_id, 
         -sample_obs, treatment)

npt_micropem <- read_rds(paste0(file_path, "Output/npt_micropem.rds"))

npt_micropem_new <- npt_micropem %>% 
  left_join(npt_micropem_log, by = c("area", "mp_file_name")) %>% 
  mutate(datetime_pm = ymd_hms(datetime_pm),
         start_datetime_pm = ymd_hms(start_datetime_pm),
         stop_datetime_pm = ymd_hms(stop_datetime_pm))

# Filter to only dates that fall within sampling timeframe
npt_micropem_new2 <- npt_micropem_new %>% 
  mutate(date_filter = if_else(datetime_pm >= start_datetime_pm & 
                               datetime_pm <= stop_datetime_pm, 1, 0))

npt_micropem_new3 <- npt_micropem_new2 %>% 
  filter(date_filter == 1)

# Save data
write_rds(npt_micropem_new3, paste0(file_path, "Output/npt_micropem_new.rds"))

check <- npt_micropem_log %>% 
  distinct(mp_file_name, .keep_all = TRUE)

# load NN data
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

nn_micropem_log <- read_rds(paste0(file_path, "Output/elders_micropem_log.rds")) %>% 
  filter(area == "NN") %>% 
  filter(!is.na(stop_date)) %>% 
  select(area, home, home_winter_id:winter_id, 
         -sample_obs, treatment)

nn_micropem <- read_rds(paste0(file_path, "Output/nn_micropem_initial.rds"))

nn_micropem_new <- nn_micropem %>% 
  left_join(nn_micropem_log, by = c("area", "mp_file_name")) %>% 
  mutate(datetime_pm = ymd_hms(datetime_pm),
         start_datetime_pm = ymd_hms(start_datetime_pm),
         stop_datetime_pm = ymd_hms(stop_datetime_pm))

# Filter to only dates that fall within sampling timeframe
nn_micropem_new2 <- nn_micropem_new %>% 
  mutate(date_filter = if_else(datetime_pm >= start_datetime_pm & 
                               datetime_pm <= stop_datetime_pm, 1, 0))

nn_micropem_new3 <- nn_micropem_new2 %>% 
  filter(date_filter == 1) %>% 
  mutate(mp_file_name = as.factor(mp_file_name),
         area = as.factor(area),
         home = as.factor(home),
         home_winter_id = as.factor(home_winter_id),
         home_id_num = as.factor(home_id_num),
         mpid = as.factor(mpid),
         sampling_visit = as.factor(sampling_visit))

# Save data
write_rds(nn_micropem_new3, paste0(file_path, "Output/nn_micropem_new.rds"))

summary(nn_micropem_new3)

check <- nn_micropem_log %>% 
  distinct(mp_file_name, .keep_all = TRUE)
```

# Summarize filtered micropem data for further cleaning
```{r}
#NPT data
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

npt_micropem_new <- read_rds(paste0(file_path, "Output/npt_micropem_new.rds"))

micropem_observations_summary <- npt_micropem_new %>% 
  group_by(area, home, winter_id, sampling_visit) %>% 
  count(mp_file_name)
head(micropem_observations_summary, 10)

micropem_pm_summary <- npt_micropem_new %>% 
  #group_by(area, winter_id, sampling_visit) %>%
  group_by(area, mp_file_name) %>% 
  summarize("n" = n(),
            "Mean PM" = mean(pm_personal, na.rm = TRUE), 
            "SD PM" = sd(pm_personal, na.rm = TRUE),
            "Min PM" = min(pm_personal, na.rm = TRUE), 
            "25%" = quantile(pm_personal, 0.25, na.rm = TRUE),
            "Median PM" = median(pm_personal, na.rm = TRUE),
            "75%" = quantile(pm_personal, 0.75, na.rm = TRUE),
            "Max PM" = max(pm_personal, na.rm = TRUE))
micropem_pm_summary

# NN data
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

nn_micropem_new <- read_rds(paste0(file_path, "Output/nn_micropem_new.rds"))

micropem_observations_summary <- nn_micropem_new %>% 
  group_by(area, home, winter_id, sampling_visit) %>% 
  count(mp_file_name)
head(micropem_observations_summary, 10)

micropem_pm_summary <- nn_micropem_new %>% 
  #group_by(area, winter_id, sampling_visit) %>%
  group_by(area, mp_file_name) %>% 
  summarize("n" = n(),
            "Mean PM" = mean(pm_personal, na.rm = TRUE), 
            "SD PM" = sd(pm_personal, na.rm = TRUE),
            "Min PM" = min(pm_personal, na.rm = TRUE), 
            "25%" = quantile(pm_personal, 0.25, na.rm = TRUE),
            "Median PM" = median(pm_personal, na.rm = TRUE),
            "75%" = quantile(pm_personal, 0.75, na.rm = TRUE),
            "Max PM" = max(pm_personal, na.rm = TRUE))
micropem_pm_summary
```

# Continue micropem cleaning
```{r}
# NPT data
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

npt_micropem <- read_rds(paste0(file_path, "Output/npt_micropem_new.rds"))

npt_micropem_new <- npt_micropem %>%   
  group_by(area, mp_file_name) %>% 
  # new variables summarizing pm2.5 within each home
  mutate(pm_total_observations = n(),
         # new datetime values to calculate sampling interval
         first_datetime = first(datetime_pm),
         last_datetime = last(datetime_pm),
         first_datetime = ymd_hms(first_datetime),
         last_datetime = ymd_hms(last_datetime),
         diff_datetime = interval(first_datetime, last_datetime),
         # total sampling interval, based on first and last datetimes
         pm_sample_interval = diff_datetime/86400,
         # expected number of observations with the sampling interval
         expected_obs = pm_sample_interval*24*60,
         # difference between actual and expected observations
         actual_vs_expected_diff = pm_total_observations - expected_obs,
         # indicator variable if actual vs expected obs > 10 or greater
         actual_vs_expected = if_else(actual_vs_expected_diff < 10 &
                                      actual_vs_expected_diff > -10, 1, 0),
         pm_mean = mean(pm_personal, na.rm = TRUE),
         pm_sd = sd(pm_personal, na.rm = TRUE),
         pm_min = min(pm_personal, na.rm = TRUE),
         pm_median = median(pm_personal, na.rm = TRUE),
         pm_max = max(pm_personal, na.rm = TRUE),
         # breaking sampling period into days
         sampling_day = if_else(datetime_pm - first_datetime < 86400, 1, 0),
         sampling_day = if_else(datetime_pm - first_datetime >= 86400 & datetime_pm - first_datetime < 172800, 2, sampling_day),
         sampling_day = if_else(datetime_pm - first_datetime > 172800, 3, sampling_day)) %>%
  filter(sampling_day != 0) %>% 
  # calculate pm means for each sampling day
  group_by(area, mp_file_name, sampling_day) %>% 
  mutate(pm_mean_daily = mean(pm_personal, na.rm = TRUE),
         pm_sampling_day = sampling_day,
         day_of_week = weekdays(date_pm)) %>%
  group_by(area, mp_file_name) %>% 
  mutate(start_datetime_pm = ymd_hms(start_datetime_pm),
         pm_start_diff = interval(first_datetime, start_datetime_pm),
         pm_start_diff_hrs = pm_start_diff/3600) %>% 
  ungroup() %>% 
  select(-sampling_day) %>% 
  arrange(area, home_winter_id, sampling_visit)

write_rds(npt_micropem_new, paste0(file_path, "Output/npt_micropem_new2.rds"))

pm_summary <- npt_micropem_new %>% 
  distinct(mp_file_name, .keep_all = TRUE) %>% 
  select(-datetime_pm, -date_pm, -time_pm, -pm_personal, -file_type,
         -x_axis, -y_axis, -z_axis, -vector_sum,
         -start_date, -start_time, -stop_date, -stop_time, -day_of_week,
         -fb_file_name, -treatment, -date_filter, -pm_mean_daily, -pm_sampling_day) %>% 
  select(mp_file_name:sampling_visit, winter_id, start_datetime_pm,
         first_datetime, stop_datetime_pm, last_datetime, diff_datetime,
         pm_sample_interval, pm_total_observations, expected_obs, actual_vs_expected_diff,
         actual_vs_expected:pm_start_diff_hrs)
write_csv(pm_summary, paste0(file_path, "Output/npt_pm_summary.csv"))

# NN data
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

nn_micropem <- read_rds(paste0(file_path, "Output/nn_micropem_new.rds"))

nn_micropem_new <- nn_micropem %>%   
  group_by(area, mp_file_name) %>% 
  # new variables summarizing pm2.5 within each home
  mutate(pm_total_observations = n(),
         # new datetime values to calculate sampling interval
         first_datetime = first(datetime_pm),
         last_datetime = last(datetime_pm),
         first_datetime = ymd_hms(first_datetime),
         last_datetime = ymd_hms(last_datetime),
         diff_datetime = interval(first_datetime, last_datetime),
         # total sampling interval, based on first and last datetimes
         pm_sample_interval = diff_datetime/86400,
         # expected number of observations with the sampling interval
         expected_obs = as.numeric(pm_sample_interval*24*60),
         # difference between actual and expected observations
         actual_vs_expected_diff = pm_total_observations - expected_obs,
         # indicator variable if actual vs expected obs > 10 or greater
         actual_vs_expected = if_else(actual_vs_expected_diff < 10 &
                                      actual_vs_expected_diff > -10, 1, 0),
         pm_mean = mean(pm_personal, na.rm = TRUE),
         pm_sd = sd(pm_personal, na.rm = TRUE),
         pm_min = min(pm_personal, na.rm = TRUE),
         pm_median = median(pm_personal, na.rm = TRUE),
         pm_max = max(pm_personal, na.rm = TRUE),
         # breaking sampling period into days
         sampling_day = if_else(datetime_pm - first_datetime < 86400, 1, 0),
         sampling_day = if_else(datetime_pm - first_datetime >= 86400 & datetime_pm - first_datetime < 172800, 2, sampling_day),
         sampling_day = if_else(datetime_pm - first_datetime > 172800, 3, sampling_day)) %>%
  filter(sampling_day != 0) %>% 
  # calculate pm means for each sampling day
  group_by(area, mp_file_name, sampling_day) %>% 
  mutate(pm_mean_daily = mean(pm_personal, na.rm = TRUE),
         pm_sampling_day = sampling_day,
         day_of_week = weekdays(date_pm)) %>%
  group_by(area, mp_file_name) %>% 
  mutate(start_datetime_pm = ymd_hms(start_datetime_pm),
         pm_start_diff = interval(first_datetime, start_datetime_pm),
         pm_start_diff_hrs = pm_start_diff/3600) %>% 
  ungroup() %>% 
  select(-sampling_day) %>% 
  arrange(area, home_winter_id, sampling_visit)

write_rds(nn_micropem_new, paste0(file_path, "Output/nn_micropem_new2.rds"))

nn_pm_summary <- nn_micropem %>% 
  distinct(mp_file_name, .keep_all = TRUE) %>% 
  select(-datetime_pm, -date_pm, -time, -pm_personal, -file_type,
         -x_axis, -y_axis, -z_axis, -vector_sum,
         -start_date, -start_time, -stop_date, -stop_time, -day_of_week,
         -fb_file_name, -treatment, -date_filter, -pm_mean_daily, -pm_sampling_day) %>% 
  select(mp_file_name:sampling_visit, winter_id, start_datetime_pm,
         first_datetime, stop_datetime_pm, last_datetime, diff_datetime,
         pm_sample_interval, pm_total_observations, expected_obs, actual_vs_expected_diff,
         actual_vs_expected:pm_start_diff_hrs)
write_csv(pm_summary, paste0(file_path, "Output/nn_pm_summary.csv"))
```

# Function to look at PM data across time
```{r}
# Plot PM by datetime to look at trends, gaps, and peaks
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")

pm_time_plot_function <- function(id, location, visit) {

pm_time_plot <- nn_micropem %>% 
  filter(home_winter_id == id & area == location & sampling_visit == visit) %>% 
  #filter(pm < 30000) %>% 
  arrange(datetime_pm) %>% 
  mutate(pm_sampling_day = as.character(pm_sampling_day),
         day_of_week = factor(day_of_week,
                              levels = c("Sunday", "Monday", "Tuesday", "Wednesday", 
                                         "Thursday", "Friday", "Saturday"))) %>%
  ggplot() + 
    geom_point(aes(datetime_pm, pm_personal, color = day_of_week), size = 1.5) +
    theme_classic() +
    labs(title = id,
         y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")")),
         x = "Datetime",
         color = "Day of week") +
    theme(axis.title.y = element_text(size = 12,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_text(size = 12),
          axis.text.x = element_blank(),
          axis.text.y = element_text(size = 12, color = "black"),
          axis.line.x = element_line(colour = "black", size = 1), 
          axis.line.y = element_line(colour = "black", size = 1), 
          axis.ticks = element_blank()) +
    scale_color_manual(values = jv_palette)

gaps <<- nn_micropem %>% 
  filter(home_winter_id == id & area == location & sampling_visit == visit) %>% 
  mutate(datetime_pm = ymd_hms(datetime_pm),
         datetime_lead = ymd_hms(lead(datetime_pm)),
         datetime_1 = seconds(datetime_pm),
         datetime_2 = seconds(datetime_lead),
         datetime_diff_minutes = as.numeric(datetime_2 - datetime_1)/60,
         datetime_diff_hours = as.numeric(datetime_diff_minutes/60)) %>% 
  arrange(desc(datetime_diff_minutes)) %>% 
  select(day_of_week, pm_sampling_day, datetime_diff_minutes, datetime_diff_hours, 
         datetime_pm, datetime_lead)

pm_desc <<- nn_micropem %>% 
  filter(home_winter_id == id & area == location & sampling_visit == visit) %>% 
  arrange(desc(pm_personal)) %>% 
  select(datetime_pm, pm_personal, pm_sampling_day)

pm_negs <<- nn_micropem %>% 
  filter(home_winter_id == id & area == location & sampling_visit == visit) %>% 
  filter(pm_personal < 0) %>% 
  arrange(desc(pm_personal)) %>% 
  select(datetime_pm, pm_personal, pm_sampling_day)

pm_time_plot

}
```

# Run function from above to look at results
## Function run for each household to check for odd trends/gaps in the data
## Notes from this process on uploaded to Box:
### EldersAIR > Ethan - data cleaning and analysis > pm_cleaning_summary
```{r}
# load data
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

npt_micropem <- read_rds(paste0(file_path, "Output/npt_micropem_new2.rds")) 
nn_micropem <- read_rds(paste0(file_path, "Output/nn_micropem_new2.rds"))

# Change function input to look at specific home
# Run all of these lines together
pm_time_plot_function(id = "81", location = "NN", visit = 1)
head(gaps, 10)
head(pm_desc, 10)
tail(pm_desc, 10)

# filter data for a single home
check_data <- nn_micropem %>% 
  filter(area == "NN" & home_winter_id == "81" & sampling_visit == 1) %>% 
  filter(pm_personal < 0)
```

# Further cleaning steps after checking individual files
```{r}
# load data
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

npt_micropem <- read_rds(paste0(file_path, "Output/npt_micropem_new2.rds"))

# https://cran.r-project.org/web/packages/rtimicropem/vignettes/chai_data_cleaning.html 
# The method in the above link holds off negative values and checks change in mean
# They check to see if mean changed by >5%, and don't use the file if it did

npt_summary <- npt_micropem %>% 
  distinct(mp_file_name, .keep_all = TRUE) %>% 
  summarize(mean_pm = mean(pm_mean, na.rm = TRUE), 
            sd_pm = sd(pm_mean, na.rm = TRUE),
            min_pm = min(pm_mean, na.rm = TRUE),
            med_pm = median(pm_mean, na.rm = TRUE),
            max_pm = max(pm_mean, na.rm = TRUE))

npt_summary_new <- npt_micropem %>% 
  group_by(mp_file_name) %>% 
  mutate(pm_personal_new = if_else(pm_personal < 0, 0, pm_personal),
         pm_mean_new = mean(pm_personal_new, na.rm = TRUE),
         pm_diff = round(pm_mean_new - pm_mean, digits = 2),
         pm_diff_perc = round(abs(pm_diff/pm_mean*100), digits = 2),
         pm_indicator = if_else(pm_diff_perc > 5 & pm_diff > 3, 1, 0)) %>% 
  ungroup() %>% 
  distinct(mp_file_name, .keep_all = TRUE) %>% 
  select(mp_file_name, area, home, home_winter_id, sampling_visit,
         pm_mean, pm_mean_new, pm_diff, pm_diff_perc, pm_indicator,
         pm_sample_interval) %>% 
  filter(pm_indicator == 0) %>% 
  filter(pm_sample_interval > 1.6)
  summarize(mean_pm = mean(pm_mean_new, na.rm = TRUE), 
            sd_pm = sd(pm_mean_new, na.rm = TRUE),
            min_pm = min(pm_mean_new, na.rm = TRUE),
            med_pm = median(pm_mean_new, na.rm = TRUE),
            max_pm = max(pm_mean_new, na.rm = TRUE))

npt_summary
npt_summary_new

# Above, I changed all negative PM values to 0 and checked the change in mean PM per file
# Some files had small absolute changes in mean PM, but since levels were so low,
# the percent changes were very high. I filtered out files with the following criteria:
# -After changing negative values to 0, if percent change in mean PM was >5%,
# -and if the absolute difference in mean PM was >3 ug/m3
# I also calculated overall summary stats before and after filtering out these files.
# Mean and median PM changed by less than 1 ug/m3. SD was essentially unchanged.
# By using this method, I filter out files that are highly impacted by negatives.
# The remaining files are not as impacted by negatives, and overall summary stats are
# impacted minimally.
## This method filtered out 7 files.
# I also filter out files that had a run time <80% expected.
## This method filtered out 9 additional files, for a total of 16 files removed.
# The steps below will use the methods on the full dataset

npt_micropem_new <- npt_micropem %>% 
  group_by(mp_file_name) %>% 
  mutate(pm_personal_new = if_else(pm_personal < 0, 0, pm_personal),
         pm_mean_new = mean(pm_personal_new, na.rm = TRUE),
         pm_diff = round(pm_mean_new - pm_mean, digits = 2),
         pm_diff_perc = round(abs(pm_diff/pm_mean*100), digits = 2),
         pm_indicator = if_else(pm_diff_perc > 5 & pm_diff > 3, 1, 0)) %>% 
  ungroup() %>% 
  filter(pm_indicator == 0) %>% 
  filter(pm_sample_interval > 1.6)

write_rds(npt_micropem_new, paste0(file_path, "Output/npt_micropem_new3.rds"))
```