### 2021-03-31 main_analysis_initial.Rmd
tables and initial model for primary analysis

---
title: 'EldersAIR: Main analysis - initial'
author: "Ethan Walker"
date: "Started 31 March 2021; Updated 31 March 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, include = TRUE,
                      message = FALSE, warning = FALSE)
```

```{r}
library(readxl)
library(naniar)
library(lubridate)
library(emmeans)
library(broom)
library(broom.mixed)
library(zoo)
library(lme4)
library(lmerTest)
library(mctest)
library(tidyverse)
library(knitr)
library(kableExtra)
library(MASS)
library(faraway)
library(DHARMa)
```

```{r}
# Load data

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

treatments_blind_nn <- read_xlsx(paste0(file_path, "Input/NN/treatments_blind.xlsx")) %>% 
  mutate(area = "NN")

treatments_blind_npt <- read_xlsx(paste0(file_path, "Input/NP/treatments_blind.xlsx")) %>% 
  mutate(area = "NPT")

treatments_blind <- rbind(treatments_blind_nn, treatments_blind_npt) %>% 
  mutate(home_id_num = as.factor(HomeID),
         treatment_blind = as.factor(CodedCondition),
         area = as.factor(area)) %>% 
  dplyr::select(area, home_id_num, treatment_blind)

elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
  left_join(treatments_blind, by = c("area", "home_id_num")) 
```

```{r}
# Prep data to run through the model
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp
## calculate mean outcome per winter
## spread data to 1 row per adult - so Winter 1/baseline outcome is a covariate
analysis_data <- elders_cleaned_dataset_by_visit %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, adult_id_num, winter_id) %>% 
  mutate(mean_outcome = mean(fvc_pp, na.rm = T)) %>% # select outcome here
  ungroup() %>% 
  distinct(area, adult_id_num, winter_id, .keep_all = T) %>% 
  group_by(area, adult_id_num) %>% 
  pivot_wider(names_from = "winter_id", values_from = "mean_outcome") %>% 
  mutate(mean_outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(!is.na(mean_outcome), lag(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, mean_outcome = NaN)) %>% 
  filter(!is.na(mean_outcome)) %>% 
  ungroup()

table(analysis_data$area, analysis_data$treatment_blind)
```

```{r}
model_results <- lmer(mean_outcome ~ treatment_blind + baseline_outcome + 
                      (1 | area), data = analysis_data)

summary(model_results)

tidy_results <- tidy(model_results, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value)
tidy_results
```