### 2021-04-09 main_analysis_initial.Rmd
Primary model updates

---
title: 'EldersAIR: Main analysis - initial'
author: "Ethan Walker"
date: "Started 31 March 2021; Updated 2 April 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, include = TRUE,
                      message = FALSE, warning = FALSE)
```

```{r}
library(readxl)
library(naniar)
library(lubridate)
library(emmeans)
library(broom)
library(broom.mixed)
library(zoo)
library(car)
library(lme4)
library(lmerTest)
library(mctest)
library(tidyverse)
library(knitr)
library(kableExtra)
library(MASS)
library(faraway)
#library(DHARMa)
```

```{r}
# Load data

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

treatments_blind_nn <- read_xlsx(paste0(file_path, "Input/NN/treatments_blind.xlsx")) %>% 
  mutate(area = "NN")

treatments_blind_npt <- read_xlsx(paste0(file_path, "Input/NP/treatments_blind.xlsx")) %>% 
  mutate(area = "NPT")

treatments_blind <- rbind(treatments_blind_nn, treatments_blind_npt) %>% 
  mutate(home_id_num = as.factor(HomeID),
         treatment_blind = as.factor(CodedCondition),
         area = as.factor(area)) %>% 
  dplyr::select(area, home_id_num, treatment_blind)

elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
  left_join(treatments_blind, by = c("area", "home_id_num")) 
```

```{r}
# Prep data to run through the model
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp
## calculate mean outcome per winter
## spread data to 1 row per adult - so Winter 1/baseline outcome is a covariate
analysis_data_1row <- elders_cleaned_dataset_by_visit %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, adult_id_char, winter_id) %>% 
  mutate(mean_outcome = mean(fvc, na.rm = T)) %>% # select outcome here
  ungroup() %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, mean_outcome) %>% 
  distinct(area, adult_id_char, winter_id, .keep_all = T) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "mean_outcome") %>% 
  mutate(mean_outcome = `2`,
         baseline_outcome = `1`) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, mean_outcome = NaN)) %>% 
  #filter(!is.na(mean_outcome)) %>% 
  ungroup()

analysis_data_2row <- elders_cleaned_dataset_by_visit %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  mutate(outcome = fvc) %>% # select outcome here
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, outcome) %>% 
  group_by(area, adult_id_char, sampling_visit) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(!is.na(outcome) | !is.na(baseline_outcome)) %>% 
  ungroup()

analysis_data_4row <- elders_cleaned_dataset_by_visit %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  mutate(outcome = fvc) %>%  # select outcome here
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, outcome)

#table(analysis_data$area, analysis_data$treatment_blind)
#table(analysis_data$area, analysis_data$cohort)
```

At this point (4/9/21), the 3 model scenarios below give similar estimates with
largely overlapping confidence intervals. The 1-row data model seems to have better
precision for some reason (narrower CIs). This holds true for all of the outcomes.
Next steps will be to review the literature again and look for pros and cons of
each method, potential sources of bias for the methods, how they handle missing
data, etc. Also refer back to discussion with Jon about missing data, REML, and
if I need to do anything else to account for missing observations.
Also think about whether a time function/variable needs to be included if the 
repeated measures models are used. Refer back to notes from 4/5/21 meeting with Jon.

# Model Code - 1 row data
```{r, echo=T}
# 1 row data
itt_health_results <- lmer(mean_outcome ~ treatment_blind + baseline_outcome + 
                      (1 | home:cohort:area), data = analysis_data_1row)

tidy_results <- tidy(itt_health_results, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value)
tidy_results
```

## Model with (1 | cohort:area): variance for cohort random term = 0
## Variance for only (1 | cohort) still = 0
## Cohort does not explain any variance beyond what the residual variance does
## Removing or leaving cohort in does not change the model results
## This issue also does not happen for all outcomes, so leaving (1 | cohort:area) in primary model
```{r, echo=T}
isSingular(itt_health_results)
#rePCA(itt_health_results)
```

# Model Summary
```{r}
summary(itt_health_results)
```

# Model Code - 2 row data
```{r, echo=T}
# 1 row data
itt_health_results <- lmer(outcome ~ treatment_blind + baseline_outcome + 
                      (1 | adult_id_char:home:cohort:area), data = analysis_data_2row)
tidy_results <- tidy(itt_health_results, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value)
tidy_results
```

```{r, echo=T}
isSingular(itt_health_results)
#rePCA(itt_health_results)
```

# Model Summary
```{r}
summary(itt_health_results)
```

# Model Code - 4 row data
```{r}
# 4 row data
itt_health_results <- lmer(outcome ~ treatment_blind*winter_id + 
                      (1 | adult_id_char:home:cohort:area), data = analysis_data_4row)

int_emmeans <- emmeans(itt_health_results, revpairwise ~ treatment_blind | winter_id)

save_contrasts <- confint(int_emmeans$contrasts) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((lower.CL), digits = 2),
         conf.high = round((upper.CL), digits = 2)) %>% 
  dplyr::select(contrast, winter_id, estimate, conf.low, conf.high)
save_contrasts
```

# Model Summary
```{r}
summary(itt_health_results)

emmeans(itt_health_results, pairwise ~ winter_id | treatment_blind)
```

# Tidy Model Results
```{r}
tidy_results <- tidy(itt_health_results, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value)
tidy_results
```

# Model Diagnostic Plots
```{r}
plot(itt_health_results, main = "ITT Primary Model", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results), main = "ITT Primary Model")
qqline(residuals(itt_health_results))
```