### 2021-04-16 main_analysis_comparisons_diagnostics.Rmd
New RMDs for main analysis

---
title: 'EldersAIR: Main analysis - model comparisons/diagnostics'
author: "Ethan Walker"
date: "Started 16 April 2021; Updated 16 April 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, include = TRUE,
                      message = FALSE, warning = FALSE)
```

```{r}
library(readxl)
library(naniar)
library(lubridate)
library(emmeans)
library(broom)
library(broom.mixed)
library(zoo)
library(car)
library(lme4)
library(lmerTest)
library(mctest)
library(tidyverse)
library(knitr)
library(kableExtra)
#library(MASS)
library(faraway)
library(splines)
library(rptR)
library(sjstats)
```

```{r}
# Load data

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

treatments_blind_nn <- read_xlsx(paste0(file_path, "Input/NN/treatments_blind.xlsx")) %>% 
  mutate(area = "NN")

treatments_blind_npt <- read_xlsx(paste0(file_path, "Input/NP/treatments_blind.xlsx")) %>% 
  mutate(area = "NPT")

treatments_blind <- rbind(treatments_blind_nn, treatments_blind_npt) %>% 
  mutate(home_id_num = as.factor(HomeID),
         treatment_blind = as.factor(CodedCondition),
         area = as.factor(area)) %>% 
  select(area, home_id_num, treatment_blind)

elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
  left_join(treatments_blind, by = c("area", "home_id_num")) 
```

```{r}
# Prep analysis dataset
analysis_data <- elders_cleaned_dataset_by_visit %>% 
  mutate(treatment = factor(treatment,
                            levels = c("Placebo", "Filter", "Education"),
                            labels = c("Placebo", "Filter", "Education")),
         visit_intervention_diff = round(difftime(sampling_date, intervention_date, 
                                            units = "days"), digits = 0),
         visit_intervention_diff = as.numeric(visit_intervention_diff),
         visit_intervention_diff = if_else(winter_id == 1, 9999,
                                           visit_intervention_diff))

#table(analysis_data$area, analysis_data$treatment_blind)
#table(analysis_data$area, analysis_data$cohort)
```

```{r}
# Prep outcome variables
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
bp_sys_data_1row <- analysis_data %>% 
  mutate(outcome = bp_sys) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, adult_id_char, winter_id) %>% 
  mutate(mean_outcome = mean(outcome, na.rm = T)) %>% 
  ungroup() %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, treatment, mean_outcome) %>% 
  distinct(area, adult_id_char, winter_id, .keep_all = T) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "mean_outcome") %>% 
  mutate(mean_outcome = `2`,
         baseline_outcome = `1`) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, mean_outcome = NaN)) %>% 
  ungroup()

bp_sys_data_2row <- analysis_data %>% 
  mutate(outcome = bp_sys) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, treatment, outcome,
                visit_intervention_diff) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

# bp_sys: try removing lines 127, 202, 95, 32, 48, 31, 17, 216, 128, 4
bp_sys_data_2row_outliers <- bp_sys_data_2row [-c(127, 202, 95, 32, 48, 31, 17, 216, 128, 4), ]

bp_sys_data_4row <- analysis_data %>% 
  mutate(outcome = bp_sys) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  mutate(visit_intervention_diff = if_else(visit_intervention_diff == 9999, 0,
                                           visit_intervention_diff)) %>%  
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, treatment, outcome,
                visit_intervention_diff)

```

\pagebreak

# Systolic Blood Pressure models
```{r, echo=T}
itt_health_results_1row <- lmer(mean_outcome ~ treatment_blind + baseline_outcome + 
                                (1 | home:cohort:area),
                                data = bp_sys_data_1row)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_1row, main = "Sys BP 1-Row", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_1row), main = "Sys BP 1-Row")
qqline(residuals(itt_health_results_1row))
```

\pagebreak

```{r, echo=T}
itt_health_results_2row <- lmer(outcome ~ treatment_blind + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_sys_data_2row)
```

```{r, fig.width=6, fig.height=3.5}
itt_health_results_2row_days <- lmer(outcome ~ treatment_blind + visit_intervention_diff + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_sys_data_2row)

itt_health_results_2row_visit <- lmer(outcome ~ treatment_blind + sampling_visit + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_sys_data_2row)

plot(itt_health_results_2row, main = "Sys BP 2-Row", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_2row), main = "Sys BP 2-Row")
qqline(residuals(itt_health_results_2row))
```

\pagebreak

```{r, echo=T}
itt_health_results_2row_outliers <- lmer(outcome ~ treatment_blind + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_sys_data_2row_outliers)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_2row_outliers, main = "Sys BP 2-Row, 10 outliers removed", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_2row_outliers), main = "Sys BP 2-Row, 10 outliers removed")
qqline(residuals(itt_health_results_2row_outliers))
```

\pagebreak

```{r, echo=T}
itt_health_results_2row_int <- lmer(outcome ~ treatment_blind*sampling_visit + 
                                    baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                    data = bp_sys_data_2row)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_2row_int, main = "Sys BP 2-Row, Visit interaction", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_2row_int), main = "Sys BP 2-Row, Visit interaction")
qqline(residuals(itt_health_results_2row_int))
```

\pagebreak

```{r, echo=T}
itt_health_results_4row <- lmer(outcome ~ treatment_blind*winter_id + 
                                (1 | adult_id_char:home:cohort:area), 
                                data = bp_sys_data_4row)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_4row, main = "Sys BP 4-Row, Winter interaction", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_4row), main = "Sys BP 4-Row, Winter interaction")
qqline(residuals(itt_health_results_4row))
```

\pagebreak

```{r, fig.width=8, fig.height=6}
# Combine model results
tidy_results_1row <- tidy(itt_health_results_1row, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "1-row",
         sampling_visit = "none") %>% 
  dplyr::select(-p.value)

tidy_results_2row <- tidy(itt_health_results_2row, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row",
         sampling_visit = "none") %>% 
  dplyr::select(-p.value)

tidy_results_2row_outliers <- tidy(itt_health_results_2row_outliers, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row, outliers",
         sampling_visit = "none") %>% 
  dplyr::select(-p.value)

tidy_results_2row_days <- tidy(itt_health_results_2row_days, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row days covar",
         sampling_visit = "days since intervention") %>% 
  dplyr::select(-p.value)

tidy_results_2row_visit <- tidy(itt_health_results_2row_visit, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row visit covar",
         sampling_visit = "covariate") %>% 
  dplyr::select(-p.value)

int_emmeans <- emmeans(itt_health_results_2row_int, revpairwise ~ treatment_blind | sampling_visit)

tidy_results_2row_int <- confint(int_emmeans$contrasts) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((lower.CL), digits = 2),
         conf.high = round((upper.CL), digits = 2)) %>% 
  dplyr::select(contrast, sampling_visit, estimate, conf.low, conf.high) %>% 
  filter(contrast == "Queso - Guac" | contrast == "Salsa - Guac") %>% 
  mutate(contrast = as.character(contrast),
         contrast = if_else(contrast == "Queso - Guac" & sampling_visit == 1, "Queso Visit 1", contrast),
         contrast = if_else(contrast == "Salsa - Guac" & sampling_visit == 1, "Salsa Visit 1", contrast),
         contrast = if_else(contrast == "Queso - Guac" & sampling_visit == 2, "Queso Visit 2", contrast),
         contrast = if_else(contrast == "Salsa - Guac" & sampling_visit == 2, "Salsa Visit 2", contrast),
         contrast = as.factor(contrast),
         model = "2-row visit interaction") %>% 
  rename(term = contrast)

int_emmeans <- emmeans(itt_health_results_4row, revpairwise ~ treatment_blind | winter_id)

tidy_results_4row <- confint(int_emmeans$contrasts) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((lower.CL), digits = 2),
         conf.high = round((upper.CL), digits = 2)) %>% 
  dplyr::select(contrast, winter_id, estimate, conf.low, conf.high) %>% 
  filter((contrast == "Queso - Guac" | contrast == "Salsa - Guac") & winter_id == 2) %>% 
  mutate(contrast = if_else(contrast == "Queso - Guac", "Queso", "Salsa"),
         model = "4-row",
         sampling_visit = "none") %>% 
  rename(term = contrast) %>% 
  dplyr::select(-winter_id)

tidy_results_combined <- rbind(tidy_results_1row, tidy_results_2row, tidy_results_2row_outliers,
                               tidy_results_2row_days, tidy_results_2row_visit,
                               tidy_results_2row_int, tidy_results_4row)

# Calculate ICC
## The ratio of the between-cluster variance to the total variance
## It can also be interpreted as the correlation among observations within the same cluster
icc_model <- lmer(outcome ~ (1 | adult_id_char), bp_sys_data_4row)
bp_sys_icc <- icc(icc_model)

# Plot combined results
itt_plot_estimates <- tidy_results_combined %>%
  mutate(model = as.factor(model)) %>% 
  ggplot(aes(group = term, shape = term)) +
  geom_point(aes(x=model, y=estimate), 
             position = position_dodge(width = 0.5), size = 4) +
  geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
  geom_hline(yintercept = 0) +
  theme_bw() +  
  ggtitle(label = "Sys BP Model comparison for ITT framework",
          subtitle = "ICC for subject = 0.371") +
  labs(y = "Estimate compared to Guac") +
  labs(x = "", group = "", shape = "") +
  theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
itt_plot_estimates
```

\pagebreak

# Diastolic Blood Pressure models
```{r}
# Prep outcome variables
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
bp_dia_data_1row <- analysis_data %>% 
  mutate(outcome = bp_dia) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, adult_id_char, winter_id) %>% 
  mutate(mean_outcome = mean(outcome, na.rm = T)) %>% 
  ungroup() %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, treatment, mean_outcome) %>% 
  distinct(area, adult_id_char, winter_id, .keep_all = T) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "mean_outcome") %>% 
  mutate(mean_outcome = `2`,
         baseline_outcome = `1`) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, mean_outcome = NaN)) %>% 
  ungroup()

bp_dia_data_2row <- analysis_data %>% 
  mutate(outcome = bp_dia) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, treatment, outcome,
                visit_intervention_diff) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

# bp_dia: try removing lines 42, 95, 192, 127, 48, 202, 191, 31, 223, 7
bp_dia_data_2row_outliers <- bp_dia_data_2row [-c(42, 95, 192, 127, 48, 202, 191, 31, 223, 7), ]

bp_dia_data_4row <- analysis_data %>% 
  mutate(outcome = bp_dia) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  mutate(visit_intervention_diff = if_else(visit_intervention_diff == 9999, 0,
                                           visit_intervention_diff)) %>%  
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, treatment, outcome,
                visit_intervention_diff)

```

```{r, echo=T}
itt_health_results_1row <- lmer(mean_outcome ~ treatment_blind + baseline_outcome + 
                                (1 | home:cohort:area),
                                data = bp_dia_data_1row)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_1row, main = "Dia BP 1-Row", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_1row), main = "Dia BP 1-Row")
qqline(residuals(itt_health_results_1row))
```

\pagebreak

```{r, echo=T}
itt_health_results_2row <- lmer(outcome ~ treatment_blind + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_dia_data_2row)
```

```{r, fig.width=6, fig.height=3.5}
itt_health_results_2row_days <- lmer(outcome ~ treatment_blind + visit_intervention_diff + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_dia_data_2row)

itt_health_results_2row_visit <- lmer(outcome ~ treatment_blind + sampling_visit + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_dia_data_2row)

plot(itt_health_results_2row, main = "Dia BP 2-Row", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_2row), main = "Dia BP 2-Row")
qqline(residuals(itt_health_results_2row))
```

\pagebreak

```{r, echo=T}
itt_health_results_2row_outliers <- lmer(outcome ~ treatment_blind + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_dia_data_2row_outliers)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_2row_outliers, main = "Dia BP 2-Row, 10 outliers removed", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_2row_outliers), main = "Dia BP 2-Row, 10 outliers removed")
qqline(residuals(itt_health_results_2row_outliers))
```

\pagebreak

```{r, echo=T}
itt_health_results_2row_int <- lmer(outcome ~ treatment_blind*sampling_visit + 
                                    baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                    data = bp_dia_data_2row)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_2row_int, main = "Dia BP 2-Row, Visit interaction", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_2row_int), main = "Dia BP 2-Row, Visit interaction")
qqline(residuals(itt_health_results_2row_int))
```

\pagebreak

```{r, echo=T}
itt_health_results_4row <- lmer(outcome ~ treatment_blind*winter_id + 
                                (1 | adult_id_char:home:cohort:area), 
                                data = bp_dia_data_4row)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_4row, main = "Dia BP 4-Row, Winter interaction", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_4row), main = "Dia BP 4-Row, Winter interaction")
qqline(residuals(itt_health_results_4row))
```

\pagebreak

```{r, fig.width=8, fig.height=6}
# Combine model results
tidy_results_1row <- tidy(itt_health_results_1row, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "1-row",
         sampling_visit = "none") %>% 
  dplyr::select(-p.value)

tidy_results_2row <- tidy(itt_health_results_2row, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row",
         sampling_visit = "none") %>% 
  dplyr::select(-p.value)

tidy_results_2row_outliers <- tidy(itt_health_results_2row_outliers, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row, outliers",
         sampling_visit = "none") %>% 
  dplyr::select(-p.value)

tidy_results_2row_days <- tidy(itt_health_results_2row_days, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row days covar",
         sampling_visit = "days since intervention") %>% 
  dplyr::select(-p.value)

tidy_results_2row_visit <- tidy(itt_health_results_2row_visit, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row visit covar",
         sampling_visit = "covariate") %>% 
  dplyr::select(-p.value)

int_emmeans <- emmeans(itt_health_results_2row_int, revpairwise ~ treatment_blind | sampling_visit)

tidy_results_2row_int <- confint(int_emmeans$contrasts) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((lower.CL), digits = 2),
         conf.high = round((upper.CL), digits = 2)) %>% 
  dplyr::select(contrast, sampling_visit, estimate, conf.low, conf.high) %>% 
  filter(contrast == "Queso - Guac" | contrast == "Salsa - Guac") %>% 
  mutate(contrast = as.character(contrast),
         contrast = if_else(contrast == "Queso - Guac" & sampling_visit == 1, "Queso Visit 1", contrast),
         contrast = if_else(contrast == "Salsa - Guac" & sampling_visit == 1, "Salsa Visit 1", contrast),
         contrast = if_else(contrast == "Queso - Guac" & sampling_visit == 2, "Queso Visit 2", contrast),
         contrast = if_else(contrast == "Salsa - Guac" & sampling_visit == 2, "Salsa Visit 2", contrast),
         contrast = as.factor(contrast),
         model = "2-row visit interaction") %>% 
  rename(term = contrast)

int_emmeans <- emmeans(itt_health_results_4row, revpairwise ~ treatment_blind | winter_id)

tidy_results_4row <- confint(int_emmeans$contrasts) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((lower.CL), digits = 2),
         conf.high = round((upper.CL), digits = 2)) %>% 
  dplyr::select(contrast, winter_id, estimate, conf.low, conf.high) %>% 
  filter((contrast == "Queso - Guac" | contrast == "Salsa - Guac") & winter_id == 2) %>% 
  mutate(contrast = if_else(contrast == "Queso - Guac", "Queso", "Salsa"),
         model = "4-row",
         sampling_visit = "none") %>% 
  rename(term = contrast) %>% 
  dplyr::select(-winter_id)

tidy_results_combined <- rbind(tidy_results_1row, tidy_results_2row, tidy_results_2row_outliers,
                               tidy_results_2row_days, tidy_results_2row_visit,
                               tidy_results_2row_int, tidy_results_4row)

# Calculate ICC
## The ratio of the between-cluster variance to the total variance
## It can also be interpreted as the correlation among observations within the same cluster
icc_model <- lmer(outcome ~ (1 | adult_id_char), bp_dia_data_4row)
bp_dia_icc <- icc(icc_model)

# Plot combined results
itt_plot_estimates <- tidy_results_combined %>%
  mutate(model = as.factor(model)) %>% 
  ggplot(aes(group = term, shape = term)) +
  geom_point(aes(x=model, y=estimate), 
             position = position_dodge(width = 0.5), size = 4) +
  geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
  geom_hline(yintercept = 0) +
  theme_bw() +  
  ggtitle(label = "Dia BP Model comparison for ITT framework",
          subtitle = "ICC for subject = 0.463") +
  labs(y = "Estimate compared to Guac") +
  labs(x = "", group = "", shape = "") +
  theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
itt_plot_estimates
```

\pagebreak

# Forced Vital Capacity models
```{r}
# Prep outcome variables
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
fvc_data_1row <- analysis_data %>% 
  mutate(outcome = fvc) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, adult_id_char, winter_id) %>% 
  mutate(mean_outcome = mean(outcome, na.rm = T)) %>% 
  ungroup() %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, treatment, mean_outcome) %>% 
  distinct(area, adult_id_char, winter_id, .keep_all = T) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "mean_outcome") %>% 
  mutate(mean_outcome = `2`,
         baseline_outcome = `1`) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, mean_outcome = NaN)) %>% 
  ungroup()

fvc_data_2row <- analysis_data %>% 
  mutate(outcome = fvc) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, treatment, outcome,
                visit_intervention_diff) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

# fvc: try removing lines 93, 94, 34, 82, 35, 164, 142, 81, 219, 32
fvc_data_2row_outliers <- fvc_data_2row [-c(93, 94, 34, 82, 35, 164, 142, 81, 219, 32), ]

fvc_data_4row <- analysis_data %>% 
  mutate(outcome = fvc) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  mutate(visit_intervention_diff = if_else(visit_intervention_diff == 9999, 0,
                                           visit_intervention_diff)) %>%  
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, treatment, outcome,
                visit_intervention_diff)

```

```{r, echo=T}
itt_health_results_1row <- lmer(mean_outcome ~ treatment_blind + baseline_outcome + 
                                (1 | home:cohort:area),
                                data = fvc_data_1row)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_1row, main = "fvc 1-Row", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_1row), main = "fvc 1-Row")
qqline(residuals(itt_health_results_1row))
```

\pagebreak

```{r, echo=T}
itt_health_results_2row <- lmer(outcome ~ treatment_blind + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = fvc_data_2row)
```

```{r, fig.width=6, fig.height=3.5}
itt_health_results_2row_days <- lmer(outcome ~ treatment_blind + visit_intervention_diff + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = fvc_data_2row)

itt_health_results_2row_visit <- lmer(outcome ~ treatment_blind + sampling_visit + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = fvc_data_2row)

plot(itt_health_results_2row, main = "fvc 2-Row", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_2row), main = "fvc 2-Row")
qqline(residuals(itt_health_results_2row))
```

\pagebreak

```{r, echo=T}
itt_health_results_2row_outliers <- lmer(outcome ~ treatment_blind + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = fvc_data_2row_outliers)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_2row_outliers, main = "fvc 2-Row, 10 outliers removed", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_2row_outliers), main = "fvc 2-Row, 10 outliers removed")
qqline(residuals(itt_health_results_2row_outliers))
```

\pagebreak

```{r, echo=T}
itt_health_results_2row_int <- lmer(outcome ~ treatment_blind*sampling_visit + 
                                    baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                    data = fvc_data_2row)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_2row_int, main = "fvc 2-Row, Visit interaction", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_2row_int), main = "fvc 2-Row, Visit interaction")
qqline(residuals(itt_health_results_2row_int))
```

\pagebreak

```{r, echo=T}
itt_health_results_4row <- lmer(outcome ~ treatment_blind*winter_id + 
                                (1 | adult_id_char:home:cohort:area), 
                                data = fvc_data_4row)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_4row, main = "fvc 4-Row, Winter interaction", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_4row), main = "fvc 4-Row, Winter interaction")
qqline(residuals(itt_health_results_4row))
```

\pagebreak

```{r, fig.width=8, fig.height=6}
# Combine model results
tidy_results_1row <- tidy(itt_health_results_1row, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "1-row",
         sampling_visit = "none") %>% 
  dplyr::select(-p.value)

tidy_results_2row <- tidy(itt_health_results_2row, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row",
         sampling_visit = "none") %>% 
  dplyr::select(-p.value)

tidy_results_2row_outliers <- tidy(itt_health_results_2row_outliers, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row, outliers",
         sampling_visit = "none") %>% 
  dplyr::select(-p.value)

tidy_results_2row_days <- tidy(itt_health_results_2row_days, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row days covar",
         sampling_visit = "days since intervention") %>% 
  dplyr::select(-p.value)

tidy_results_2row_visit <- tidy(itt_health_results_2row_visit, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row visit covar",
         sampling_visit = "covariate") %>% 
  dplyr::select(-p.value)

int_emmeans <- emmeans(itt_health_results_2row_int, revpairwise ~ treatment_blind | sampling_visit)

tidy_results_2row_int <- confint(int_emmeans$contrasts) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((lower.CL), digits = 2),
         conf.high = round((upper.CL), digits = 2)) %>% 
  dplyr::select(contrast, sampling_visit, estimate, conf.low, conf.high) %>% 
  filter(contrast == "Queso - Guac" | contrast == "Salsa - Guac") %>% 
  mutate(contrast = as.character(contrast),
         contrast = if_else(contrast == "Queso - Guac" & sampling_visit == 1, "Queso Visit 1", contrast),
         contrast = if_else(contrast == "Salsa - Guac" & sampling_visit == 1, "Salsa Visit 1", contrast),
         contrast = if_else(contrast == "Queso - Guac" & sampling_visit == 2, "Queso Visit 2", contrast),
         contrast = if_else(contrast == "Salsa - Guac" & sampling_visit == 2, "Salsa Visit 2", contrast),
         contrast = as.factor(contrast),
         model = "2-row visit interaction") %>% 
  rename(term = contrast)

int_emmeans <- emmeans(itt_health_results_4row, revpairwise ~ treatment_blind | winter_id)

tidy_results_4row <- confint(int_emmeans$contrasts) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((lower.CL), digits = 2),
         conf.high = round((upper.CL), digits = 2)) %>% 
  dplyr::select(contrast, winter_id, estimate, conf.low, conf.high) %>% 
  filter((contrast == "Queso - Guac" | contrast == "Salsa - Guac") & winter_id == 2) %>% 
  mutate(contrast = if_else(contrast == "Queso - Guac", "Queso", "Salsa"),
         model = "4-row",
         sampling_visit = "none") %>% 
  rename(term = contrast) %>% 
  dplyr::select(-winter_id)

tidy_results_combined <- rbind(tidy_results_1row, tidy_results_2row, tidy_results_2row_outliers,
                               tidy_results_2row_days, tidy_results_2row_visit,
                               tidy_results_2row_int, tidy_results_4row)

# Calculate ICC
## The ratio of the between-cluster variance to the total variance
## It can also be interpreted as the correlation among observations within the same cluster
icc_model <- lmer(outcome ~ (1 | adult_id_char), fvc_data_4row)
fvc_icc <- icc(icc_model)

# Plot combined results
itt_plot_estimates <- tidy_results_combined %>%
  mutate(model = as.factor(model)) %>% 
  ggplot(aes(group = term, shape = term)) +
  geom_point(aes(x=model, y=estimate), 
             position = position_dodge(width = 0.5), size = 4) +
  geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
  geom_hline(yintercept = 0) +
  theme_bw() +  
  ggtitle(label = "FVC Model comparison for ITT framework",
          subtitle = "ICC for subject = 0.799") +
  labs(y = "Estimate compared to Guac") +
  labs(x = "", group = "", shape = "") +
  theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
itt_plot_estimates
```

\pagebreak

# Forced Exporatory Volume models
```{r}
# Prep outcome variables
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
fev1_data_1row <- analysis_data %>% 
  mutate(outcome = fev1) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, adult_id_char, winter_id) %>% 
  mutate(mean_outcome = mean(outcome, na.rm = T)) %>% 
  ungroup() %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, treatment, mean_outcome) %>% 
  distinct(area, adult_id_char, winter_id, .keep_all = T) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "mean_outcome") %>% 
  mutate(mean_outcome = `2`,
         baseline_outcome = `1`) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, mean_outcome = NaN)) %>% 
  ungroup()

fev1_data_2row <- analysis_data %>% 
  mutate(outcome = fev1) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, treatment, outcome,
                visit_intervention_diff) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

# fev1: try removing lines 79, 34, 24, 80, 25, 82, 26, 230, 5, 6
fev1_data_2row_outliers <- fev1_data_2row [-c(79, 34, 24, 80, 25, 82, 26, 230, 5, 6), ]

fev1_data_4row <- analysis_data %>% 
  mutate(outcome = fev1) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  mutate(visit_intervention_diff = if_else(visit_intervention_diff == 9999, 0,
                                           visit_intervention_diff)) %>%  
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, treatment, outcome,
                visit_intervention_diff)

```

```{r, echo=T}
itt_health_results_1row <- lmer(mean_outcome ~ treatment_blind + baseline_outcome + 
                                (1 | home:cohort:area),
                                data = fev1_data_1row)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_1row, main = "fev1 1-Row", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_1row), main = "fev1 1-Row")
qqline(residuals(itt_health_results_1row))
```

\pagebreak

```{r, echo=T}
itt_health_results_2row <- lmer(outcome ~ treatment_blind + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = fev1_data_2row)
```

```{r, fig.width=6, fig.height=3.5}
itt_health_results_2row_days <- lmer(outcome ~ treatment_blind + visit_intervention_diff + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = fev1_data_2row)

itt_health_results_2row_visit <- lmer(outcome ~ treatment_blind + sampling_visit + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = fev1_data_2row)

plot(itt_health_results_2row, main = "fev1 2-Row", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_2row), main = "fev1 2-Row")
qqline(residuals(itt_health_results_2row))
```

\pagebreak

```{r, echo=T}
itt_health_results_2row_outliers <- lmer(outcome ~ treatment_blind + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = fev1_data_2row_outliers)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_2row_outliers, main = "fev1 2-Row, 10 outliers removed", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_2row_outliers), main = "fev1 2-Row, 10 outliers removed")
qqline(residuals(itt_health_results_2row_outliers))
```

\pagebreak

```{r, echo=T}
itt_health_results_2row_int <- lmer(outcome ~ treatment_blind*sampling_visit + 
                                    baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                    data = fev1_data_2row)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_2row_int, main = "fev1 2-Row, Visit interaction", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_2row_int), main = "fev1 2-Row, Visit interaction")
qqline(residuals(itt_health_results_2row_int))
```

\pagebreak

```{r, echo=T}
itt_health_results_4row <- lmer(outcome ~ treatment_blind*winter_id + 
                                (1 | adult_id_char:home:cohort:area), 
                                data = fev1_data_4row)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_4row, main = "fev1 4-Row, Winter interaction", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_4row), main = "fev1 4-Row, Winter interaction")
qqline(residuals(itt_health_results_4row))
```

\pagebreak

```{r, fig.width=8, fig.height=6}
# Combine model results
tidy_results_1row <- tidy(itt_health_results_1row, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "1-row",
         sampling_visit = "none") %>% 
  dplyr::select(-p.value)

tidy_results_2row <- tidy(itt_health_results_2row, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row",
         sampling_visit = "none") %>% 
  dplyr::select(-p.value)

tidy_results_2row_outliers <- tidy(itt_health_results_2row_outliers, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row, outliers",
         sampling_visit = "none") %>% 
  dplyr::select(-p.value)

tidy_results_2row_days <- tidy(itt_health_results_2row_days, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row days covar",
         sampling_visit = "days since intervention") %>% 
  dplyr::select(-p.value)

tidy_results_2row_visit <- tidy(itt_health_results_2row_visit, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row visit covar",
         sampling_visit = "covariate") %>% 
  dplyr::select(-p.value)

int_emmeans <- emmeans(itt_health_results_2row_int, revpairwise ~ treatment_blind | sampling_visit)

tidy_results_2row_int <- confint(int_emmeans$contrasts) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((lower.CL), digits = 2),
         conf.high = round((upper.CL), digits = 2)) %>% 
  dplyr::select(contrast, sampling_visit, estimate, conf.low, conf.high) %>% 
  filter(contrast == "Queso - Guac" | contrast == "Salsa - Guac") %>% 
  mutate(contrast = as.character(contrast),
         contrast = if_else(contrast == "Queso - Guac" & sampling_visit == 1, "Queso Visit 1", contrast),
         contrast = if_else(contrast == "Salsa - Guac" & sampling_visit == 1, "Salsa Visit 1", contrast),
         contrast = if_else(contrast == "Queso - Guac" & sampling_visit == 2, "Queso Visit 2", contrast),
         contrast = if_else(contrast == "Salsa - Guac" & sampling_visit == 2, "Salsa Visit 2", contrast),
         contrast = as.factor(contrast),
         model = "2-row visit interaction") %>% 
  rename(term = contrast)

int_emmeans <- emmeans(itt_health_results_4row, revpairwise ~ treatment_blind | winter_id)

tidy_results_4row <- confint(int_emmeans$contrasts) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((lower.CL), digits = 2),
         conf.high = round((upper.CL), digits = 2)) %>% 
  dplyr::select(contrast, winter_id, estimate, conf.low, conf.high) %>% 
  filter((contrast == "Queso - Guac" | contrast == "Salsa - Guac") & winter_id == 2) %>% 
  mutate(contrast = if_else(contrast == "Queso - Guac", "Queso", "Salsa"),
         model = "4-row",
         sampling_visit = "none") %>% 
  rename(term = contrast) %>% 
  dplyr::select(-winter_id)

tidy_results_combined <- rbind(tidy_results_1row, tidy_results_2row, tidy_results_2row_outliers,
                               tidy_results_2row_days, tidy_results_2row_visit,
                               tidy_results_2row_int, tidy_results_4row)

# Calculate ICC
## The ratio of the between-cluster variance to the total variance
## It can also be interpreted as the correlation among observations within the same cluster
icc_model <- lmer(outcome ~ (1 | adult_id_char), fev1_data_4row)
fev1_icc <- icc(icc_model)

# Plot combined results
itt_plot_estimates <- tidy_results_combined %>%
  mutate(model = as.factor(model)) %>% 
  ggplot(aes(group = term, shape = term)) +
  geom_point(aes(x=model, y=estimate), 
             position = position_dodge(width = 0.5), size = 4) +
  geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
  geom_hline(yintercept = 0) +
  theme_bw() +  
  ggtitle(label = "FEV1 Model comparison for ITT framework",
          subtitle = "ICC for subject = 0.796") +
  labs(y = "Estimate compared to Guac") +
  labs(x = "", group = "", shape = "") +
  theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
itt_plot_estimates
```