### 2021-04-26 main_analysis_final.Rmd
main analysis updates

---
title: 'EldersAIR: Main analysis - final version and manuscript results'
author: "Ethan Walker"
date: "Started 16 April 2021; Updated 26 April 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, include = TRUE,
                      message = FALSE, warning = FALSE)
```

```{r}
library(readxl)
library(naniar)
library(lubridate)
library(emmeans)
library(broom)
library(broom.mixed)
library(zoo)
library(car)
library(lme4)
library(lmerTest)
library(mctest)
library(tidyverse)
library(knitr)
library(kableExtra)
#library(MASS)
library(faraway)
library(splines)
library(influence.ME)
library(sjstats)
```

```{r}
# Load data

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) 
```

```{r}
analysis_data <- elders_cleaned_dataset_by_visit %>% 
  mutate(treatment = factor(treatment,
                            levels = c("Placebo", "Filter", "Education"),
                            labels = c("Placebo", "Filter", "Education")),
         visit_intervention_diff = round(difftime(sampling_date, intervention_date, 
                                            units = "days"), digits = 0),
         visit_intervention_diff = as.numeric(visit_intervention_diff),
         visit_intervention_diff = if_else(winter_id == 1, 9999,
                                           visit_intervention_diff)) 
```

# ITT Model Code - Data formatted as 2 rows per person  
outcome = outcome from Winter 2 visits (up to 2 repeated measures)  
baseline_outcome = mean outcome from Winter 1

# Systolic Blood Pressure
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
bp_sys_data_2row <- analysis_data %>% 
  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = bp_sys) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

itt_health_results_primary <- lmer(outcome ~ treatment + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_sys_data_2row)

#isSingular(itt_health_results_primary)

#summary(itt_health_results_primary)

tidy_results_primary <- tidy(itt_health_results_primary, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "Primary") %>% 
  dplyr::select(-p.value)
kable(tidy_results_primary)

itt_health_results_confounders <- lmer(outcome ~ treatment + 
                                education + residents_smoke + home_sqm_2level +
                                burn_level + gender + chimney_clean + stove_grade + 
                                age + bmi + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_sys_data_2row)

tidy_results_confounders <- tidy(itt_health_results_confounders, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "Confounders") %>% 
  dplyr::select(-p.value)

itt_health_results_nohome <- lmer(outcome ~ treatment + 
                                baseline_outcome + (1 | adult_id_char:cohort:area), 
                                data = bp_sys_data_2row)

tidy_results_nohome <- tidy(itt_health_results_nohome, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "No Home Term") %>% 
  dplyr::select(-p.value)

# Remove 10 highest outliers based on Cook's distance
bp_sys_data_outliers <- bp_sys_data_2row [-c(95, 48, 127, 17, 96, 128, 202, 254, 109, 216), ]

itt_health_results_outliers <- lmer(outcome ~ treatment + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_sys_data_outliers)

tidy_results_outliers <- tidy(itt_health_results_outliers, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "Outliers Removed") %>% 
  dplyr::select(-p.value)

tidy_results_combined <- rbind(tidy_results_primary, tidy_results_confounders,
                               tidy_results_nohome, tidy_results_outliers)

# Calculate ICC
## The ratio of the between-cluster variance to the total variance
## It can also be interpreted as the correlation among observations within the same cluster
icc_model <- lmer(outcome ~ (1 | adult_id_char), bp_sys_data_2row)
bp_sys_icc <- icc(icc_model)

itt_plot_estimates <- tidy_results_combined %>%
  mutate(model = factor(model,
                        levels = c("Primary", "Confounders", "No Home Term",
                                   "Outliers Removed"))) %>% 
  ggplot(aes(group = term, shape = term)) +
    geom_point(aes(x=model, y=estimate), 
               position = position_dodge(width = 0.5), size = 4) +
    geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
                  position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
    geom_hline(yintercept = 0) +
    theme_bw() +  
    ggtitle(label = "Sys BP Model - ITT framework",
            subtitle = "ICC for subject = 0.58") +
    labs(y = "Estimate compared to Placebo") +
    labs(x = "", group = "", shape = "") +
    theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
itt_plot_estimates

plot(itt_health_results_primary, main = "ITT Primary Model", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_primary), main = "ITT Primary Model")
qqline(residuals(itt_health_results_primary))
```

\pagebreak

# Diastolic Blood Pressure
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
bp_dia_data_2row <- analysis_data %>% 
  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = bp_dia) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

itt_health_results_primary <- lmer(outcome ~ treatment + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_dia_data_2row)

#isSingular(itt_health_results_primary)

#summary(itt_health_results_primary)

tidy_results_primary <- tidy(itt_health_results_primary, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "Primary") %>% 
  dplyr::select(-p.value)
kable(tidy_results_primary)

itt_health_results_confounders <- lmer(outcome ~ treatment + 
                                education + residents_smoke + home_sqm_2level +
                                burn_level + gender + chimney_clean + stove_grade + 
                                age + bmi + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_dia_data_2row)

tidy_results_confounders <- tidy(itt_health_results_confounders, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "Confounders") %>% 
  dplyr::select(-p.value)

itt_health_results_nohome <- lmer(outcome ~ treatment + 
                                baseline_outcome + (1 | adult_id_char:cohort:area), 
                                data = bp_dia_data_2row)

tidy_results_nohome <- tidy(itt_health_results_nohome, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "No Home Term") %>% 
  dplyr::select(-p.value)

# Remove 10 highest outliers based on Cook's distance
bp_dia_data_outliers <- bp_dia_data_2row [-c(42, 95, 48, 192, 127, 223, 202, 128, 204, 52), ]

itt_health_results_outliers <- lmer(outcome ~ treatment + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_dia_data_outliers)

tidy_results_outliers <- tidy(itt_health_results_outliers, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "Outliers Removed") %>% 
  dplyr::select(-p.value)

tidy_results_combined <- rbind(tidy_results_primary, tidy_results_confounders,
                               tidy_results_nohome, tidy_results_outliers)

# Calculate ICC
## The ratio of the between-cluster variance to the total variance
## It can also be interpreted as the correlation among observations within the same cluster
icc_model <- lmer(outcome ~ (1 | adult_id_char), bp_dia_data_2row)
bp_dia_icc <- icc(icc_model)

itt_plot_estimates <- tidy_results_combined %>%
  mutate(model = factor(model,
                        levels = c("Primary", "Confounders", "No Home Term",
                                   "Outliers Removed"))) %>% 
  ggplot(aes(group = term, shape = term)) +
    geom_point(aes(x=model, y=estimate), 
               position = position_dodge(width = 0.5), size = 4) +
    geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
                  position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
    geom_hline(yintercept = 0) +
    theme_bw() +  
    ggtitle(label = "Dia BP Model - ITT framework",
            subtitle = "ICC for subject = 0.61") +
    labs(y = "Estimate compared to Placebo") +
    labs(x = "", group = "", shape = "") +
    theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
itt_plot_estimates

plot(itt_health_results_primary, main = "ITT Primary Model", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_primary), main = "ITT Primary Model")
qqline(residuals(itt_health_results_primary))
```

\pagebreak

# FVC
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
fvc_data_2row <- analysis_data %>% 
  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = fvc) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

fvc_data_acceptable <- analysis_data %>% 
  filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = fvc) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

itt_health_results_primary <- lmer(outcome ~ treatment + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = fvc_data_2row)

#isSingular(itt_health_results_primary)

#summary(itt_health_results_primary)

tidy_results_primary <- tidy(itt_health_results_primary, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "Primary") %>% 
  dplyr::select(-p.value)
kable(tidy_results_primary)

itt_health_results_acceptable <- lmer(outcome ~ treatment + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = fvc_data_acceptable)

tidy_results_acceptable <- tidy(itt_health_results_acceptable, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "Acceptable Sessions") %>% 
  dplyr::select(-p.value)

itt_health_results_confounders <- lmer(outcome ~ treatment + 
                                education + residents_smoke + home_sqm_2level +
                                burn_level + gender + chimney_clean + stove_grade + 
                                age + bmi + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = fvc_data_2row)

tidy_results_confounders <- tidy(itt_health_results_confounders, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "Confounders") %>% 
  dplyr::select(-p.value)

itt_health_results_nohome <- lmer(outcome ~ treatment + 
                                baseline_outcome + (1 | adult_id_char:cohort:area), 
                                data = fvc_data_2row)

tidy_results_nohome <- tidy(itt_health_results_nohome, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "No Home Term") %>% 
  dplyr::select(-p.value)

# Remove 10 highest outliers based on Cook's distance
fvc_data_outliers <- fvc_data_2row [-c(93, 94, 82, 34, 142, 164, 35, 219, 89, 81), ]

itt_health_results_outliers <- lmer(outcome ~ treatment + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = fvc_data_outliers)

tidy_results_outliers <- tidy(itt_health_results_outliers, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "Outliers Removed") %>% 
  dplyr::select(-p.value)

tidy_results_combined <- rbind(tidy_results_primary, tidy_results_acceptable, 
                               tidy_results_confounders,
                               tidy_results_nohome, tidy_results_outliers)

# Calculate ICC
## The ratio of the between-cluster variance to the total variance
## It can also be interpreted as the correlation among observations within the same cluster
icc_model <- lmer(outcome ~ (1 | adult_id_char), fvc_data_2row)
fvc_icc <- icc(icc_model)

itt_plot_estimates <- tidy_results_combined %>%
  mutate(model = factor(model,
                        levels = c("Primary", "Acceptable Sessions",
                                   "Confounders", "No Home Term",
                                   "Outliers Removed"))) %>% 
  ggplot(aes(group = term, shape = term)) +
    geom_point(aes(x=model, y=estimate), 
               position = position_dodge(width = 0.5), size = 4) +
    geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
                  position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
    geom_hline(yintercept = 0) +
    theme_bw() +  
    ggtitle(label = "FVC Model - ITT framework",
            subtitle = "ICC for subject = 0.86") +
    labs(y = "Estimate compared to Placebo") +
    labs(x = "", group = "", shape = "") +
    theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
itt_plot_estimates

plot(itt_health_results_primary, main = "ITT Primary Model", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_primary), main = "ITT Primary Model")
qqline(residuals(itt_health_results_primary))
```

\pagebreak

# FEV1
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
fev1_data_2row <- analysis_data %>% 
  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = fev1) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

fev1_data_acceptable <- analysis_data %>% 
  filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = fev1) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

itt_health_results_primary <- lmer(outcome ~ treatment + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = fev1_data_2row)

#isSingular(itt_health_results_primary)

#summary(itt_health_results_primary)

tidy_results_primary <- tidy(itt_health_results_primary, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "Primary") %>% 
  dplyr::select(-p.value)
kable(tidy_results_primary)

itt_health_results_acceptable <- lmer(outcome ~ treatment + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = fev1_data_acceptable)

tidy_results_acceptable <- tidy(itt_health_results_acceptable, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "Acceptable Sessions") %>% 
  dplyr::select(-p.value)

itt_health_results_confounders <- lmer(outcome ~ treatment + 
                                education + residents_smoke + home_sqm_2level +
                                burn_level + gender + chimney_clean + stove_grade + 
                                age + bmi + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = fev1_data_2row)

tidy_results_confounders <- tidy(itt_health_results_confounders, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "Confounders") %>% 
  dplyr::select(-p.value)

itt_health_results_nohome <- lmer(outcome ~ treatment + 
                                baseline_outcome + (1 | adult_id_char:cohort:area), 
                                data = fev1_data_2row)

tidy_results_nohome <- tidy(itt_health_results_nohome, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "No Home Term") %>% 
  dplyr::select(-p.value)

# Remove 10 highest outliers based on Cook's distance
fev1_data_outliers <- fev1_data_2row [-c(79, 219, 34, 82, 80, 24, 142, 81, 164, 14), ]

itt_health_results_outliers <- lmer(outcome ~ treatment + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = fev1_data_outliers)

tidy_results_outliers <- tidy(itt_health_results_outliers, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "Outliers Removed") %>% 
  dplyr::select(-p.value)

tidy_results_combined <- rbind(tidy_results_primary, tidy_results_acceptable, 
                               tidy_results_confounders,
                               tidy_results_nohome, tidy_results_outliers)

# Calculate ICC
## The ratio of the between-cluster variance to the total variance
## It can also be interpreted as the correlation among observations within the same cluster
icc_model <- lmer(outcome ~ (1 | adult_id_char), fev1_data_2row)
fev1_icc <- icc(icc_model)

itt_plot_estimates <- tidy_results_combined %>%
  mutate(model = factor(model,
                        levels = c("Primary", "Acceptable Sessions",
                                   "Confounders", "No Home Term",
                                   "Outliers Removed"))) %>% 
  ggplot(aes(group = term, shape = term)) +
    geom_point(aes(x=model, y=estimate), 
               position = position_dodge(width = 0.5), size = 4) +
    geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
                  position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
    geom_hline(yintercept = 0) +
    theme_bw() +  
    ggtitle(label = "FEV1 Model - ITT framework",
            subtitle = "ICC for subject = 0.91") +
    labs(y = "Estimate compared to Placebo") +
    labs(x = "", group = "", shape = "") +
    theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
itt_plot_estimates

plot(itt_health_results_primary, main = "ITT Primary Model", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_primary), main = "ITT Primary Model")
qqline(residuals(itt_health_results_primary))
```

\pagebreak

# PM2.5
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
pm_data_2row <- analysis_data %>% 
  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = pm_mean_visit) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

pm_at_home_data <- analysis_data %>% 
  mutate(outcome = pm_mean_at_home_visit) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

itt_pm_results_primary <- lmer(log(outcome) ~ treatment + 
                                baseline_outcome + (1 | home:cohort:area), 
                                data = pm_data_2row)

#isSingular(itt_pm_results_primary)

#summary(itt_pm_results_primary)

tidy_results_primary <- tidy(itt_pm_results_primary, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
         conf.low = round((exp(conf.low)-1)*100, digits = 2),
         conf.high = round((exp(conf.high)-1)*100, digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "Primary") %>% 
  dplyr::select(-p.value)
kable(tidy_results_primary)

itt_pm_results_athome <- lmer(log(outcome) ~ treatment + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = pm_at_home_data)

tidy_results_athome <- tidy(itt_pm_results_athome, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
         conf.low = round((exp(conf.low)-1)*100, digits = 2),
         conf.high = round((exp(conf.high)-1)*100, digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "At-home PM") %>% 
  dplyr::select(-p.value)

itt_pm_results_confounders <- lmer(log(outcome) ~ treatment + 
                                education + residents_smoke + home_sqm_2level +
                                burn_level + gender + chimney_clean + stove_grade + 
                                age + bmi + 
                                baseline_outcome + (1 | home:cohort:area), 
                                data = pm_data_2row)

tidy_results_confounders <- tidy(itt_pm_results_confounders, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
         conf.low = round((exp(conf.low)-1)*100, digits = 2),
         conf.high = round((exp(conf.high)-1)*100, digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "Confounders") %>% 
  dplyr::select(-p.value)

tidy_results_combined <- rbind(tidy_results_primary, tidy_results_athome, 
                               tidy_results_confounders)

# Calculate ICC
## The ratio of the between-cluster variance to the total variance
## It can also be interpreted as the correlation among observations within the same cluster
icc_model <- lmer(outcome ~ (1 | home), pm_data_2row)
pm_icc <- icc(icc_model)

itt_plot_estimates <- tidy_results_combined %>%
  mutate(model = factor(model,
                        levels = c("Primary", "At-home PM",
                                   "Confounders"))) %>% 
  ggplot(aes(group = term, shape = term)) +
    geom_point(aes(x=model, y=estimate), 
               position = position_dodge(width = 0.5), size = 4) +
    geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
                  position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
    geom_hline(yintercept = 0) +
    theme_bw() +  
    ggtitle(label = "PM Model - ITT framework",
            subtitle = "ICC for subject = 0.44") +
    labs(y = "% difference compared to Placebo") +
    labs(x = "", group = "", shape = "") +
    theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
itt_plot_estimates

plot(itt_pm_results_primary, main = "ITT Primary Model", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_pm_results_primary), main = "ITT Primary Model")
qqline(residuals(itt_pm_results_primary))
```

```{r, eval=F}
# ITT primary model identify potential influential datapoints

influential <- influence(itt_health_results_primary, obs = TRUE)
cooks <- cooks.distance.estex(influential, sort = TRUE)
plot.estex(influential, which = "cook")

# bp_sys: try removing lines 95, 48, 127, 17, 96, 128, 202, 254, 109, 216
# bp_dia: try removing lines 42, 95, 48, 192, 127, 223, 202, 128, 204, 52
# fvc: try removing lines 93, 94, 82, 34, 142, 164, 35, 219, 89, 81
# fev1: try removing lines 79, 219, 34, 82, 80, 24, 142, 81, 164, 14

# fev1_fvc_ratio: try removing lines 67, 113, 140, 49, 82, 27, 4, 3, 105, 151
# fvc_pp: try removing lines 27, 60, 66, 137, 41, 140, 42, 115, 113, 90
# fev1_pp: try removing lines 27, 113, 60, 67, 137, 66, 42, 41, 115, 90
```

```{r, eval=F}
# ITT Effect modification model code

itt_health_results_2row_int <- lmer(outcome ~ treatment_blind*sampling_visit + 
                                    baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                    data = analysis_data_2row)
```