### 2021-05-26 manuscript_tables_figures.Rmd

---
title: 'EldersAIR: manuscript tables/figures'
author: "Ethan Walker"
date: "Started 1 March 2021, Updated 24 May 2021"
output:
  powerpoint_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE,
                      fig.height = 6, fig.width = 8)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
               "#0072B2", "#D55E00", "#CC79A7")
```

```{r}
# Load data

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

treatments_blind_nn <- read_xlsx(paste0(file_path, "Input/NN/treatments_blind.xlsx")) %>% 
  mutate(area = "NN")

treatments_blind_npt <- read_xlsx(paste0(file_path, "Input/NP/treatments_blind.xlsx")) %>% 
  mutate(area = "NPT")

treatments_blind <- rbind(treatments_blind_nn, treatments_blind_npt) %>% 
  mutate(home_id_num = as.factor(HomeID),
         treatment_blind = as.factor(CodedCondition),
         area = as.factor(area)) %>% 
  select(area, home_id_num, treatment_blind)

elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
  left_join(treatments_blind, by = c("area", "home_id_num")) 
```

# Number of homes in Figure 1
```{r}
# Step 1: Winter 1 homes
demographics_check <- elders_demographics %>% 
  filter(winter_id == 1) %>% 
  distinct(area, home, .keep_all = T) %>% 
  count(area)

demographics_check

# Step 2: Randomized homes
demographics_check <- elders_demographics_full %>% 
  arrange(area, home, winter_id) %>% 
  mutate(filter_var = if_else(home == "NPE5197" | home == "NPE5198" | home == "NPE5199" |
                              home == "FDE314" | home == "FDE316" | home == "FDE318" |
                              home == "FDE413" | home == "FDE412B" | home == "NPE5314" |
                              home == "NPE5317", 1, 0)) %>% 
  filter(filter_var == 0 | is.na(filter_var)) %>% 
  filter(!is.na(home)) %>% 
  filter(!is.na(treatment)) %>% 
  filter(winter_id == 2) %>% 
  distinct(area, home, .keep_all = T) %>% 
  count(treatment, area)

demographics_check

# Step 3: Winter 2 homes
demographics_check <- elders_demographics %>% 
  filter(winter_id == 2) %>% 
  distinct(area, home, .keep_all = T) %>% 
  count(treatment, area)

demographics_check
```

# Number of participants in Figure 1
```{r}
# Step 1: Winter 1 participants
demographics_check <- elders_demographics %>% 
  filter(winter_id == 1) %>% 
  distinct(area, adult_id_char, .keep_all = T) %>% 
  count(area)

demographics_check

# Step 2: Randomized participants
demographics_check <- elders_demographics_full %>% 
  arrange(area, home, winter_id) %>% 
  mutate(filter_var = if_else(home == "NPE5197" | home == "NPE5198" | home == "NPE5199" |
                              home == "FDE314" | home == "FDE316" | home == "FDE318" |
                              home == "FDE413" | home == "FDE412B" | home == "NPE5314" |
                              home == "NPE5317", 1, 0)) %>% 
  filter(filter_var == 0 | is.na(filter_var)) %>% 
  filter(!is.na(home)) %>% 
  filter(!is.na(treatment)) %>% 
  filter(winter_id == 2) %>% 
  distinct(area, adult_id_char, .keep_all = T) %>% 
  count(treatment, area)

demographics_check

# Step 3: Winter 2 participants
demographics_check <- elders_demographics %>% 
  filter(winter_id == 2) %>% 
  distinct(area, adult_id_char, .keep_all = T) %>% 
  count(treatment, area)

demographics_check
```

# Number of med visits in Figure 1
```{r}
# Step 1: Winter 1 visits
medvisit_check <- elders_medvisit_clean %>% 
  filter(winter_id == 1 & !is.na(sampling_visit)) %>% 
  filter_at(vars(sp02:spiro_result), any_vars(!is.na(.))) %>% 
  count(area)

medvisit_check

# Step 2: Winter 2 visits
medvisit_check <- elders_medvisit_clean %>% 
  filter(!is.na(treatment)) %>% 
  filter(winter_id == 2 & !is.na(sampling_visit)) %>% 
  filter_at(vars(sp02:spiro_result), any_vars(!is.na(.))) %>% 
  count(treatment, area)

medvisit_check
```

# Number of exposure samples in Figure 1
```{r} 
# Step 1: Winter 1 samples
exposure_check <- elders_cleaned_dataset_by_visit %>% 
  arrange(area, home, winter_id, sampling_visit) %>% 
  distinct(area, home, winter_id, sampling_visit, .keep_all = T) %>% 
  filter(!is.na(mp_mean_visit)) %>% # mp_mean_visit or pm_mean_visit
  filter(winter_id == 1) %>% 
  count(area)

exposure_check

# Step 2: Winter 2 samples
exposure_check <- elders_cleaned_dataset_by_visit %>% 
  filter(!is.na(treatment)) %>% 
  arrange(area, home, winter_id, sampling_visit) %>% 
  distinct(area, home, winter_id, sampling_visit, .keep_all = T) %>% 
  filter(!is.na(mp_mean_visit)) %>% # mp_mean_visit or pm_mean_visit
  filter(winter_id == 2) %>% 
  count(treatment, area)

exposure_check
```

```{r}
# summary data for manuscript tables
char_funct_total <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  rename(new_var = var) %>% 
  #filter(treatment_assigned == "Placebo") %>%
  #filter(area == "NPT") %>% 
  arrange(home, winter_id, sampling_visit) %>% 
  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
  mutate(total_n = n()) %>% 
  group_by(area) %>% 
  mutate(area_n = n()) %>% 
  group_by(new_var) %>% 
  mutate(n = n(),
         percent = n()/total_n) %>% 
  distinct(new_var, .keep_all = TRUE) %>% 
  dplyr::select(new_var, n, percent) %>% 
  arrange(new_var)
demographics_summary
}

char_funct_treatment <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  rename(new_var = var) %>% 
  #filter(treatment_assigned == "Placebo") %>% 
  #filter(area == "NPT") %>% 
  arrange(home, winter_id, sampling_visit) %>% 
  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
  mutate(total_n = n()) %>% 
  group_by(treatment) %>% 
  mutate(treatment_n = n()) %>% 
  group_by(treatment, new_var) %>% 
  mutate(n = n(),
         percent = n()/treatment_n) %>% 
  distinct(new_var, .keep_all = TRUE) %>% 
  dplyr::select(new_var, n, percent) %>% 
  arrange(treatment, new_var)
demographics_summary
}

# distinct by participant
char_funct_total("gender")
char_funct_treatment("gender")
char_funct_total("hispanic")
char_funct_treatment("hispanic")
char_funct_total("race")
char_funct_treatment("race")
char_funct_total("education")
char_funct_treatment("education")

#distinct by home
char_funct_total("residents_over55")
char_funct_treatment("residents_over55")
char_funct_total("income_3level")
char_funct_treatment("income_3level")
char_funct_total("home_floors_2level")
char_funct_treatment("home_floors_2level")
char_funct_total("home_sqm_2level")
char_funct_treatment("home_sqm_2level")
char_funct_total("home_year_built_2level")
char_funct_treatment("home_year_built_2level")
char_funct_total("home_pets_3level")
char_funct_treatment("home_pets_3level")
char_funct_total("residents_smoke")
char_funct_treatment("residents_smoke")
char_funct_total("residents_smoke_inside")
char_funct_treatment("residents_smoke_inside")

char_funct_total("stove_age")
char_funct_treatment("stove_age")
char_funct_total("chimney_clean")
char_funct_treatment("chimney_clean")
char_funct_total("wood_collect_method")
char_funct_treatment("wood_collect_method")
char_funct_total("wood_collect")
char_funct_treatment("wood_collect")
char_funct_total("stove_grade")
char_funct_treatment("stove_grade")
char_funct_total("burn_level")
char_funct_treatment("burn_level")
```

```{r}
# summary data for manuscript tables - 1 response from Winter 1
num_funct_total <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  rename(new_var = var) %>% 
  mutate(new_var = as.numeric(new_var)) %>% 
  #mutate(new_var = (as.numeric(new_var)-32)*5/9) %>% 
  #filter(area == "NPT") %>% 
  filter(winter_id == 1) %>% 
  arrange(home, winter_id, sampling_visit) %>% 
  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
  filter(!is.na(new_var)) %>% 
  dplyr::select(new_var) %>% 
  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
demographics_summary
}

num_funct_treatment <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  rename(new_var = var) %>%
  mutate(new_var = as.numeric(new_var)) %>% 
  #mutate(new_var = (as.numeric(new_var)-32)*5/9) %>%
  #filter(area == "NPT") %>% 
  filter(winter_id == 1) %>% 
  arrange(home, winter_id, sampling_visit) %>% 
  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
  filter(!is.na(new_var)) %>% 
  dplyr::select(treatment, new_var) %>% 
  group_by(treatment) %>% 
  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
demographics_summary
}

# distinct by participant
num_funct_total("age")
num_funct_treatment("age")

num_funct_total("sum_act_smoking")
num_funct_treatment("sum_act_smoking")
num_funct_total("sum_act_windows")
num_funct_treatment("sum_act_windows")
num_funct_total("sum_act_sweep")
num_funct_treatment("sum_act_sweep")
num_funct_total("sum_act_door")
num_funct_treatment("sum_act_door")

# distinct by home
num_funct_total("home_bedrooms")
num_funct_treatment("home_bedrooms")

# summary data for manuscript tables - averaged over multiple Winter 1 visits
num_funct_total <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  filter(area == "NN") %>% 
  rename(new_var = var) %>% 
  mutate(new_var = as.numeric(new_var)) %>% 
  group_by(home, winter_id) %>% 
  mutate(mean_var = mean(new_var, na.rm = T)) %>% 
  ungroup() %>% 
  filter(winter_id == 1) %>% 
  arrange(home, winter_id, sampling_visit) %>% 
  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
  filter(!is.na(mean_var)) %>% 
  dplyr::select(mean_var) %>% 
  summarize(mean_new_var = mean(mean_var), sd_var = sd(mean_var), n_var = n(), 
            min_var = min(mean_var), med_var = median(mean_var), max_var = max(mean_var))
demographics_summary
}

num_funct_treatment <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  filter(area == "NN") %>% 
  rename(new_var = var) %>%
  mutate(new_var = as.numeric(new_var)) %>% 
  group_by(home, winter_id) %>% 
  mutate(mean_var = mean(new_var, na.rm = T)) %>% 
  ungroup() %>% 
  filter(winter_id == 1) %>% 
  arrange(home, winter_id, sampling_visit) %>% 
  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
  filter(!is.na(mean_var)) %>% 
  dplyr::select(treatment, mean_var) %>% 
  group_by(treatment) %>% 
  summarize(mean_new_var = mean(mean_var), sd_var = sd(mean_var), n_var = n(), 
            min_var = min(mean_var), med_var = median(mean_var), max_var = max(mean_var))
demographics_summary
}

# distinct by participant
num_funct_total("bp_sys")
num_funct_treatment("bp_sys")
num_funct_total("bp_dia")
num_funct_treatment("bp_dia")
num_funct_total("bmi")
num_funct_treatment("bmi")
num_funct_total("fvc")
num_funct_treatment("fvc")
num_funct_total("fev1")
num_funct_treatment("fev1")
num_funct_total("fev1_fvc_ratio")
num_funct_treatment("fev1_fvc_ratio")
num_funct_total("fvc_pp")
num_funct_treatment("fvc_pp")
num_funct_total("fev1_pp")
num_funct_treatment("fev1_pp")
num_funct_total("mp_mean_winter")
num_funct_treatment("mp_mean_winter")

num_funct_total("home_act_smoking")
num_funct_treatment("home_act_smoking")
num_funct_total("home_act_windows")
num_funct_treatment("home_act_windows")
num_funct_total("home_act_sweep")
num_funct_treatment("home_act_sweep")
num_funct_total("home_act_door")
num_funct_treatment("home_act_door")

# distinct by home
num_funct_total("pm_mean_winter")
num_funct_treatment("pm_mean_winter")
num_funct_total("moisture_ave")
num_funct_treatment("moisture_ave")
```

# Check sociodemographic, health, exposure values
# Among those who missed a session/dropped and those who completed all sessions
```{r}
char_funct_4visits <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  rename(new_var = var) %>% 
  arrange(adult_id_char, winter_id, sampling_visit) %>% 
  group_by(adult_id_char) %>% 
  mutate(participant_n = n()) %>% 
  ungroup() %>% 
  filter(participant_n == 4) %>% 
  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
  distinct(adult_id_char, .keep_all = TRUE) %>% 
  mutate(total_n = n()) %>% 
  group_by(new_var) %>% 
  mutate(n = n(),
         percent = n()/total_n) %>% 
  distinct(new_var, .keep_all = TRUE) %>% 
  dplyr::select(new_var, n, percent) %>% 
  arrange(new_var)
demographics_summary
}

char_funct_miss_winter2 <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  rename(new_var = var) %>% 
  arrange(adult_id_char, winter_id, sampling_visit) %>% 
  group_by(adult_id_char) %>% 
  mutate(winter_id = as.numeric(winter_id)) %>% 
  mutate(winter_check = if_else(sum(winter_id)<3, "missed2", "present2")) %>% 
  ungroup() %>% 
  filter(winter_check == "missed2") %>% 
  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
  distinct(adult_id_char, .keep_all = TRUE) %>% 
  mutate(total_n = n()) %>% 
  group_by(new_var) %>% 
  mutate(n = n(),
         percent = n()/total_n) %>% 
  distinct(new_var, .keep_all = TRUE) %>% 
  dplyr::select(new_var, n, percent) %>% 
  arrange(new_var)
demographics_summary
}

char_funct_miss_visit <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  rename(new_var = var) %>% 
  arrange(adult_id_char, winter_id, sampling_visit) %>% 
  group_by(adult_id_char) %>% 
  mutate(participant_n = n()) %>% 
  ungroup() %>% 
  filter(participant_n < 4) %>% 
  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
  distinct(adult_id_char, .keep_all = TRUE) %>% 
  mutate(total_n = n()) %>% 
  group_by(new_var) %>% 
  mutate(n = n(),
         percent = n()/total_n) %>% 
  distinct(new_var, .keep_all = TRUE) %>% 
  dplyr::select(new_var, n, percent) %>% 
  arrange(new_var)
demographics_summary
}

# distinct by participant
char_funct_4visits("gender")
char_funct_miss_winter2("gender")
char_funct_miss_visit("gender")
char_funct_4visits("education")
char_funct_miss_winter2("education")
char_funct_miss_visit("education")
char_funct_4visits("income_3level")
char_funct_miss_winter2("income_3level")
char_funct_miss_visit("income_3level")
char_funct_4visits("home_floors_2level")
char_funct_miss_winter2("home_floors_2level")
char_funct_miss_visit("home_floors_2level")
char_funct_4visits("home_pets_3level")
char_funct_miss_winter2("home_pets_3level")
char_funct_miss_visit("home_pets_3level")
char_funct_4visits("residents_smoke")
char_funct_miss_winter2("residents_smoke")
char_funct_miss_visit("residents_smoke")
char_funct_4visits("stove_age")
char_funct_miss_winter2("stove_age")
char_funct_miss_visit("stove_age")
char_funct_4visits("chimney_clean")
char_funct_miss_winter2("chimney_clean")
char_funct_miss_visit("chimney_clean")
char_funct_4visits("wood_collect_method")
char_funct_miss_winter2("wood_collect_method")
char_funct_miss_visit("wood_collect_method")
char_funct_4visits("wood_collect")
char_funct_miss_winter2("wood_collect")
char_funct_miss_visit("wood_collect")
char_funct_4visits("stove_grade")
char_funct_miss_winter2("stove_grade")
char_funct_miss_visit("stove_grade")
char_funct_4visits("burn_level")
char_funct_miss_winter2("burn_level")
char_funct_miss_visit("burn_level")
```

# Check sociodemographic, health, exposure values
# Among those who missed a session/dropped and those who completed all sessions
```{r}
num_funct_4visits <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  rename(new_var = var) %>% 
  mutate(new_var = as.numeric(new_var)) %>% 
  arrange(adult_id_char, winter_id, sampling_visit) %>% 
  group_by(adult_id_char) %>% 
  mutate(participant_n = n()) %>% 
  ungroup() %>% 
  filter(participant_n == 4) %>% 
  filter(winter_id == 1) %>% 
  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
  distinct(adult_id_char, .keep_all = TRUE) %>% 
  filter(!is.na(new_var)) %>% 
  dplyr::select(new_var) %>% 
  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
demographics_summary
}

num_funct_miss_winter2 <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  rename(new_var = var) %>% 
  mutate(new_var = as.numeric(new_var)) %>% 
  arrange(adult_id_char, winter_id, sampling_visit) %>% 
  group_by(adult_id_char) %>% 
  mutate(winter_id = as.numeric(winter_id)) %>% 
  mutate(winter_check = if_else(sum(winter_id)<3, "missed2", "present2")) %>% 
  ungroup() %>% 
  filter(winter_check == "missed2") %>% 
  filter(winter_id == 1) %>% 
  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
  distinct(adult_id_char, .keep_all = TRUE) %>% 
  filter(!is.na(new_var)) %>% 
  dplyr::select(new_var) %>% 
  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
demographics_summary
}

num_funct_miss_visit <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  rename(new_var = var) %>% 
  mutate(new_var = as.numeric(new_var)) %>% 
  arrange(adult_id_char, winter_id, sampling_visit) %>% 
  group_by(adult_id_char) %>% 
  mutate(participant_n = n()) %>% 
  ungroup() %>% 
  filter(participant_n < 4) %>% 
  filter(winter_id == 1) %>% 
  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
  distinct(adult_id_char, .keep_all = TRUE) %>% 
  filter(!is.na(new_var)) %>% 
  dplyr::select(new_var) %>% 
  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
demographics_summary
}

num_funct_4visits("age")
num_funct_miss_winter2("age")
num_funct_miss_visit("age")
num_funct_4visits("bp_sys")
num_funct_miss_winter2("bp_sys")
num_funct_miss_visit("bp_sys")
num_funct_4visits("bp_dia")
num_funct_miss_winter2("bp_dia")
num_funct_miss_visit("bp_dia")
num_funct_4visits("bmi")
num_funct_miss_winter2("bmi")
num_funct_miss_visit("bmi")
num_funct_4visits("fvc")
num_funct_miss_winter2("fvc")
num_funct_miss_visit("fvc")
num_funct_4visits("fev1")
num_funct_miss_winter2("fev1")
num_funct_miss_visit("fev1")
num_funct_4visits("fev1_fvc_ratio")
num_funct_miss_winter2("fev1_fvc_ratio")
num_funct_miss_visit("fev1_fvc_ratio")
num_funct_4visits("fvc_pp")
num_funct_miss_winter2("fvc_pp")
num_funct_miss_visit("fvc_pp")
num_funct_4visits("fev1_pp")
num_funct_miss_winter2("fev1_pp")
num_funct_miss_visit("fev1_pp")
num_funct_4visits("pm_mean_winter")
num_funct_miss_winter2("pm_mean_winter")
num_funct_miss_visit("pm_mean_winter")
```

```{r}
kw_filtertype_data <- elders_cleaned_dataset_by_visit %>% 
  select(area, home, adult_id_char, winter_id, sampling_visit, treatment,
         filter_type, kw_perc_expected) %>% 
  mutate(treatment_filter_type = as.character(treatment),
         filter_type = as.character(filter_type),
         treatment_filter_type2 = if_else(treatment_filter_type == "Filter", filter_type,
                                         treatment_filter_type),
         treatment_filter_type = as.factor(treatment_filter_type2),
         filter_type = as.factor(filter_type)) %>% 
  select(-treatment_filter_type2) 
  group_by(treatment) %>% 
  summarize(mean(kw_perc_expected, na.rm = T), median(kw_perc_expected, na.rm = T))
kw_filtertype_data

summary(kw_filtertype_data)

compliance_summary <- elders_cleaned_dataset_by_visit %>% 
  filter(winter_id == 2) %>% 
  distinct(area, home, .keep_all = T) %>% 
  filter(treatment == "Filter") %>% 
  filter(!is.na(kw_perc_expected)) %>% 
  summarize("Mean KW % Exp" = mean(kw_perc_expected),
            "SD KW % Exp" = sd(kw_perc_expected),
            "N KW % Exp" = n(),
            "Min KW % Exp" = min(kw_perc_expected),
            "25Q KW % Exp" = quantile(kw_perc_expected, 0.25),
            "Med KW % Exp" = median(kw_perc_expected),
            "75Q KW % Exp" = quantile(kw_perc_expected, 0.75),
            "Max KW % Exp" = max(kw_perc_expected))
compliance_summary
```

```{r}
pm_cutoff_table <- elders_cleaned_dataset_by_visit %>% 
  filter(!is.na(pm_mean_winter)) %>% 
  distinct(area, home, winter_id, .keep_all = T) %>% 
  mutate(epa_12 = if_else(pm_mean_winter > 12, "Above", "Below"),
         epa_35 = if_else(pm_mean_winter > 35, "Above", "Below")) %>% 
  group_by(winter_id, treatment, epa_35) %>% 
  count()
pm_cutoff_table
```