### 2021-07-16 sums_data_cleaning.Rmd

---
title: "EldersAIR initial SUMs data work"
author: "Ethan Walker"
date: "Started 16 July 2021, Updated 16 July 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
```

# Add linked IDs to NPT SUMs data logs
# Use these throughout the data cleaning process to check start times
```{r}
# Input file from Box: ibutton.xlsx
npt_sums_data_log <- read_xlsx(paste0(file_path, "Input/NPT/ibutton.xlsx"), 
                              na = c("NULL", "")) %>%
  mutate(home_winter_id = as.character(HomeWinterID),
         ibid = iButtonID,
         date_log = InstallDate,
         time_log = InstallTime,
         logtag_location = Location) %>% 
  separate(time_log, into = c("trash", "time_log"), sep = " ") %>% 
  select(home_winter_id, ibid, date_log, time_log, logtag_location)

elders_linked_ids <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
  filter(area == "NPT") %>% 
  select(home_winter_id, winter_id, home, home_id_num)

npt_sums_log <- npt_sums_data_log %>% 
  right_join(elders_linked_ids, by = "home_winter_id") %>% 
  select(home_winter_id, winter_id, home, home_id_num, date_log, time_log, 
         ibid, logtag_location) %>% 
  arrange(home, date_log) %>% 
  mutate(date_log = ymd(date_log))

write_rds(npt_sums_log, paste0(file_path, "Output/npt_sums_data_log.rds"))

# Input file from Box: ibuttonupd.xlsx
npt_sums_data_log_updated <- read_xlsx(paste0(file_path, "Input/NPT/ibuttonupd.xlsx"), 
                                      na = c("NULL", "")) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         date = ChangeDate,
         time = ChangeTime,
         alarms = Alarms,
         file_name = ibutFile,
         sampling_visit = SamplingVisit) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  select(home_winter_id, date, time, alarms, file_name, 
         sampling_visit, QAQC, QAQCComments)

npt_sums_log <- npt_sums_data_log_updated %>% 
  right_join(elders_linked_ids, by = "home_winter_id") %>% 
  select(home_winter_id, home, home_id_num, date, time, alarms, file_name, 
         sampling_visit, QAQC, QAQCComments) %>% 
  arrange(home, date) %>% 
  mutate(date = ymd(date))
```

# Add linked IDs to NN SUMs data logs
# Use these throughout the data cleaning process to check start times
```{r}
# Input file from Box: ibutton.xlsx
nn_sums_data_log <- read_xlsx(paste0(file_path, "Input/NN/ibutton.xlsx"), 
                              na = c("NULL", "")) %>%
  mutate(home_winter_id = as.character(HomeWinterID),
         ibid = iButtonID,
         date_log = InstallDate,
         time_log = InstallTime,
         logtag_location = Location) %>% 
  separate(time_log, into = c("trash", "time_log"), sep = " ") %>% 
  select(home_winter_id, ibid, date_log, time_log, logtag_location)

elders_linked_ids <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
  filter(area == "NN") %>% 
  select(home_winter_id, winter_id, home, home_id_num)

nn_sums_log <- nn_sums_data_log %>% 
  right_join(elders_linked_ids, by = "home_winter_id") %>% 
  select(home_winter_id, winter_id, home, home_id_num, date_log, time_log, 
         ibid, logtag_location) %>% 
  arrange(home, date_log) %>% 
  mutate(date_log = ymd(date_log))

write_rds(nn_sums_log, paste0(file_path, "Output/nn_sums_data_log.rds"))

# Input file from Box: ibuttonupd.xlsx
nn_sums_data_log_updated <- read_xlsx(paste0(file_path, "Input/NN/ibuttonupd.xlsx"), 
                                      na = c("NULL", "")) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         date = ChangeDate,
         time = ChangeTime,
         alarms = Alarms,
         file_name = ibutFile,
         sampling_visit = SamplingVisit) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  select(home_winter_id, date, time, alarms, file_name, 
         sampling_visit, QAQC, QAQCComments)

nn_sums_log <- nn_sums_data_log_updated %>% 
  right_join(elders_linked_ids, by = "home_winter_id") %>% 
  select(home_winter_id, home, home_id_num, date, time, alarms, file_name, 
         sampling_visit, QAQC, QAQCComments) %>% 
  arrange(home, date) %>% 
  mutate(date = ymd(date))
```

# Pull in files with several different formats and combine
```{r}
file_path2 <- c("C:/Users/ethan.walker/Box/NoonanGroupData/EldersAIR/Ethan/Logtag Data/Clean data/")

###### NPT Files #######

#### File type: header_1col_date_m.d.y_time_ampm_c

list_files <- list.files(paste0(file_path2, "Nez Perce Tribe/header_1col_date_m.d.y_time_ampm_c"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "Nez Perce Tribe/header_1col_date_m.d.y_time_ampm_c"))
initial_data <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, skip = 1,
                            col_names = c("data"),
                            col_types = c("text"))) %>%
  unnest(save_data) %>% 
  separate(data, into = c("datetime", "c_f", "temp_c"), sep = ",") %>% 
  mutate(datetime = mdy_hms(datetime),
         temp_c = as.numeric(temp_c))

#### File type: header_1col_index_date_m.d.y_time_ampm_f

list_files <- list.files(paste0(file_path2, "Nez Perce Tribe/header_1col_index_date_m.d.y_time_ampm_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "Nez Perce Tribe/header_1col_index_date_m.d.y_time_ampm_f"))
initial_data <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_csv, skip = 1,
                            col_names = c("data"),
                            col_types = list(col_character()))) %>%
  unnest(save_data) %>% 
  mutate(data = gsub('"', "", data)) %>% 
  separate(data, into = c("trash", "date", "time", "temp_f"), sep = ",") %>% 
  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
  mutate(datetime = mdy_hms(datetime),
         temp_f = as.numeric(temp_f),
         temp_c = (temp_f - 32) * (5/9)) %>% 
  select(-trash, -temp_f)

#### File type: header_4col_date_d.m.y_time24h_c

list_files <- list.files(paste0(file_path2, "Nez Perce Tribe/header_4col_date_d.m.y_time24h_c"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "Nez Perce Tribe/header_4col_date_d.m.y_time24h_c"))
initial_data <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel)) %>%
  unnest(save_data) %>% 
  mutate(date = Date,
         time = `Time (24hr UTC-07:00)`,
         temp_c = `Temperature (°C)`) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_c = as.numeric(temp_c)) %>% 
  select(file_name, datetime, temp_c)

#### File type: header_5col_date_d.m.y_time_ampm_c_f

list_files <- list.files(paste0(file_path2, "Nez Perce Tribe/header_5col_date_d.m.y_time_ampm_c_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "Nez Perce Tribe/header_5col_date_d.m.y_time_ampm_c_f"))
initial_data <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel)) %>%
  unnest(save_data) %>% 
  mutate(date = ...1,
         time = ...2,
         ampm = ...3,
         temp_c = C) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  unite(c("time", "ampm"), col = "time", sep = " ") %>% 
  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_c = as.numeric(temp_c)) %>% 
  select(file_name, datetime, temp_c)

#### File type: header_5col_index_date_d.m.y_time_ampm_f

list_files <- list.files(paste0(file_path2, "Nez Perce Tribe/header_5col_index_date_d.m.y_time_ampm_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "Nez Perce Tribe/header_5col_index_date_d.m.y_time_ampm_f"))
initial_data <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel,
                            col_names = c("index", "date", "time", "temp_f", "type"),
                            col_types = c("date", "date", "text", "numeric", "text"))) %>%
  unnest(save_data)

practice_data <- initial_data %>% 
  mutate(date = ...1,
         time = ...2,
         ampm = ...3,
         temp_c = C) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  unite(c("time", "ampm"), col = "time", sep = " ") %>% 
  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_c = as.numeric(temp_c)) %>% 
  select(file_name, datetime, temp_c)
```