### 2021-07-20 sums_data_cleaning.Rmd
new logtag cleaning and code to work on CPHR server

---
title: "EldersAIR initial SUMs data work"
author: "Ethan Walker"
date: "Started 16 July 2021, Updated 16 July 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
```

# Add linked IDs to NPT SUMs data logs
# Use these throughout the data cleaning process to check start times
```{r}
# Input file from Box: ibutton.xlsx
npt_sums_data_log <- read_xlsx(paste0(file_path, "Input/NPT/ibutton.xlsx"), 
                              na = c("NULL", "")) %>%
  mutate(home_winter_id = as.character(HomeWinterID),
         ibid = iButtonID,
         date_log = InstallDate,
         time_log = InstallTime,
         logtag_location = Location) %>% 
  separate(time_log, into = c("trash", "time_log"), sep = " ") %>% 
  select(home_winter_id, ibid, date_log, time_log, logtag_location)

elders_linked_ids <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
  filter(area == "NPT") %>% 
  select(area, home_winter_id, winter_id, home, home_id_num)

npt_sums_log <- npt_sums_data_log %>% 
  right_join(elders_linked_ids, by = "home_winter_id") %>% 
  select(area, home_winter_id, winter_id, home, home_id_num, date_log, time_log, 
         ibid, logtag_location) %>% 
  arrange(home, date_log) %>% 
  mutate(date_log = ymd(date_log))

write_rds(npt_sums_log, paste0(file_path, "Output/npt_sums_data_log.rds"))

# Input file from Box: ibuttonupd.xlsx
npt_sums_data_log_updated <- read_xlsx(paste0(file_path, "Input/NPT/ibuttonupd.xlsx"), 
                                      na = c("NULL", "")) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         date = ChangeDate,
         time = ChangeTime,
         alarms = Alarms,
         file_name = ibutFile,
         sampling_visit = SamplingVisit) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  select(home_winter_id, date, time, alarms, file_name, 
         sampling_visit, QAQC, QAQCComments)

npt_sums_log2 <- npt_sums_data_log_updated %>% 
  right_join(elders_linked_ids, by = "home_winter_id") %>% 
  select(area, home_winter_id, home, home_id_num, date, time, alarms, file_name, 
         sampling_visit, QAQC, QAQCComments) %>% 
  arrange(home, date) %>% 
  mutate(date = ymd(date))
```

# Add linked IDs to NN SUMs data logs
# Use these throughout the data cleaning process to check start times
```{r}
# Input file from Box: ibutton.xlsx
nn_sums_data_log <- read_xlsx(paste0(file_path, "Input/NN/ibutton.xlsx"), 
                              na = c("NULL", "")) %>%
  mutate(home_winter_id = as.character(HomeWinterID),
         ibid = iButtonID,
         date_log = InstallDate,
         time_log = InstallTime,
         logtag_location = Location) %>% 
  separate(time_log, into = c("trash", "time_log"), sep = " ") %>% 
  select(home_winter_id, ibid, date_log, time_log, logtag_location)

elders_linked_ids <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
  filter(area == "NN") %>% 
  select(area, home_winter_id, winter_id, home, home_id_num)

nn_sums_log <- nn_sums_data_log %>% 
  right_join(elders_linked_ids, by = "home_winter_id") %>% 
  select(area, home_winter_id, winter_id, home, home_id_num, date_log, time_log, 
         ibid, logtag_location) %>% 
  arrange(home, date_log) %>% 
  mutate(date_log = ymd(date_log))

write_rds(nn_sums_log, paste0(file_path, "Output/nn_sums_data_log.rds"))

# Input file from Box: ibuttonupd.xlsx
nn_sums_data_log_updated <- read_xlsx(paste0(file_path, "Input/NN/ibuttonupd.xlsx"), 
                                      na = c("NULL", "")) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         date = ChangeDate,
         time = ChangeTime,
         alarms = Alarms,
         file_name = ibutFile,
         sampling_visit = SamplingVisit) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  select(home_winter_id, date, time, alarms, file_name, 
         sampling_visit, QAQC, QAQCComments)

nn_sums_log2 <- nn_sums_data_log_updated %>% 
  right_join(elders_linked_ids, by = "home_winter_id") %>% 
  select(area, home_winter_id, home, home_id_num, date, time, alarms, file_name, 
         sampling_visit, QAQC, QAQCComments) %>% 
  arrange(home, date) %>% 
  mutate(date = ymd(date))
```

# Open SSH server connection to work with files below
```{r}
library(ssh)

# Connect to VPN or work from campus

session <- ssh_connect("cphrstor1.health.umt.edu")
# Passward in email from Allen
# Need phone for Duo Authentication

print(session)

ssh_tunnel(session, port = 9090, target = "cphrstor1.health.umt.edu:22")

ssh_disconnect(session = session)

# ask for current working directory
ssh_exec_wait(session, command = 'pwd')

# make new directory, then use command line to transfer files
ssh_exec_wait(session, command = 'mkdir /home/ethan.walker/eldersair')
ssh_exec_wait(session, command = 'mkdir /home/ethan.walker/eldersair/logtag')

###### Command line code for sftp file transfer ######

# sftp ethan.walker@cphrstor1.health.umt.edu # open sftp connection
# pwd # print working directory
# cd /home/ethan.walker/eldersair/logtag # change directory - specify filepath
# put C:/R/EldersAIR/logtag_cleaning.r # transfer specific file
# put -r C:/R/EldersAIR/Logtag # use this to transfer a directory
# ls # list files in a directory
```

# Pull in files with several different formats and combine
```{r}
# Save SSH working directory from above
file_path2 <- c("C:/R/EldersAIR/Logtag/")

###### NPT Files #######

#### File type: header_1col_date_m.d.y_time_ampm_c

list_files <- list.files(paste0(file_path2, "NPT/header_1col_date_m.d.y_time_ampm_c"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/header_1col_date_m.d.y_time_ampm_c"))
initial_data1 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, skip = 1,
                            col_names = c("data"),
                            col_types = c("text"))) %>%
  unnest(save_data) %>% 
  separate(data, into = c("datetime", "c_f", "temp_c"), sep = ",") %>% 
  mutate(datetime = mdy_hms(datetime),
         temp_c = as.numeric(temp_c),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: header_1col_index_date_m.d.y_time_ampm_f

list_files <- list.files(paste0(file_path2, "NPT/header_1col_index_date_m.d.y_time_ampm_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/header_1col_index_date_m.d.y_time_ampm_f"))
initial_data2 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_csv, skip = 1,
                            col_names = c("data"),
                            col_types = list(col_character()))) %>%
  unnest(save_data) %>% 
  mutate(data = gsub('"', "", data)) %>% 
  separate(data, into = c("trash", "date", "time", "temp_f"), sep = ",") %>% 
  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
  mutate(datetime = mdy_hms(datetime),
         temp_f = as.numeric(temp_f),
         temp_c = (temp_f - 32) * (5/9),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: header_4col_date_d.m.y_time24h_c

list_files <- list.files(paste0(file_path2, "NPT/header_4col_date_d.m.y_time24h_c"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/header_4col_date_d.m.y_time24h_c"))
initial_data3 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel)) %>%
  unnest(save_data) %>% 
  mutate(date = Date,
         time = `Time (24hr UTC-07:00)`,
         time2 = `Time (24hr UTC-08:00)`,
         time = if_else(is.na(time), time2, time),
         temp_c = `Temperature (°C)`) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_c = as.numeric(temp_c),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: header_5col_date_d.m.y_time_ampm_c_f

list_files <- list.files(paste0(file_path2, "NPT/header_5col_date_d.m.y_time_ampm_c_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/header_5col_date_d.m.y_time_ampm_c_f"))
initial_data4 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel)) %>%
  unnest(save_data) %>% 
  mutate(date = ...1,
         time = ...2,
         ampm = ...3,
         temp_c = C) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  unite(c("time", "ampm"), col = "time", sep = " ") %>% 
  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_c = as.numeric(temp_c),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: header_5col_index_date_d.m.y_time_ampm_f

list_files <- list.files(paste0(file_path2, "NPT/header_5col_index_date_d.m.y_time_ampm_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/header_5col_index_date_d.m.y_time_ampm_f"))
initial_data5 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, skip = 1, range = "B2:D14000",
                            col_names = c("date", "time", "temp_f"),
                            col_types = c("date", "date", "numeric"))) %>%
  unnest(save_data) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_c = (temp_f - 32) * (5/9),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: header22_2col_date_d.m.ytime_c

list_files <- list.files(paste0(file_path2, "NPT/header22_2col_date_d.m.ytime_c"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/header22_2col_date_d.m.ytime_c"))
initial_data6 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, skip = 22,
                            col_names = c("datetime", "temp_c"),
                            col_types = c("date", "numeric"))) %>%
  unnest(save_data) %>% 
  mutate(datetime = ymd_hms(datetime),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: noheader_1col_am_pm_f

list_files <- list.files(paste0(file_path2, "NPT/noheader_1col_am_pm_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/noheader_1col_am_pm_f"))
initial_data7 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, 
                            col_names = c("data"),
                            col_types = c("text"))) %>%
  unnest(save_data) %>% 
  separate(data, into = c("datetime", "c_f", "temp_f"), sep = ",") %>% 
  mutate(datetime = mdy_hms(datetime),
         temp_f = as.numeric(temp_f),
         temp_c = (temp_f - 32) * (5/9),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: noheader_2col_date_d.m.y_t

list_files <- list.files(paste0(file_path2, "NPT/noheader_2col_date_d.m.y_t"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/noheader_2col_date_d.m.y_t"))
initial_data8 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, 
                            col_names = c("datetime", "temp_c"),
                            col_types = c("date", "numeric"))) %>%
  unnest(save_data) %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_c = as.numeric(temp_c),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: noheader_3col_date_d.m.y_time_f

list_files <- list.files(paste0(file_path2, "NPT/noheader_3col_date_d.m.y_time_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/noheader_3col_date_d.m.y_time_f"))
initial_data9 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, 
                            col_names = c("datetime", "c_f", "temp_f"),
                            col_types = c("date", "text", "numeric"))) %>%
  unnest(save_data) %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_f = as.numeric(temp_f),
         temp_c = (temp_f - 32) * (5/9),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: noheader_4col_date_d.m.y_time_ampm_f

list_files <- list.files(paste0(file_path2, "NPT/noheader_4col_date_d.m.y_time_ampm_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/noheader_4col_date_d.m.y_time_ampm_f"))
initial_data10 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, 
                            col_names = c("date", "time", "c_f", "temp_f"),
                            col_types = c("date", "date", "text", "numeric"))) %>%
  unnest(save_data) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_f = as.numeric(temp_f),
         temp_c = (temp_f - 32) * (5/9),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: noheader_4col_date_d.m.y_time_f

list_files <- list.files(paste0(file_path2, "NPT/noheader_4col_date_d.m.y_time_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/noheader_4col_date_d.m.y_time_f"))
initial_data11 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, 
                            col_names = c("date", "time", "c_f", "temp_f"),
                            col_types = c("date", "date", "text", "numeric"))) %>%
  unnest(save_data) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_f = as.numeric(temp_f),
         temp_c = (temp_f - 32) * (5/9),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: noheader_5col_date_d.m.y_time_ampm_f

list_files <- list.files(paste0(file_path2, "NPT/noheader_5col_date_d.m.y_time_ampm_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/noheader_5col_date_d.m.y_time_ampm_f"))
initial_data12 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, 
                            col_names = c("date", "time", "ampm", "c_f", "temp_f"),
                            col_types = c("date", "date", "text", "text", "numeric"))) %>%
  unnest(save_data) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  unite(c("date", "time", "ampm"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_f = as.numeric(temp_f),
         temp_c = (temp_f - 32) * (5/9),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: noheader_5col_date_d.m.y_time_ampm_t_t

list_files <- list.files(paste0(file_path2, "NPT/noheader_5col_date_d.m.y_time_ampm_t_t"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/noheader_5col_date_d.m.y_time_ampm_t_t"))
initial_data13 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, 
                            col_names = c("date", "time", "ampm", "temp_c", "temp_f"),
                            col_types = c("date", "date", "text", "numeric", "numeric"))) %>%
  unnest(save_data) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  unite(c("date", "time", "ampm"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_f = as.numeric(temp_f),
         temp_c = (temp_f - 32) * (5/9),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

###### NN Files #######

#### File type: noheader_1col_date_m.d.y_time_ampm_f

list_files <- list.files(paste0(file_path2, "NN/noheader_1col_date_m.d.y_time_ampm_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NN/noheader_1col_date_m.d.y_time_ampm_f"))
initial_data14 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, 
                            col_names = c("data"),
                            col_types = c("text"))) %>%
  unnest(save_data) %>% 
  separate(data, into = c("datetime", "c_f", "temp_f"), sep = ",") %>% 
  mutate(datetime = mdy_hms(datetime),
         temp_f = as.numeric(temp_f),
         temp_c = (temp_f - 32) * (5/9),
         area = "NN") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: noheader_1col_date_m.d.y_time_ampm_c

list_files <- list.files(paste0(file_path2, "NN/noheader_1col_date_m.d.y_time_ampm_c"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NN/noheader_1col_date_m.d.y_time_ampm_c"))
initial_data15 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, 
                            col_names = c("data"),
                            col_types = c("text"))) %>%
  unnest(save_data) %>% 
  separate(data, into = c("datetime", "c_f", "temp_c"), sep = ",") %>% 
  mutate(datetime = mdy_hms(datetime),
         temp_c = as.numeric(temp_c),
         area = "NN") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: noheader_4col_index_date_y.m.d_time24h_t

list_files <- list.files(paste0(file_path2, "NN/noheader_4col_index_date_y.m.d_time24h_t"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NN/noheader_4col_index_date_y.m.d_time24h_t"))
initial_data16 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_csv,
                            col_names = c("index", "date", "time", "temp_c"),
                            col_types = list(col_character(),
                                             col_character(),
                                             col_character(),
                                             col_double()))) %>%
  unnest(save_data) %>% 
  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_c = as.numeric(temp_c),
         area = "NN") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: noheader_5col_date_d.m.y_time_ampm_f

list_files <- list.files(paste0(file_path2, "NN/noheader_5col_date_d.m.y_time_ampm_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NN/noheader_5col_date_d.m.y_time_ampm_f"))
initial_data17 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, 
                            col_names = c("date", "time", "ampm", "c_f", "temp_f"),
                            col_types = c("date", "date", "text", "text", "numeric"))) %>%
  unnest(save_data) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  unite(c("date", "time", "ampm"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_f = as.numeric(temp_f),
         temp_c = (temp_f - 32) * (5/9),
         area = "NN") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: noheader_5col_date_d.m.y_time_ampm_c

list_files <- list.files(paste0(file_path2, "NN/noheader_5col_date_d.m.y_time_ampm_c"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NN/noheader_5col_date_d.m.y_time_ampm_c"))
initial_data18 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, 
                            col_names = c("date", "time", "ampm", "c_f", "temp_c"),
                            col_types = c("date", "date", "text", "text", "numeric"))) %>%
  unnest(save_data) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  unite(c("date", "time", "ampm"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_c = as.numeric(temp_c),
         area = "NN") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))
```

```{r}
# Bind all data files together
all_sums_raw <- rbind(initial_data1, initial_data2, initial_data3, initial_data4,
                      initial_data5, initial_data6, initial_data7, initial_data8,
                      initial_data9, initial_data10, initial_data11, initial_data12,
                      initial_data13, initial_data14, initial_data15, initial_data16,
                      initial_data17, initial_data18) %>% 
  mutate(area = as.factor(area),
         file_name = as.factor(file_name)) %>% 
  rename(datetime_sums = datetime,
         temp_c_sums = temp_c) %>% 
  filter(!is.na(datetime_sums))

summary(all_sums_raw)

write_rds(all_sums_raw, paste0(file_path, "Output/all_sums_raw.rds"))
```

```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

all_sums_raw <- read_rds(paste0(file_path, "Output/all_sums_raw.rds")) %>% 
  ungroup() %>% 
  arrange(area, file_name, datetime_sums)

# Next steps: identify start/stop times for logtag data, filter, then clean up and create variables
```