### 2021-09-07 sums_data_cleaning.Rmd

---
title: "EldersAIR initial SUMs data work"
author: "Ethan Walker"
date: "Started 16 July 2021, Updated 21 July 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)

jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
```

# Add linked IDs to NPT SUMs data logs
# Use these throughout the data cleaning process to check start times
```{r}
# Input file from Box: ibutton.xlsx
npt_sums_data_log <- read_xlsx(paste0(file_path, "Input/NPT/ibutton.xlsx"), 
                              na = c("NULL", "")) %>%
  mutate(home_winter_id = as.character(HomeWinterID),
         ibid = iButtonID,
         date_log = InstallDate,
         time_log = InstallTime,
         logtag_location = Location) %>% 
  separate(time_log, into = c("trash", "time_log"), sep = " ") %>% 
  select(home_winter_id, ibid, date_log, time_log, logtag_location)

elders_linked_ids <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
  filter(area == "NPT") %>% 
  select(area, home_winter_id, winter_id, home, home_id_num)

npt_sums_log <- npt_sums_data_log %>% 
  right_join(elders_linked_ids, by = "home_winter_id") %>% 
  select(area, home_winter_id, winter_id, home, home_id_num, date_log, time_log, 
         ibid, logtag_location) %>% 
  arrange(home, date_log) %>% 
  mutate(date_log = ymd(date_log))

write_rds(npt_sums_log, paste0(file_path, "Output/npt_sums_data_log.rds"))

# Input file from Box: ibuttonupd.xlsx
npt_sums_data_log_updated <- read_xlsx(paste0(file_path, "Input/NPT/ibuttonupd.xlsx"), 
                                      na = c("NULL", "")) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         date = ChangeDate,
         time = ChangeTime,
         alarms = Alarms,
         file_name = ibutFile,
         sampling_visit = SamplingVisit) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  select(home_winter_id, date, time, alarms, file_name, 
         sampling_visit, QAQC, QAQCComments)

npt_sums_log2 <- npt_sums_data_log_updated %>% 
  right_join(elders_linked_ids, by = "home_winter_id") %>% 
  select(area, home_winter_id, home, home_id_num, date, time, alarms, file_name, 
         sampling_visit, QAQC, QAQCComments) %>% 
  arrange(home, date) %>% 
  mutate(date = ymd(date))
```

# Add linked IDs to NN SUMs data logs
# Use these throughout the data cleaning process to check start times
```{r}
# Input file from Box: ibutton.xlsx
nn_sums_data_log <- read_xlsx(paste0(file_path, "Input/NN/ibutton.xlsx"), 
                              na = c("NULL", "")) %>%
  mutate(home_winter_id = as.character(HomeWinterID),
         ibid = iButtonID,
         date_log = InstallDate,
         time_log = InstallTime,
         logtag_location = Location) %>% 
  separate(time_log, into = c("trash", "time_log"), sep = " ") %>% 
  select(home_winter_id, ibid, date_log, time_log, logtag_location)

elders_linked_ids <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
  filter(area == "NN") %>% 
  select(area, home_winter_id, winter_id, home, home_id_num)

nn_sums_log <- nn_sums_data_log %>% 
  right_join(elders_linked_ids, by = "home_winter_id") %>% 
  select(area, home_winter_id, winter_id, home, home_id_num, date_log, time_log, 
         ibid, logtag_location) %>% 
  arrange(home, date_log) %>% 
  mutate(date_log = ymd(date_log))

write_rds(nn_sums_log, paste0(file_path, "Output/nn_sums_data_log.rds"))

# Input file from Box: ibuttonupd.xlsx
nn_sums_data_log_updated <- read_xlsx(paste0(file_path, "Input/NN/ibuttonupd.xlsx"), 
                                      na = c("NULL", "")) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         date = ChangeDate,
         time = ChangeTime,
         alarms = Alarms,
         file_name = ibutFile,
         sampling_visit = SamplingVisit) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  select(home_winter_id, date, time, alarms, file_name, 
         sampling_visit, QAQC, QAQCComments)

nn_sums_log2 <- nn_sums_data_log_updated %>% 
  right_join(elders_linked_ids, by = "home_winter_id") %>% 
  select(area, home_winter_id, home, home_id_num, date, time, alarms, file_name, 
         sampling_visit, QAQC, QAQCComments) %>% 
  arrange(home, date) %>% 
  mutate(date = ymd(date))
```

# Combine SUMs logs
```{r}
elders_sums_log <- rbind(npt_sums_log2, nn_sums_log2)

write_rds(elders_sums_log, paste0(file_path, "Output/elders_sums_log.rds"))
```

# Open SSH server connection to work with files below
```{r}
library(ssh)

# Connect to VPN or work from campus

session <- ssh_connect("cphrstor1.health.umt.edu")
# Passward in email from Allen
# Need phone for Duo Authentication

print(session)

ssh_tunnel(session, port = 9090, target = "cphrstor1.health.umt.edu:22")

ssh_disconnect(session = session)

# ask for current working directory
ssh_exec_wait(session, command = 'pwd')

# make new directory, then use command line to transfer files
ssh_exec_wait(session, command = 'mkdir /home/ethan.walker/eldersair')
ssh_exec_wait(session, command = 'mkdir /home/ethan.walker/eldersair/logtag')

###### Command line code for sftp file transfer ######

# sftp ethan.walker@cphrstor1.health.umt.edu # open sftp connection
# pwd # print working directory
# cd /home/ethan.walker/eldersair/logtag # change directory - specify filepath
# put C:/R/EldersAIR/logtag_cleaning.r # transfer specific file
# put -r C:/R/EldersAIR/Logtag # use this to transfer a directory
# ls # list files in a directory
```

# Pull in files with several different formats and combine
```{r}
# Save SSH working directory from above
file_path2 <- c("C:/R/EldersAIR/Logtag/")

###### NPT Files #######

#### File type: header_1col_date_m.d.y_time_ampm_c

list_files <- list.files(paste0(file_path2, "NPT/header_1col_date_m.d.y_time_ampm_c"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/header_1col_date_m.d.y_time_ampm_c"))
initial_data1 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, skip = 1,
                            col_names = c("data"),
                            col_types = c("text"))) %>%
  unnest(save_data) %>% 
  separate(data, into = c("datetime", "c_f", "temp_c"), sep = ",") %>% 
  mutate(datetime = mdy_hms(datetime),
         temp_c = as.numeric(temp_c),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: header_1col_index_date_m.d.y_time_ampm_f

list_files <- list.files(paste0(file_path2, "NPT/header_1col_index_date_m.d.y_time_ampm_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/header_1col_index_date_m.d.y_time_ampm_f"))
initial_data2 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_csv, skip = 1,
                            col_names = c("data"),
                            col_types = list(col_character()))) %>%
  unnest(save_data) %>% 
  mutate(data = gsub('"', "", data)) %>% 
  separate(data, into = c("trash", "date", "time", "temp_f"), sep = ",") %>% 
  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
  mutate(datetime = mdy_hms(datetime),
         temp_f = as.numeric(temp_f),
         temp_c = (temp_f - 32) * (5/9),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: header_4col_date_d.m.y_time24h_c

list_files <- list.files(paste0(file_path2, "NPT/header_4col_date_d.m.y_time24h_c"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/header_4col_date_d.m.y_time24h_c"))
initial_data3 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel)) %>%
  unnest(save_data) %>% 
  mutate(date = Date,
         time = `Time (24hr UTC-07:00)`,
         time2 = `Time (24hr UTC-08:00)`,
         time = if_else(is.na(time), time2, time),
         temp_c = `Temperature (°C)`) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_c = as.numeric(temp_c),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: header_5col_date_d.m.y_time_ampm_c_f

list_files <- list.files(paste0(file_path2, "NPT/header_5col_date_d.m.y_time_ampm_c_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/header_5col_date_d.m.y_time_ampm_c_f"))
initial_data4 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel)) %>%
  unnest(save_data) %>% 
  mutate(date = ...1,
         time = ...2,
         ampm = ...3,
         temp_c = C) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  unite(c("time", "ampm"), col = "time", sep = " ") %>% 
  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_c = as.numeric(temp_c),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: header_5col_index_date_d.m.y_time_ampm_f

list_files <- list.files(paste0(file_path2, "NPT/header_5col_index_date_d.m.y_time_ampm_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/header_5col_index_date_d.m.y_time_ampm_f"))
initial_data5 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, skip = 1, range = "B2:D14000",
                            col_names = c("date", "time", "temp_f"),
                            col_types = c("date", "date", "numeric"))) %>%
  unnest(save_data) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_c = (temp_f - 32) * (5/9),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: header22_2col_date_d.m.ytime_c

list_files <- list.files(paste0(file_path2, "NPT/header22_2col_date_d.m.ytime_c"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/header22_2col_date_d.m.ytime_c"))
initial_data6 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, skip = 22,
                            col_names = c("datetime", "temp_c"),
                            col_types = c("date", "numeric"))) %>%
  unnest(save_data) %>% 
  mutate(datetime = ymd_hms(datetime),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: noheader_1col_am_pm_f

list_files <- list.files(paste0(file_path2, "NPT/noheader_1col_am_pm_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/noheader_1col_am_pm_f"))
initial_data7 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, 
                            col_names = c("data"),
                            col_types = c("text"))) %>%
  unnest(save_data) %>% 
  separate(data, into = c("datetime", "c_f", "temp_f"), sep = ",") %>% 
  mutate(datetime = mdy_hms(datetime),
         temp_f = as.numeric(temp_f),
         temp_c = (temp_f - 32) * (5/9),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: noheader_2col_date_d.m.y_t

list_files <- list.files(paste0(file_path2, "NPT/noheader_2col_date_d.m.y_t"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/noheader_2col_date_d.m.y_t"))
initial_data8 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, 
                            col_names = c("datetime", "temp_c"),
                            col_types = c("date", "numeric"))) %>%
  unnest(save_data) %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_c = as.numeric(temp_c),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: noheader_3col_date_d.m.y_time_f

list_files <- list.files(paste0(file_path2, "NPT/noheader_3col_date_d.m.y_time_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/noheader_3col_date_d.m.y_time_f"))
initial_data9 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, 
                            col_names = c("datetime", "c_f", "temp_f"),
                            col_types = c("date", "text", "numeric"))) %>%
  unnest(save_data) %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_f = as.numeric(temp_f),
         temp_c = (temp_f - 32) * (5/9),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: noheader_4col_date_d.m.y_time_ampm_f

list_files <- list.files(paste0(file_path2, "NPT/noheader_4col_date_d.m.y_time_ampm_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/noheader_4col_date_d.m.y_time_ampm_f"))
initial_data10 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, 
                            col_names = c("date", "time", "c_f", "temp_f"),
                            col_types = c("date", "date", "text", "numeric"))) %>%
  unnest(save_data) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_f = as.numeric(temp_f),
         temp_c = (temp_f - 32) * (5/9),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: noheader_4col_date_d.m.y_time_f

list_files <- list.files(paste0(file_path2, "NPT/noheader_4col_date_d.m.y_time_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/noheader_4col_date_d.m.y_time_f"))
initial_data11 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, 
                            col_names = c("date", "time", "c_f", "temp_f"),
                            col_types = c("date", "date", "text", "numeric"))) %>%
  unnest(save_data) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_f = as.numeric(temp_f),
         temp_c = (temp_f - 32) * (5/9),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: noheader_5col_date_d.m.y_time_ampm_f

list_files <- list.files(paste0(file_path2, "NPT/noheader_5col_date_d.m.y_time_ampm_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/noheader_5col_date_d.m.y_time_ampm_f"))
initial_data12 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, 
                            col_names = c("date", "time", "ampm", "c_f", "temp_f"),
                            col_types = c("date", "date", "text", "text", "numeric"))) %>%
  unnest(save_data) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  unite(c("date", "time", "ampm"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_f = as.numeric(temp_f),
         temp_c = (temp_f - 32) * (5/9),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: noheader_5col_date_d.m.y_time_ampm_t_t

list_files <- list.files(paste0(file_path2, "NPT/noheader_5col_date_d.m.y_time_ampm_t_t"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NPT/noheader_5col_date_d.m.y_time_ampm_t_t"))
initial_data13 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, 
                            col_names = c("date", "time", "ampm", "temp_c", "temp_f"),
                            col_types = c("date", "date", "text", "numeric", "numeric"))) %>%
  unnest(save_data) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  unite(c("date", "time", "ampm"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_f = as.numeric(temp_f),
         temp_c = (temp_f - 32) * (5/9),
         area = "NPT") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

###### NN Files #######

#### File type: noheader_1col_date_m.d.y_time_ampm_f

list_files <- list.files(paste0(file_path2, "NN/noheader_1col_date_m.d.y_time_ampm_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NN/noheader_1col_date_m.d.y_time_ampm_f"))
initial_data14 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, 
                            col_names = c("data"),
                            col_types = c("text"))) %>%
  unnest(save_data) %>% 
  separate(data, into = c("datetime", "c_f", "temp_f"), sep = ",") %>% 
  mutate(datetime = mdy_hms(datetime),
         temp_f = as.numeric(temp_f),
         temp_c = (temp_f - 32) * (5/9),
         area = "NN") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: noheader_1col_date_m.d.y_time_ampm_c

list_files <- list.files(paste0(file_path2, "NN/noheader_1col_date_m.d.y_time_ampm_c"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NN/noheader_1col_date_m.d.y_time_ampm_c"))
initial_data15 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, 
                            col_names = c("data"),
                            col_types = c("text"))) %>%
  unnest(save_data) %>% 
  separate(data, into = c("datetime", "c_f", "temp_c"), sep = ",") %>% 
  mutate(datetime = mdy_hms(datetime),
         temp_c = as.numeric(temp_c),
         area = "NN") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: noheader_4col_index_date_y.m.d_time24h_t

list_files <- list.files(paste0(file_path2, "NN/noheader_4col_index_date_y.m.d_time24h_t"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NN/noheader_4col_index_date_y.m.d_time24h_t"))
initial_data16 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_csv,
                            col_names = c("index", "date", "time", "temp_c"),
                            col_types = list(col_character(),
                                             col_character(),
                                             col_character(),
                                             col_double()))) %>%
  unnest(save_data) %>% 
  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_c = as.numeric(temp_c),
         area = "NN") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: noheader_5col_date_d.m.y_time_ampm_f

list_files <- list.files(paste0(file_path2, "NN/noheader_5col_date_d.m.y_time_ampm_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NN/noheader_5col_date_d.m.y_time_ampm_f"))
initial_data17 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, 
                            col_names = c("date", "time", "ampm", "c_f", "temp_f"),
                            col_types = c("date", "date", "text", "text", "numeric"))) %>%
  unnest(save_data) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  unite(c("date", "time", "ampm"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_f = as.numeric(temp_f),
         temp_c = (temp_f - 32) * (5/9),
         area = "NN") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))

#### File type: noheader_5col_date_d.m.y_time_ampm_c

list_files <- list.files(paste0(file_path2, "NN/noheader_5col_date_d.m.y_time_ampm_c"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "NN/noheader_5col_date_d.m.y_time_ampm_c"))
initial_data18 <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, 
                            col_names = c("date", "time", "ampm", "c_f", "temp_c"),
                            col_types = c("date", "date", "text", "text", "numeric"))) %>%
  unnest(save_data) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  unite(c("date", "time", "ampm"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_c = as.numeric(temp_c),
         area = "NN") %>% 
  select(area, file_name, datetime, temp_c) %>% 
  filter(!is.na(temp_c))
```

```{r}
# Bind all data files together
all_sums_raw <- rbind(initial_data1, initial_data2, initial_data3, initial_data4,
                      initial_data5, initial_data6, initial_data7, initial_data8,
                      initial_data9, initial_data10, initial_data11, initial_data12,
                      initial_data13, initial_data14, initial_data15, initial_data16,
                      initial_data17, initial_data18) %>% 
  mutate(area = as.factor(area),
         file_name = as.factor(file_name)) %>% 
  rename(datetime_sums = datetime,
         temp_c_sums = temp_c) %>% 
  filter(!is.na(datetime_sums))

summary(all_sums_raw)

write_rds(all_sums_raw, paste0(file_path, "Output/all_sums_raw.rds"))
```

```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

all_sums_raw <- read_rds(paste0(file_path, "Output/all_sums_raw.rds")) %>% 
  ungroup() %>% 
  arrange(area, file_name, datetime_sums) %>% 
  mutate(file_name = gsub(".csv", "", file_name)) %>% 
  mutate(file_name = gsub(".xlsx", "", file_name))

elders_sums_log <- read_rds(paste0(file_path, "Output/elders_sums_log.rds"))  %>% 
  separate(file_name, c("trash", "file_name"), sep = "uploads/") %>% 
  mutate(file_name = gsub(".csv", "", file_name)) %>% 
  mutate(file_name = gsub(".ltd", "", file_name)) %>% 
  mutate(file_name = gsub(".xlsx", "", file_name))

elders_demographics <- read_rds(paste0(file_path, "Output/elders_demographics_full.rds")) %>% 
  unite(c("sampling_date", "sampling_time"), col = "start_datetime", sep = " ") %>% 
  unite(c("sampling_date2", "sampling_time2"), col = "start_datetime2", sep = " ") %>% 
  unite(c("pickup_date", "pickup_time"), col = "stop_datetime1", sep = " ") %>% 
  unite(c("pickup_date2", "pickup_time2"), col = "stop_datetime", sep = " ") %>% 
  mutate(start_datetime = ymd_hms(start_datetime),
         stop_datetime = ymd_hms(stop_datetime),
         start_datetime2 = ymd_hms(start_datetime2),
         stop_datetime1 = ymd_hms(stop_datetime1),
         stop_datetime = if_else(is.na(stop_datetime), stop_datetime1, stop_datetime)) %>% 
  select(area, winter_id, home_winter_id, start_datetime, stop_datetime) %>% 
  filter(!is.na(home_winter_id))

all_sums_clean <- all_sums_raw %>% 
  left_join(elders_sums_log, by = c("area", "file_name")) %>% 
  filter(!is.na(home_winter_id)) %>% 
  left_join(elders_demographics, by = c("area", "home_winter_id")) %>% 
  mutate(datetime_filter = if_else((datetime_sums >= start_datetime) &
                                   (datetime_sums <= stop_datetime), 1, 0)) %>% 
  filter(datetime_filter == 1) %>% 
  distinct(area, home_winter_id, datetime_sums, .keep_all = T) %>% 
  select(area, home_id_num, home, home_winter_id, winter_id, sampling_visit, 
         file_name, datetime_sums, temp_c_sums, QAQC:stop_datetime) %>% 
  arrange(area, home_winter_id, datetime_sums) %>% 
  group_by(area, home, winter_id) %>% 
  mutate(lag_filter = if_else(lag(file_name) != file_name, 1, 0),
         sum_filter = sum(lag_filter, na.rm = T)) %>% 
  filter(lag_filter != 1) %>% 
  select(-lag_filter, -sum_filter) %>% 
  group_by(area, home, winter_id) %>% 
  mutate(lag_time_diff = (datetime_sums - lag(datetime_sums)),
         temp_c_sums = if_else(temp_c_sums < -11, 9999, temp_c_sums),
         temp_c_sums = na_if(temp_c_sums, 9999),
         sums_winter_n = n(),
         sums_winter_mean = mean(temp_c_sums, na.rm = T)) %>% 
  group_by(area, home) %>% 
  mutate(sums_total_n = n(),
         sums_total_mean = mean(temp_c_sums, na.rm = T),
         lag_temp_diff = if_else(lag_time_diff < 26, lag(temp_c_sums), 9999),
         lag_temp_diff = na_if(lag_temp_diff, 9999)) %>% 
  ungroup() %>% 
  arrange(area, home_winter_id, datetime_sums) %>% 
  mutate(area = as.factor(area),
         home = as.factor(home),
         home_winter_id = as.factor(home_winter_id),
         sampling_visit = as.factor(sampling_visit),
         day_of_week = weekdays(datetime_sums),
         month_sums = as.factor(month(datetime_sums, label = TRUE, abbr = TRUE)))

sums_clean_distinct_winter <- all_sums_clean %>% 
  distinct(area, home_winter_id, .keep_all = T) %>% 
  select(area, home_winter_id, start_datetime, stop_datetime, sums_winter_n:sums_total_mean)

summary(all_sums_clean)
summary(sums_clean_distinct_winter)

write_rds(all_sums_clean, paste0(file_path, "Output/all_sums_clean.rds"))
write_rds(sums_clean_distinct_winter, paste0(file_path, "Output/sums_clean_distinct_winter.rds"))

write_csv(all_sums_clean, paste0(file_path, "Output/elders_logtag.csv"))
write_csv(sums_clean_distinct_winter, paste0(file_path, "Output/elders_logtag_distinct_winter.csv"))
```

# Plot temp over time for individual home to check for trends visually
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

sums_clean <- read_rds(paste0(file_path, "Output/all_sums_clean.rds"))
sums_n_check <- sums_clean %>% 
  filter(temp_c_sums < 0) %>% 
  count(area, home_winter_id)

sums_time_trends_plot <- sums_clean %>% 
  filter(area == "NN" & home_winter_id == "106") %>% 
  ggplot() + 
    geom_point(aes(datetime_sums, temp_c_sums, color = sampling_visit), size = 1.5) +
    theme_classic() +
    labs(y = "Temperature (degrees C)",
         x = "Datetime",
         color = "Sampling Visit") +
    theme(axis.title.y = element_text(size = 12,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_text(size = 12),
          axis.text.x = element_blank(),
          axis.text.y = element_text(size = 12, color = "black"),
          axis.line.x = element_line(colour = "black", size = 1), 
          axis.line.y = element_line(colour = "black", size = 1), 
          axis.ticks = element_blank()) +
    scale_color_manual(values = jv_palette)
sums_time_trends_plot

summary(sums_clean)
```

# Thresholds for heating events
```{r}
# Load SUMs data
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

sums_clean <- read_rds(paste0(file_path, "Output/all_sums_clean.rds"))

sums_events_30 <- sums_clean %>% 
  group_by(area, home, winter_id) %>%
  mutate(threshold_temp = if_else(temp_c_sums >= 30, 1, 0),
         threshold_temp_event = if_else(threshold_temp == 1 & lag(threshold_temp) == 0, 
                                        "start", "Temp < threshold"),
         threshold_temp_event = if_else(is.na(threshold_temp_event), "start", threshold_temp_event),
         threshold_temp_event = if_else(threshold_temp == 1 & threshold_temp_event == "Temp < threshold",
                                      "Temp >= threshold", threshold_temp_event),
         threshold_temp_event = if_else(threshold_temp_event == "Temp >= threshold" &
                                      lead(threshold_temp_event) == "Temp < threshold", 
                                      "event_end", threshold_temp_event)) %>% 
  filter(threshold_temp_event == "event_end" | threshold_temp_event == "start") %>% 
  separate(datetime_sums, c("date_sums", "time_sums"), sep = " ", remove = F) %>% 
  mutate(time_sums = as.character(time_sums),
         date_sums = as.character(date_sums)) %>%  
  spread(threshold_temp_event, datetime_sums) %>% 
  mutate(event_end = lead(event_end)) %>% 
  filter(!is.na(start)) %>% 
  mutate(hrs_till_next_event = (lead(start)-event_end),
         hrs_till_next_event = as.numeric(hrs_till_next_event)/3600,
         event_end2 = if_else(hrs_till_next_event <= 1.5, lead(event_end), event_end),
         event_end3 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5, 
                              lead(event_end, 2), event_end2),
         event_end4 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5
                              & lead(hrs_till_next_event, 2) <= 1.5, 
                              lead(event_end, 3), event_end3),
         event_end5 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5
                              & lead(hrs_till_next_event, 2) <= 1.5 & lead(hrs_till_next_event, 3) <= 1.5, 
                              lead(event_end, 4), event_end4),
         event_end6 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5
                              & lead(hrs_till_next_event, 2) <= 1.5 & lead(hrs_till_next_event, 3) <= 1.5
                              & lead(hrs_till_next_event, 4) <= 1.5, 
                              lead(event_end, 5), event_end5)) %>% 
  distinct(event_end6, .keep_all = TRUE) %>% 
  mutate(event_length_min = (event_end6 - start),
         event_length_min = as.numeric(event_length_min/60),
         hrs_till_next_event = (lead(start)-event_end6),
         hrs_till_next_event = as.numeric(hrs_till_next_event)/3600) %>% 
  filter(!is.na(event_length_min)) %>% 
  group_by(area, home, winter_id, date_sums) %>% 
  mutate(event_start_datetime = start,
         event_end_datetime = event_end6,
         sums_event = as.numeric(1)) %>% 
  ungroup() %>% 
  select(area, home, home_winter_id, winter_id, date_sums, 
         event_start_datetime, event_end_datetime, sums_event,
         event_length_min, hrs_till_next_event)

summary(sums_events_30)

count_sums_events_30 <- sums_events_30 %>% 
  count(area, home_winter_id)
count_sums_events_30

mean_events_30 <- count_sums_events_30 %>% 
  group_by(area) %>% 
  summarize(mean(n))
mean_events_30
```