### 2021-12-09 letter_results.Rmd
results report updates

---
output: 
  word_document:
    reference_docx: results_report_template.docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE,
                      fig.height = 2.75, fig.width = 8)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
               "#0072B2", "#D55E00", "#CC79A7")
```

```{r}
# Load data

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

analysis_data <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
  mutate(cohort = factor(cohort, levels = c("2015", "2016", "2017", "2018", "2019"))) %>% 
  filter(area == "NPT") %>% 
  filter(pm_mean_winter < 200) %>% 
  mutate("Winter:" = factor(winter_id, levels = c(1, 2),
                                  labels = c("Pre-Intervention", "Post-Intervention")),
         treatment = factor(treatment, 
                            levels = c("Placebo", "Education", "Filter"),
                            labels = c("Placebo Group", "Education Group", "Filter Group")),
         pm_mean_winter = round(pm_mean_winter, digits = 0),
         mp_mean_winter = round(mp_mean_winter, digits = 0))

data_pm <- adult_data %>% 
  distinct(winter_id, .keep_all = T)

data_mp <- adult_data %>% 
  distinct(winter_id, .keep_all = T)

data_sbp <- adult_data %>% 
  group_by(winter_id) %>% 
  mutate(bp_sys_mean = round(mean(bp_sys, na.rm = T), digits = 0)) %>% 
  ungroup() %>% 
  distinct(winter_id, .keep_all = T)

data_dbp <- adult_data %>% 
  group_by(winter_id) %>% 
  mutate(bp_dia_mean = round(mean(bp_dia, na.rm = T), digits = 0)) %>% 
  ungroup() %>% 
  distinct(winter_id, .keep_all = T)

data_fvc <- adult_data %>% 
  group_by(winter_id) %>% 
  mutate(fvc_mean = round(mean(fvc, na.rm = T), digits = 1)) %>% 
  ungroup() %>% 
  distinct(winter_id, .keep_all = T)

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
               "#0072B2", "#D55E00", "#CC79A7")
```

![](npt_logo.png)

## Your Study Results

Your results are indicated by the triangles in the figures. The label next to each 
triangle is your result from the study. The smaller circles in the figure 
are results from the rest of the participants from the community. Please see the 
fact sheet handout for frequently asked questions related to your results.

### Air Pollution Results

#### Indoor PM2.5  
The smoke from wood burning is made up of a complex mixture of gases and particles. 
The biggest health threat from smoke is from fine particles, also called fine particulate matter or PM2.5. 
PM2.5 is measured in micrograms per cubic meter (ug/m3) of air. In the study, we 
measured Indoor PM2.5 using the stationary monitor in your home near the wood stove.  
```{r}
indoor_pm_plot <- analysis_data %>% 
  filter(!is.na(pm_mean_winter)) %>% 
  filter(!is.na(treatment)) %>% 
  distinct(home, winter_id, .keep_all = T) %>% 
  ggplot() + 
    #geom_boxplot(aes(treatment, pm_mean_winter, fill = `Winter:`, size = 1), 
     #            lwd = 1.2, colour = "black", 
      #           fatten = 1, outlier.size = 1.5, width = 0.35) +
    geom_point(aes(treatment, pm_mean_winter, fill = `Winter:`), 
               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
    geom_label(aes(treatment, pm_mean_winter, label = pm_mean_winter, 
                   color = `Winter:`), 
               data = data_pm, size = 5, position=position_dodge(1.6)) +
    geom_point(aes(treatment, pm_mean_winter, fill = `Winter:`), 
               data = data_pm, shape = 24, size = 4, stroke = 3, 
               position=position_dodge(0.9)) +
    labs(y = "Indoor PM2.5") +
    theme_minimal() +
    theme(plot.title = element_text(size = 18, colour = "black"),
          strip.text = element_text(size = 16, colour = "black"),
          axis.text.x = element_text(size = 16, colour = "black", 
                                     vjust = 0.5),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16, colour = "black"),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          legend.title = element_text(size = 16, colour = "black"),
          legend.text = element_text(size = 16, colour = "black"),
          legend.position = "top",
          panel.grid.major.x = element_blank(),
          axis.ticks = element_blank()) +
    scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
    scale_color_manual(values = c("#E69F00", "#56B4E9")) 
indoor_pm_plot
```

#### Personal PM2.5  
Personal PM2.5 was measured using the monitor in a small bag that you wore during the study.  
```{r}
personal_pm_plot <- analysis_data %>% 
  filter(!is.na(mp_mean_winter)) %>% 
  filter(!is.na(treatment)) %>% 
  distinct(home, winter_id, .keep_all = T) %>% 
  ggplot() + 
    geom_point(aes(treatment, mp_mean_winter, fill = `Winter:`), 
               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
    geom_label(aes(treatment, mp_mean_winter, label = mp_mean_winter, 
                   color = `Winter:`), 
               data = data_mp, size = 5, position=position_dodge(1.6)) +
    geom_point(aes(treatment, mp_mean_winter, fill = `Winter:`), 
               data = data_mp, shape = 24, size = 4, stroke = 3, 
               position=position_dodge(0.9)) +
    labs(y = "Personal PM2.5") +
    theme_minimal() +
    theme(plot.title = element_text(size = 18, colour = "black"),
          strip.text = element_text(size = 16, colour = "black"),
          axis.text.x = element_text(size = 16, colour = "black", 
                                     vjust = 0.5),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16, colour = "black"),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          legend.title = element_text(size = 16, colour = "black"),
          legend.text = element_text(size = 16, colour = "black"),
          legend.position = "top",
          panel.grid.major.x = element_blank(),
          axis.ticks = element_blank()) +
    scale_fill_manual(values = c("#E69F00", "#56B4E9")) 
personal_pm_plot
```

\pagebreak

### Health Results

#### Systolic Blood Pressure  
Blood pressure is an important indicator of heart health and is measured in 
millimeters of mercury (mmHg). Systolic blood pressure is the top number in your blood pressure, 
and a value of 130 or above is considered high.  
```{r}
sbp_plot <- analysis_data %>% 
  filter(!is.na(bp_sys)) %>% 
  filter(!is.na(treatment)) %>% 
  group_by(adult_id_char, winter_id) %>% 
  mutate(bp_sys_mean = mean(bp_sys, na.rm = T)) %>% 
  ungroup() %>% 
  distinct(adult_id_char, winter_id, .keep_all = T) %>% 
  ggplot() + 
    geom_point(aes(treatment, bp_sys_mean, fill = `Winter:`), 
               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
    geom_label(aes(treatment, bp_sys_mean, label = bp_sys_mean, 
                   color = `Winter:`), 
               data = data_sbp, size = 5, position=position_dodge(1.7)) +
    geom_point(aes(treatment, bp_sys_mean, fill = `Winter:`), 
               data = data_sbp, shape = 24, size = 4, stroke = 3, 
               position=position_dodge(0.9)) +
    labs(y = "Systolic blood pressure") +
    theme_minimal() +
    theme(plot.title = element_text(size = 18, colour = "black"),
          strip.text = element_text(size = 16, colour = "black"),
          axis.text.x = element_text(size = 16, colour = "black", 
                                     vjust = 0.5),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16, colour = "black"),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          legend.title = element_text(size = 16, colour = "black"),
          legend.text = element_text(size = 16, colour = "black"),
          legend.position = "top",
          panel.grid.major.x = element_blank(),
          axis.ticks = element_blank()) +
    scale_fill_manual(values = c("#E69F00", "#56B4E9")) 
sbp_plot
```

#### Diastolic Blood Pressure  
Diastolic blood pressure is the bottom number in your blood pressure, and a value 
of 80 or above is considered high.  
```{r}
dbp_plot <- analysis_data %>% 
  filter(!is.na(bp_dia)) %>% 
  filter(!is.na(treatment)) %>% 
  group_by(adult_id_char, winter_id) %>% 
  mutate(bp_dia_mean = mean(bp_dia, na.rm = T)) %>% 
  ungroup() %>% 
  distinct(adult_id_char, winter_id, .keep_all = T) %>% 
  ggplot() + 
    geom_point(aes(treatment, bp_dia_mean, fill = `Winter:`), 
               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
    geom_label(aes(treatment, bp_dia_mean, label = bp_dia_mean, 
                   color = `Winter:`), 
               data = data_dbp, size = 5, position=position_dodge(1.6)) +
    geom_point(aes(treatment, bp_dia_mean, fill = `Winter:`), 
               data = data_dbp, shape = 24, size = 4, stroke = 3, 
               position=position_dodge(0.9)) +
    labs(y = "Diastolic blood pressure") +
    theme_minimal() +
    theme(plot.title = element_text(size = 18, colour = "black"),
          strip.text = element_text(size = 16, colour = "black"),
          axis.text.x = element_text(size = 16, colour = "black", 
                                     vjust = 0.5),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16, colour = "black"),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          legend.title = element_text(size = 16, colour = "black"),
          legend.text = element_text(size = 16, colour = "black"),
          legend.position = "top",
          panel.grid.major.x = element_blank(),
          axis.ticks = element_blank()) +
    scale_fill_manual(values = c("#E69F00", "#56B4E9")) 
dbp_plot
```

#### Forced Vital Capacity    
Forced vital capacity is the total lung volume that you exhaled when your lung 
function was tested. Forced vital capacity is measures in Liters (L) of air.  
```{r}
fvc_plot <- analysis_data %>% 
  filter(!is.na(fvc)) %>% 
  filter(!is.na(treatment)) %>% 
  group_by(adult_id_char, winter_id) %>% 
  mutate(fvc_mean = mean(fvc, na.rm = T)) %>% 
  ungroup() %>% 
  distinct(adult_id_char, winter_id, .keep_all = T) %>% 
  ggplot() + 
    geom_point(aes(treatment, fvc_mean, fill = `Winter:`), 
               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
    geom_label(aes(treatment, fvc_mean, label = fvc_mean, 
                   color = `Winter:`), 
               data = data_fvc, size = 5, position=position_dodge(1.6)) +
    geom_point(aes(treatment, fvc_mean, fill = `Winter:`), 
               data = data_fvc, shape = 24, size = 4, stroke = 3, 
               position=position_dodge(0.9)) +
    labs(y = "Forced vital capacity") +
    theme_minimal() +
    theme(plot.title = element_text(size = 18, colour = "black"),
          strip.text = element_text(size = 16, colour = "black"),
          axis.text.x = element_text(size = 16, colour = "black", 
                                     vjust = 0.5),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16, colour = "black"),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          legend.title = element_text(size = 16, colour = "black"),
          legend.text = element_text(size = 16, colour = "black"),
          legend.position = "top",
          panel.grid.major.x = element_blank(),
          axis.ticks = element_blank()) +
    scale_fill_manual(values = c("#E69F00", "#56B4E9")) 
fvc_plot
```