### 2022-01-28 letter_results.Rmd
Update figures for NPT community results presentation

---
output: 
  word_document:
    reference_docx: results_report_template.docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE,
                      fig.height = 2.75, fig.width = 8)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
               "#0072B2", "#D55E00", "#CC79A7")
```

```{r}
# Load data

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

analysis_data <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
  mutate(cohort = factor(cohort, levels = c("2015", "2016", "2017", "2018", "2019"))) %>% 
  filter(area == "NPT") %>% 
  #filter(pm_mean_winter < 200) %>% 
  #filter(home == "NPE5301") %>% 
  mutate("Winter:" = factor(winter_id, levels = c(1, 2),
                                  labels = c("1st Winter/Pre-Intervention", "2nd Winter/Post-Intervention")),
         treatment = factor(treatment, 
                            levels = c("Placebo", "Education", "Filter"),
                            labels = c("Placebo Group", "Education Group", "Filter Group")),
         pm_mean_winter = round(pm_mean_winter, digits = 0),
         mp_mean_winter = round(mp_mean_winter, digits = 0))

data_pm <- adult_data %>% 
  distinct(winter_id, .keep_all = T)

data_mp <- adult_data %>% 
  distinct(winter_id, .keep_all = T)

data_sbp <- adult_data %>% 
  group_by(winter_id) %>% 
  mutate(bp_sys_mean = round(mean(bp_sys, na.rm = T), digits = 0)) %>% 
  ungroup() %>% 
  distinct(winter_id, .keep_all = T)

data_dbp <- adult_data %>% 
  group_by(winter_id) %>% 
  mutate(bp_dia_mean = round(mean(bp_dia, na.rm = T), digits = 0)) %>% 
  ungroup() %>% 
  distinct(winter_id, .keep_all = T)

data_fvc <- adult_data %>% 
  group_by(winter_id) %>% 
  mutate(fvc_mean = round(mean(fvc, na.rm = T), digits = 1)) %>% 
  ungroup() %>% 
  distinct(winter_id, .keep_all = T)

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
               "#0072B2", "#D55E00", "#CC79A7")
```

![](npt_logo2.png)

### Your Nimiipuu EldersAIR Study Results  
Your results may be different from other participants' results depending on 1) how often and/or how efficiently you used your stove and 2) your study group and how consistently you used the education or air filter interventions. See the **Overview** for a detailed description of the three study groups and other frequently asked questions related to your results.

**Graph:** The triangles in the graph show 1) your study results and 2) which study group you were in. Next to each triangle is the numerical value for each measurement. The small circles represent the other Nimiipuu participant results.

**Terms:** Pre-intervention (or baseline) - your first winter in the study with air sampling and health tests, but no intervention. Post-intervention – your second winter in the study that occurred AFTER your randomized assignment to an intervention group.

#### Indoor PM2.5 Air Pollution  
We measured indoor PM2.5 using the stationary DustTrak monitor near your wood stove.  
```{r, fig.height = 2.5, fig.width = 8}
indoor_pm_plot <- analysis_data %>% 
  filter(!is.na(pm_mean_winter)) %>% 
  filter(!is.na(treatment)) %>% 
  distinct(home, winter_id, .keep_all = T) %>% 
  ggplot() + 
    #geom_boxplot(aes(treatment, pm_mean_winter, fill = `Winter:`, size = 1), 
     #            lwd = 1.2, colour = "black", 
      #           fatten = 1, outlier.size = 1.5, width = 0.35) +
    #geom_hline(yintercept = 35, color = "red", linetype = 2) + 
    geom_point(aes(treatment, pm_mean_winter, fill = `Winter:`), 
               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
    geom_label(aes(treatment, pm_mean_winter, label = pm_mean_winter, 
                   color = `Winter:`), 
               data = data_pm, size = 5, position=position_dodge(1.6)) +
    geom_point(aes(treatment, pm_mean_winter, fill = `Winter:`), 
               data = data_pm, shape = 24, size = 4, stroke = 3, 
               position=position_dodge(0.9)) +
    labs(y = "Indoor PM2.5", fill = "", color = "") +
    theme_minimal() +
    theme(plot.title = element_text(size = 18, colour = "black"),
          strip.text = element_text(size = 16, colour = "black"),
          axis.text.x = element_text(size = 16, colour = "black", 
                                     vjust = 0.5),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16, colour = "black"),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          legend.title = element_text(size = 16, colour = "black"),
          legend.text = element_text(size = 16, colour = "black"),
          legend.position = "top",
          panel.grid.major.x = element_blank(),
          axis.ticks = element_blank()) +
    scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
    scale_color_manual(values = c("#E69F00", "#56B4E9")) 
indoor_pm_plot
```

#### Personal PM2.5 Air Pollution  
We measured personal PM2.5 using the portable MicroPem monitor in the small pouch.  
```{r, fig.height = 2.5, fig.width = 8}
personal_pm_plot <- analysis_data %>% 
  filter(!is.na(mp_mean_winter)) %>% 
  filter(!is.na(treatment)) %>% 
  distinct(home, winter_id, .keep_all = T) %>% 
  ggplot() + 
    #geom_hline(yintercept = 35, color = "red", linetype = 2) + 
    geom_point(aes(treatment, mp_mean_winter, fill = `Winter:`), 
               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
    geom_label(aes(treatment, mp_mean_winter, label = mp_mean_winter, 
                   color = `Winter:`), 
               data = data_mp, size = 5, position=position_dodge(1.6)) +
    geom_point(aes(treatment, mp_mean_winter, fill = `Winter:`), 
               data = data_mp, shape = 24, size = 4, stroke = 3, 
               position=position_dodge(0.9)) +
    labs(y = "Personal PM2.5", fill = "", color = "") +
    theme_minimal() +
    theme(plot.title = element_text(size = 18, colour = "black"),
          strip.text = element_text(size = 16, colour = "black"),
          axis.text.x = element_text(size = 16, colour = "black", 
                                     vjust = 0.5),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16, colour = "black"),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          legend.title = element_text(size = 16, colour = "black"),
          legend.text = element_text(size = 16, colour = "black"),
          legend.position = "top",
          panel.grid.major.x = element_blank(),
          axis.ticks = element_blank()) +
    scale_fill_manual(values = c("#E69F00", "#56B4E9")) 
personal_pm_plot
```

\pagebreak

#### Systolic Blood Pressure  
Blood pressure is an important indicator of heart health and is measured in millimeters of mercury (mmHg). Systolic blood pressure is the top number in your blood pressure; a value of 130 or above is considered high and is indicated by the dotted RED line in the graph.  
```{r}
sbp_plot <- analysis_data %>% 
  filter(!is.na(bp_sys)) %>% 
  filter(!is.na(treatment)) %>% 
  group_by(adult_id_char, winter_id) %>% 
  mutate(bp_sys_mean = mean(bp_sys, na.rm = T)) %>% 
  ungroup() %>% 
  distinct(adult_id_char, winter_id, .keep_all = T) %>% 
  ggplot() + 
    geom_hline(yintercept = 130, color = "red", linetype = 2) + 
    geom_point(aes(treatment, bp_sys_mean, fill = `Winter:`), 
               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
    geom_label(aes(treatment, bp_sys_mean, label = bp_sys_mean, 
                   color = `Winter:`), 
               data = data_sbp, size = 5, position=position_dodge(1.7)) +
    geom_point(aes(treatment, bp_sys_mean, fill = `Winter:`), 
               data = data_sbp, shape = 24, size = 4, stroke = 3, 
               position=position_dodge(0.9)) +
    labs(y = "Systolic blood pressure", fill = "", color = "") +
    theme_minimal() +
    theme(plot.title = element_text(size = 18, colour = "black"),
          strip.text = element_text(size = 16, colour = "black"),
          axis.text.x = element_text(size = 16, colour = "black", 
                                     vjust = 0.5),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16, colour = "black"),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          legend.title = element_text(size = 16, colour = "black"),
          legend.text = element_text(size = 16, colour = "black"),
          legend.position = "top",
          panel.grid.major.x = element_blank(),
          axis.ticks = element_blank()) +
    scale_fill_manual(values = c("#E69F00", "#56B4E9")) 
sbp_plot
```

#### Diastolic Blood Pressure  
The bottom number in your blood pressure; a value of 80 or above is considered high and is indicated by the dotted RED line in the graph.  
```{r}
dbp_plot <- analysis_data %>% 
  filter(!is.na(bp_dia)) %>% 
  filter(!is.na(treatment)) %>% 
  group_by(adult_id_char, winter_id) %>% 
  mutate(bp_dia_mean = mean(bp_dia, na.rm = T)) %>% 
  ungroup() %>% 
  distinct(adult_id_char, winter_id, .keep_all = T) %>% 
  ggplot() + 
    geom_hline(yintercept = 80, color = "red", linetype = 2) + 
    geom_point(aes(treatment, bp_dia_mean, fill = `Winter:`), 
               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
    geom_label(aes(treatment, bp_dia_mean, label = bp_dia_mean, 
                   color = `Winter:`), 
               data = data_dbp, size = 5, position=position_dodge(1.6)) +
    geom_point(aes(treatment, bp_dia_mean, fill = `Winter:`), 
               data = data_dbp, shape = 24, size = 4, stroke = 3, 
               position=position_dodge(0.9)) +
    labs(y = "Diastolic blood pressure", fill = "", color = "") +
    theme_minimal() +
    theme(plot.title = element_text(size = 18, colour = "black"),
          strip.text = element_text(size = 16, colour = "black"),
          axis.text.x = element_text(size = 16, colour = "black", 
                                     vjust = 0.5),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16, colour = "black"),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          legend.title = element_text(size = 16, colour = "black"),
          legend.text = element_text(size = 16, colour = "black"),
          legend.position = "top",
          panel.grid.major.x = element_blank(),
          axis.ticks = element_blank()) +
    scale_fill_manual(values = c("#E69F00", "#56B4E9")) 
dbp_plot
```

#### Forced Vital Capacity    
The total lung volume that you exhaled when your lung function was tested with the EasyOne Air Spirometer. Forced vital capacity is measures in Liters (L) of air.   
```{r}
fvc_plot <- analysis_data %>% 
  filter(!is.na(fvc)) %>% 
  filter(!is.na(treatment)) %>% 
  group_by(adult_id_char, winter_id) %>% 
  mutate(fvc_mean = mean(fvc, na.rm = T)) %>% 
  ungroup() %>% 
  distinct(adult_id_char, winter_id, .keep_all = T) %>% 
  ggplot() + 
    geom_point(aes(treatment, fvc_mean, fill = `Winter:`), 
               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
    geom_label(aes(treatment, fvc_mean, label = fvc_mean, 
                   color = `Winter:`), 
               data = data_fvc, size = 5, position=position_dodge(1.6)) +
    geom_point(aes(treatment, fvc_mean, fill = `Winter:`), 
               data = data_fvc, shape = 24, size = 4, stroke = 3, 
               position=position_dodge(0.9)) +
    labs(y = "Forced vital capacity", fill = "", color = "") +
    theme_minimal() +
    theme(plot.title = element_text(size = 18, colour = "black"),
          strip.text = element_text(size = 16, colour = "black"),
          axis.text.x = element_text(size = 16, colour = "black", 
                                     vjust = 0.5),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16, colour = "black"),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          legend.title = element_text(size = 16, colour = "black"),
          legend.text = element_text(size = 16, colour = "black"),
          legend.position = "top",
          panel.grid.major.x = element_blank(),
          axis.ticks = element_blank()) +
    scale_fill_manual(values = c("#E69F00", "#56B4E9")) 
fvc_plot
```